### 2008-07-10 doc/logparser-database-schema.tex
Corrections from my last proofreading.

\section{Database Schema}
\label{database schema}

The database is an integral part of the parser presented here: it stores
the rules and the data gleaned by applying those rules to Postfix log
files.  Understanding the database schema is important in understanding the
actions of the parser, and essential to developing further applications
which utilise the data gathered.

\subsection{Introduction}

The database schema can be conceptually divided in two: the rules, which
are used to parse log files, and the data saved from the parsing of log
files.  Rules have the fields required to parse the log lines, extract data
to be saved, and the action to be executed; they also have several fields
which aid the user in understanding the meaning of the log lines parsed by
each rule.  The rules are described in detail in \sectionref{rules} but the
fields are covered in \sectionref{rule attributes}.

The data saved from parsing the log files is also divided into two tables
as described below: connections and results.  The connections table
contains a row for every mail accepted and every connection where there was
a rejection; the individual fields will be described in
\sectionref{connections table}.  There will be at least one row in the
results table for each row in the connections table; the fields will be
covered in detail in \sectionref{results table}.

An important but easily overlooked benefit of storing the rules in the
database is the connection between rules and results: if more information
is required when examining a result, the rule which produced the result is
available for inspection --- each result references the rule which created
it.  There is no ambiguity about which rule resulted in a particular
result, eliminating one potential source of confusion.

\subsection{Rules table}

\label{rule attributes}

Rules are discussed in detail in \sectionref{rules}, but the structure of
the rules table is covered here.  Rules are created by the user, not the
parser, and will not be modified by the parser (except for the hits and
hits\_total fields).  Rules parse the individual log lines, extracting data
to be saved in the connections and results tables, and specifying the
action to take for that log line.

The following attributes are defined for each rule:

\begin{description}

    \item [id] A unique identifier for each rule which other tables can use
        to refer to a specific rule.

    \item [name] A short name for the rule.

    \item [description] Something must have occurred to cause Postfix to
        log each line (e.g.\ a remote client connecting causes a connection
        line to be logged).  This field describes the event causing the log
        lines this rule matches.

    \item [restriction\_name] The restriction which caused the mail to be
        rejected.  Only applicable to rules which have an action of
        \texttt{REJECTION}, other rules should have an empty string.

    \item [postfix\_action] This is the action Postfix must have taken to
        generate this log line.  This field is mostly ignored, but two
        values (IGNORED and INFO) have special meaning, as described below
        in the list of typical values.\label{postfix_action}

        \begin{description}

            \item [ACCEPTED] Postfix has accepted a mail, and will
                subsequently attempt to deliver it.

            \item [BOUNCED] The mail has bounced, due to a mail loop,
                delivery failure, or five day timeout.

            \item [DELETED] The mail was deleted from the queue by an
                administrator.

            \item [DISCARDED] Postfix discarded the mail it was in the
                process of accepting, because it was either larger than the
                limit set by the administrator, or the client timed out or
                disconnected.

            \item [EXPIRED] The mail has been in the queue for five
                days\footnote{As with much of Postfix's behaviour, this is
                the default value but can be changed by the administrator
                if they choose.} without successful delivery.  A bounce
                mail will be generated and sent to the sender address.

            \item [INFO] Represents an unspecified intermediate action that
                the parser is not interested in per se, but which does log
                useful information, supplementing other log lines.

            \item [IGNORED] An action which is not only uninteresting in
                itself, but which also provides no useful data.

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all existing \daemon{smtpd}
                processes will be terminated.  This does not negatively
                impact on the log files or mail queued by Postfix for
                delivery.

            \item [PROCESSING] \daemon{cleanup} is processing a mail
                (see~\cite{postfix-cleanup} for details).

            \item [REJECTED] Postfix rejected a command from the remote
                client, causing at least one recipient to be rejected.

            \item [SENT] Postfix has successfully sent a mail.

        \end{description}

        Uninteresting log lines are parsed so that any lines the parser
        is not capable of handling become immediately obvious errors.

    \item [program] The program (\daemon{smtpd}, \daemon{qmgr}, etc.) whose
        log lines the rule applies to.  This avoids needlessly trying rules
        which will not match the log line, or worse, might match
        unintentionally.  Rules whose program is \texttt{*} will be tried
        against any log lines which are not parsed by program specific
        rules.

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.

        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        caching compiled \regexes{} covered in \sectionref{Caching each
        regex}.

    \item [result\_cols, connection\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \newline \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right hand side.  Data specified by
        \texttt{result\_cols} will be entered in the \texttt{results}
        table, while \texttt{connection\_cols} populates the
        \texttt{connection} table.  The list of acceptable variable names
        is:

        \texttt{connection\_cols: client\_hostname, client\_ip, server\_ip,
        \newline \tab{} server\_hostname} and \texttt{helo.\newline}
        \texttt{result\_cols: sender, recipient, smtp\_code, message\_id,
        \newline \tab{} size,} and \texttt{data}

        Additionally \texttt{child} and \texttt{pid} are accepted and used
        respectively by the \texttt{TRACK}, \texttt{BOUNCE} and
        \texttt{SMTPD\_DIED} actions, but do not cause data to be entered
        in the database.

    \item [result\_data, connection\_data] Sometimes rules need to supply a
        piece of data which is not present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format and allowed
        variables are the same as for \texttt{result\_cols} and
        \texttt{connection\_cols}, except that arbitrary
        data\footnote{Commas and semi-colons cannot be escaped and thus
        cannot be used.  This feature is intended for use with small
        amounts of data rather than large amounts in any one rule, so
        dealing with escape sequences was deemed unnecessary.} is permitted
        on the right hand side of the assignment.

    \item [action] The action that will be invoked when this rule matches a
        log line; a full list of actions and the parameters they are
        invoked with can be found in \sectionref{actions-in-detail}.

    \item [queueid] Specifies the capturing field in the \regex{}
        containing the queueid, or zero if the log line does not contain a
        queueid.  Many log lines will not contain a queueid, e.g.\
        rejections logged before a mail has been accepted (a queueid will
        not have been allocated), or log lines which are not tied to one
        particular mail.

    \item [hits] This counter is maintained for every rule and incremented
        each time the rule successfully matches.  At the start of each run
        the program sorts the rules in descending order of hits, and at the
        end of the run updates every rule's hits field in the database.
        Assuming that the distribution of log lines is reasonably
        consistent between log files, rules matching more commonly
        occurring log lines will be tried before rules matching less
        commonly occurring log lines, reducing the program's execution
        time.  Rule ordering for efficiency is discussed in
        \sectionref{rule ordering for efficiency}.

    \item [hits\_total] The total number of hits for this rule over all
        runs of the parser.

    \item [priority] This is the user-configurable companion to hits: when
        the list of rules is sorted, priority overrides hits.  This allows
        more specific rules to take precedence over more general rules
        (described in \sectionref{overlapping rules}).

    \item [cluster\_group] A reference to the \texttt{cluster\_group}
        table.  That table is used by the Decision Tree algorithm described
        in a separate document.

\end{description}

\subsection{Connections table}

\label{connections table}

Every accepted mail and every connection where there was a rejection will
have a single entry in the connections table containing the following
fields:

\begin{description}

    \item [id] This field uniquely identifies the row.

    \item [server\_ip] The \IP{} address (IPv4 or IPv6) of the server: the
        local address when receiving mail, the remote address when sending
        mail.

    \item [server\_hostname] The hostname of the server, it will be
        \texttt{unknown} if the \IP{} address could not be resolved to a
        hostname via \DNS{}\@.

    \item [client\_ip] The client \IP{} address (IPv4 or IPv6): the remote
        address when receiving mail, the local address when sending mail.

    \item [client\_hostname] The hostname of the client, it will be
        \texttt{unknown} if the \IP{} address could not be resolved to a
        hostname via \DNS{}\@.

    \item [helo] The hostname used in the HELO command.  The HELO hostname
        occasionally changes during a connection, presumably because spam
        or virus senders think it is a good idea.  By default Postfix only
        logs the HELO hostname when it rejects an \SMTP{} command, but it
        is quite easy to rectify this, as described in \sectionref{logging
        helo}.

    \item [queueid] The queueid of the mail if the connection represents an
        accepted mail, or \texttt{NOQUEUE} otherwise.

    \item [start] The timestamp of the first log line, in seconds since the
        epoch (explained in the glossary, \sectionref{Glossary}).

    \item [end] The timestamp of the last log line, in seconds since the
        epoch.

\end{description}

\subsection{Results table}

\label{results table}

Every interesting\footnote{All log lines are interesting unless the action
of the rule which parses them is \texttt{INFO} or \texttt{IGNORE}, as
described in \sectionref{actions-in-detail}.} log line will have an entry
in the results table, e.g.\ rejecting an \SMTP{} command, delivering a
mail, or bouncing a mail (see \sectionref{postfix_action} for typical
values of \texttt{postfix\_action}).  Each row is associated with a single
connection, though there may be many results per connection.

\begin{description}

    \item [connection\_id] The id of the row in the connections table this
        result is associated with.

    \item [rule\_id] The id of the entry in the rules table which matched
        the log line and created this result.

    \item [id] A unique identifier for this result.

    \item [warning] Postfix can be configured to log a warning instead of
        enforcing a restriction that would reject an \SMTP{} command --- a
        facility that is quite useful for testing new restrictions.  This
        field will be 1 if the log line parsed was a warning rather than a
        real rejection, or 0 for a real rejection.  This field should be
        ignored if the result is not a rejection, i.e.\ the action field of
        the associated rule is not \texttt{REJECTION}.

    \item [smtp\_code] The \SMTP{} code associated with the log line.  In
        general an \SMTP{} code is only present for a rejection or final
        delivery; results initially missing an \SMTP{} code will duplicate
        the \SMTP{} code of other results in the connection.  Some final
        delivery log lines do not contain an \SMTP{} code: in those cases
        the code is specified by the rule, based on the success or failure
        represented by the log line.

    \item [sender] The sender's email address.  Because multiple mails may
        be delivered during one connection, there may be different sender
        addresses in the results for one connection; however there should
        not be different sender addresses in the results for one email.
        Mails sent over the same connection can be distinguished by their
        queueid.

    \item [recipient] The recipient address; there may be multiple
        recipient addresses per mail, but only one per result.

    \item [size] The size of the mail; it will only be present for
        delivered mails.

    \item [message\_id] The message-id of the accepted mail, or
        \texttt{NULL} if no mail was accepted.

    \item [data] A field available for anything not covered by other
        fields, e.g.\ the rejection message from an \DNSBL{}\@.

    \item [timestamp] The time the log line was logged, in seconds since
        the epoch.

\end{description}

\subsection{Summary}

The table containing the rules used by the parser and both tables
containing the data extracted from the Postfix log files were described,
with the purpose of each field discussed in detail.  A clear,
comprehensible schema is essential when using the extracted data; it is
more important when using the data than when storing it, because storing
the data is a write-once operation, whereas utilising the data requires
frequent searching, sorting and manipulation of the data to produce
customised reports and/or statistics.