### 2008-12-03 doc/sgai-2008-slides.tex
Lots of improvements to the slides.

\documentclass{beamer}
%\mode<presentation>
%{
%  \usetheme{Warsaw}
%  % or ...
%
%  \setbeamercovered{transparent}
%  % or whatever (possibly just delete it)
%}
\useoutertheme{infolines}
\defbeamertemplate*{headline}{JT theme}{}

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\newcommand{\timingnote}[1]{%
    \textbf{#1}%
}

\title{A User-Extensible and Adaptable Parser Architecture}

\author{John Tobin \and Carl Vogel}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Trinity College]
{
  School of Computer Science and Statistics\\
  Trinity College, University of Dublin
}

\date[SGAI 2008]{Twenty-eighth SGAI International Conference on Artificial Intelligence}

\subject{Theoretical Computer Science}

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

\begin{document}

\begin{frame}
    \timingnote{30 seconds}
    \titlepage{}
\end{frame}

\begin{frame}{Outline of this talk}
    \timingnote{30 seconds --- 1 minute}
    \tableofcontents{}
\end{frame}

% Structuring a talk is a difficult task and the following structure
% may not be suitable. Here are some rules that apply for this
% solution:

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.

% - A conference audience is likely to know very little of what you
%   are going to talk about. So *simplify*!
% - In a 20min talk, getting the main ideas across is hard
%   enough. Leave out details, even if it means being less precise than
%   you think necessary.
% - If you omit details that are vital to the proof/implementation,
%   just say so once. Everybody will be happy with that.

\section{Background}

\begin{frame}{Background}

    \timingnote{1--2 minutes}

    This parser architecture was developed as part of a larger project to
    improve anti-spam techniques.

    \begin{description}

        \item [Currently] Most anti-spam techniques are content-based, and
            require each mail to be accepted before determining its spam
            status.

        \item [However] Rejecting mail at SMTP-time is more efficient, and
            senders of non-spam mails that are incorrectly rejected will be
            informed immediately.

        \item [Project Aim] To improve SMTP-time anti-spam techniques, by
            providing a platform for reasoning about anti-spam techniques.

        \item [Method] Parse Postfix log files and analyse the data
            extracted.

        \item [XXX] This requires a parser for Postfix log files.

    \end{description}

\end{frame}

\section{Input characteristics}

\begin{frame}{Input characteristics}

    \timingnote{1 minute}

    \begin{itemize}

        \item Postfix log lines change over time and from site to site, so
            it must be easy for the end user to parse new log lines.

        \item The processing required for a type of log line (e.g.\ a
            remote server connecting to send mail) rarely changes, but it
            can be quite complex.

        \item By separating parsing of log lines from processing of log
            lines, we can make parsing new or changed log lines as easy as
            possible, while still allowing for complicated processing when
            necessary.

    \end{itemize}

\end{frame}

\section{Architecture}

\begin{frame}{Architecture}

    \timingnote{2 minutes}

    The architecture is divided into three parts:

    \begin{description}

        \item [Framework] The framework handles loading rules and providing
            services (XXX BETTER WORD\@?) to actions, e.g.\ data storage,
            inter-action communication, XXX ANOTHER EXAMPLE\@.  For each
            input log line, the framework tries each rule in turn until one
            matches, then it invokes the action specified by the rule.

        \item [Actions] Each action performs the processing required, e.g.\
            XXX GIVE SOME EXAMPLES\@.

        \item [Rules] Each rule matches one type of log line, e.g.\ the log
            line resulting from client connecting, a mail successfully
            delivered, or a mail rejected by an anti-spam measure.

    \end{description}

\end{frame}

\section{Features of the Architecture}

\begin{frame}{Features of the Architecture}

    \timingnote{2 minutes}

    \begin{itemize}

        \item Cascaded Parsing.  XXX SHOULD I HAVE A REFERENCE\@?

        \item The parser does not need to be recompiled when rules are
            changed, allowing: quicker testing of modified rules; different
            rulesets can easily be used; combining two or more rulesets.

        \item Rules can be changed or added by actions while the parser is
            running, allowing dynamically changing parsing, e.g.\ macros.
            This feature must be used with care.

        \item The architecture is capable of parsing inputs that are not
            completely understood, e.g.\ when developing a parser: the
            framework will warn about unparsed inputs but will continue on
            parsing as best it can.

        \item Parsing can optionally be stateful, e.g.\ parsing a
            programming language.

        \item XXX MORE FEATURES\@?

    \end{itemize}

\end{frame}

\section{Efficiency}

\begin{frame}{Efficiency}

    \timingnote{2 minutes}

    \begin{itemize}

        \item Median throughput when parsing 93 log files is 80.85~MB
            (480,569 log lines) parsed per minute.

        \item Rules are tagged with the name of a Postfix daemon, and will
            only be used to parse log lines produced by that daemon.

        \item Rules are sorted by how frequently they match, resulting in a
            mean reduction of 14.78\% in parsing time compared to using
            unsorted rules.

        \item The architecture scales extremely well as the number of rules
            increases: a 46.05\% increase in Postfix ruleset resulted in a
            mean increase of 0.63\% in parsing time.

        \item Parsing time scales linearly with input size, but is affected
            by input content.

        \item XXX MORE\@?

    \end{itemize}

\end{frame}

\section{Results}

\begin{frame}{Results}

    \timingnote{1--2 minutes}

    \begin{itemize}

        \item 100\% coverage of log lines.

        \item 100\% coverage of all mails accepted, delivered, or rejected.

        \item 169 rules: the top 10\% match 85.03\% of log lines, the
            remainder tailing off similar to a Power Law distribution.

        \item XXX MORE\@?

    \end{itemize}

\end{frame}

\section{Future Directions}

\begin{frame}{Future Directions}

    \timingnote{1--2 minutes}

    \begin{itemize}

        \item Use Machine Learning instead of writing rules; Instance Based
            Learning techniques a requirement due to the size of the data
            set.

        \item Detection of overlapping rules.

        \item XXX REFERENCE IMPLEMENTATION\@?

    \end{itemize}

\end{frame}

\section*{Summary}

\begin{frame}{Summary}

    \timingnote{1 minute}

    \begin{itemize}

        \item East to develop a flexible yet strict parser.

        \item Easy for users to parse new or changed inputs.

        \item Parser can still be efficient and fast.

    \end{itemize}

    \timingnote{remainder}

    \begin{itemize}

        \item XXX MAYBE ON A SEPARATE SLIDE\@?

        \item Questions?

        \item Comments?

        \item Suggestions?

    \end{itemize}

\end{frame}

\end{document}