### 2009-03-14 doc/logparser.tex
Dealt with some XXX items.

% $Id$
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

% Aligned list environments.
\usepackage{eqlist}
% This voodoo below gives us an eqlist environment with bold items.
\newcommand{\bolditem}[1]{%
    \bfseries#1%
}
\newenvironment{boldeqlist}
    {\begin{eqlist}[\renewcommand{\makelabel}{\bolditem}]}
    {\end{eqlist}}

% Enable customised lists.
\usepackage{enumitem}
% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

% Extra packages recommended by Springer.
%\usepackage{mathptmx}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{makeidx}
%\usepackage{multicol}

% Embed SVN Id etc.
\usepackage{svn}

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Warnings rather than errors when references cross page boundaries.
\vrefwarning{}
% Errors when loops are encountered
\vrefshowerrors{}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

% Sort numbers where there are multiple citations.  Does not appear to have
% the expected effect (probably clashes with varioref or hyperref), though
% it does reduce the space between numbers.
\usepackage{cite}
% Check for unused references, and how the citation key (e.g.\ slct-paper)
% in the margin beside the reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used; neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the defaul spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[3]{
    \begin{table}[ht]
        \caption{#2}\label{#3}
        \input{#1}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \hline%
    \noalign{\smallskip}%
}

\newcommand{\tablebottomline}[0]{%
    \noalign{\smallskip}%
    \hline%
}

\newcommand{\tablemiddleline}[0]{%
    \noalign{\smallskip}%
    \hline%
    \noalign{\smallskip}%
}

% A command to format a Postfix daemon's name
\newcommand{\daemon}[1]{%
    \texttt{postfix/#1}%
}

\newcommand{\flowchart}[2]{%
    (action: #1\@; flowchart:~#2)%
}

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}

\newcommand{\tab}[0]{%
    \hspace*{1em}%
}

% Constant values.
\newcommand{\numberOFlogFILES}[0]{%
    93%
}

\newcommand{\numberOFlogFILESall}[0]{%
    522%
}

\newcommand{\numberOFrules}[0]{%
    185%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    116%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.70\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    59.48\%%
}

\newcommand{\numberOFlogLINES}[0]{%
    60,721,709%
}

\newcommand{\numberOFlogLINEShuman}[0]{%
    60.72 million%
}

\newcommand{\numberOFactions}[0]{%
    21%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    17,020%
}

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\acronym{PLP}}
\newcommand{\parsernames}{\acronyms{PLP}}
\newcommand{\parsernamelong}{Postfix Log Parser}

\begin{document}

\title{Parsing Postfix log files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

\begin{abstract}

    \setcounter{page}{2}
    \phantomsection{}
    \addcontentsline{toc}{chapter}{Abstract}

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding the log files, and thus accuracy in reconstructing the
    actions taken by Postfix to generate the log files.  This paper
    describes the creation of a parser for Postfix log files, documenting
    the architecture developed and the parser implementing it, explaining
    the difficulties encountered and the solutions developed, with
    reference to an implementation that stores data gleaned from the log
    files in an SQL database.  The gathered data can then be used to
    optimise current anti-spam measures, provide a baseline to test new
    anti-spam measures against, or to produce statistics showing how
    effective those measures are.

    \SVN$Id$
    \begin{center}
        \SVNId{}
    \end{center}

    XXX CHECK ALL ``UNINTERESTING'' RULES TO SEE IF THEY COULD BE USED\@.

    XXX ADD A $\backslash{}$action\{\} COMMAND TO FORMAT ACTION NAMES
    CONSISTENTLY\@.

\end{abstract}

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\setcounter{page}{3}
\tableofcontents
\listoffigures
\listoftables

\newpage

% WSUIPA fonts.
\input{ipamacs}

\input{logparser-introduction}
\input{logparser-background}
\input{logparser-other-parsers-review}
\input{logparser-architecture}
\input{logparser-implementation}
\input{logparser-results}
\input{logparser-conclusion}

\appendix

\bibliographystyle{logparser-bibliography-style}
\bibliography{logparser-bibliography}

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

\end{document}