### 2009-03-26 doc/logparser-complications.tex
More corrections from proofreading.

\section{Complications Encountered While Writing \parsernamelong{}}

\label{complications}

It was initially expected that parsing Postfix log files would be a
relatively simple task, requiring a couple of months of work.  The author
had found Postfix log files useful when investigating problems reported by
users, and an examination of the log files gave the impression that they
would be straightforward to parse, process, and understand.  The large
variation in log lines was not apparent, because the majority of log lines
are recognised by a small set of rules, as shown in \figureref{rule hits
graph}.  Most of the myriad complications and difficulties documented in
this section were discovered during \parsernames{} development, but the
first three complications were identified during the planning and design
phase of this project, influencing the architecture's design.

Each of these complications caused \parsername{} to operate incorrectly,
generate warning messages, or leave mails in the state table.  The
complications are listed in the order in which they were overcome during
development of \parsername{}, with the first complication occurring several
orders of magnitude more frequently than the last.  When deciding which
problem to tackle next, the problem causing the highest number of warning
messages or mails incorrectly remaining in the state tables was always
chosen, because that approach yielded the biggest improvement in the
parser, and made the remaining problems more apparent.

\subsection{Queueid Vs Pid}

\label{queueid vs pid}

A delivery attempt lacks a queueid until one recipient has been accepted,
so log lines must first be correlated by \daemon{smtpd} \acronym{pid}, then
transition to being correlated by their queueid.  This is relatively minor,
but does require:

\begin{itemize}

    \item Two versions of several functions: \texttt{by\_pid} and
        \texttt{by\_queueid}.

    \item Two state tables to hold data structures.

    \item Most importantly: every section of code must know which state
        table it should query.

\end{itemize}

\subsection{Connection Reuse}

\label{connection reuse}

Multiple independent mails may be delivered across one connection: this
requires \parsername{} to clone the connection's data as soon as a mail is
accepted, so that subsequent mails will not overwrite each other's data.
This must be done every time a mail is accepted, as it is impossible to
tell in advance which connections will accept multiple mails.  Once a mail
has been accepted its log lines will not be correlated by \acronym{pid} any
more, its queueid will be used instead, as described in \sectionref{queueid
vs pid}.  If the original connection has any useful data (e.g.\ rejections)
it will be saved to the database when the client disconnects.  One unsolved
difficulty is distinguishing between different groups of rejections, e.g.\
when dealing with the following sequence:

\begin{enumerate}

    \item The client attempts to deliver a mail, but it is rejected.

    \item The client issues the RSET command to reset the \acronym{SMTP}
        session.

    \item The client attempts to deliver another mail, likewise rejected.

\end{enumerate}

There should ideally be two separate entries in the database resulting from
the above sequence, but currently there will only be one.

\subsection{Re-injected Mails}

\label{Re-injected mails}

\label{tracking re-injected mail}

Mails sent to local addresses are not always delivered directly to a
mailbox: sometimes they are sent to and accepted for a local address, but
need to be delivered to one or more remote addresses due to aliasing.  When
this occurs, a child mail will be injected into the Postfix queue, but
without the explicit logging that mails injected by \daemon{smtpd} or
\daemon{postdrop} have.  Thus the source of the mail is not immediately
discernible from the log line in which the mail's queueid first appears:
from a strictly chronological reading of the log lines it usually appears
as if the child mail has been created without a source.  Subsequently the
parent mail will log the creation of the child mail, e.g.\ parent mail
\texttt{3FF7C4317} creates child mail \texttt{56F5B43FD}\@:

\texttt{3FF7C4317: to=<username@example.com>, relay=local, \hfill{}
\newline{} \tab{} \tab{} delay=0, status=sent (forwarded as 56F5B43FD)}

Unfortunately, while all log lines from an individual process appear in
chronological order, the order in which log lines from different processes
are interleaved is subject to the vagaries of process scheduling.  In
addition, the first log line belonging to the child mail (the example log
line above belongs to the parent mail) is logged by either \daemon{qmgr} or
\daemon{cleanup}, so the order also depends on how soon they process the
new mail.

Because of the uncertain order that log lines can appear in, \parsername{}
cannot complain when it encounters a log line from either \daemon{qmgr} or
\daemon{cleanup} for a previously unseen mail; instead it must flag the
mail as coming from an unknown origin, and subsequently clear that flag if
and when the origin of the mail becomes clear.  Obviously the parser could
omit checking where mails originate from, but requiring an explicit source
helps to expose bugs in the parser; such checks helped to identify the
complications described in \sectionref{discarding cleanup log lines} and
\sectionref{pickup logging after cleanup}.

Process scheduling can have a still more confusing effect: quite often the
child mail will be created, delivered, and entirely finished with,
\textit{before\/} the parent mail logs its creation!  Thus, mails flagged
as coming from an unknown origin cannot be entered into the database when
their final log line is processed, nor can \parsername{} warn the user;
instead they must be marked as ready for entry and subsequently entered
once their origin has been identified.  The crux of this complication is
that re-injected mails appear in the log files without explicit logging
indicating their source.

Tracking re-injected mail requires \parsername{} to do the following in the
COMMIT action:

\begin{enumerate}

    \item If a mail is tagged with the unknown origin flag, it is assumed
        to be a child mail whose parent has not yet been identified.  The
        mail is tagged as ready to be entered in the database, but entry is
        postponed until the parent is identified.  The child mail should
        not have any subsequent log lines: only its parent will refer to
        it.

    \item If the mail is a child mail whose parent has been identified, it
        is entered in the database as usual, then removed from its parent's
        list of children.  If this child is the last mail on that list, and
        the parent has already been entered in the database, the parent
        will be removed from the state tables.

    \item If the mail is a parent, it is entered in the database as usual
        because there will be no further log lines for it.  There may be
        child mails waiting to be entered in the database; these are
        entered as normal, and removed from the state tables.  If there are
        incomplete child mails, the parent's removal from the state tables
        will be postponed until the last child has been entered.

\end{enumerate}

\subsection{Identifying Bounce Notifications}

\label{identifying bounce notifications}

Postfix 2.2.x (and presumably previous versions) does not generate a log
line when it generates a bounce notification; suddenly there will be log
entries for a mail that lacks an obvious source.  There are similarities to
the problem of identifying re-injected mails discussed in
\sectionref{tracking re-injected mail}, but unlike the solution described
therein bounce notifications do not eventually have a log line that
identifies their source.  Heuristics must be used to identify bounce
notifications:

\begin{enumerate}

    \item The sender address is \verb!<>!.\glsadd{<>}

    \item Neither \daemon{smtpd} nor \daemon{pickup} have logged any
        messages associated with the mail, indicating it was generated
        internally by Postfix, not accepted via \acronym{SMTP} or submitted
        locally by \daemon{postdrop}.

    \item The message-id has a specific format: \newline{}
        \tab{} \texttt{YYYYMMDDhhmmss.queueid@server\_hostname} \newline{}
        e.g.\ \texttt{20070321125732.D168138A1@smtp.example.com}

    \item The queueid in the message-id must be the same as the queueid of
        the mail: this is what distinguishes a new bounce notification from
        a bounce notification that is being re-injected as a result of
        aliasing.  In the latter case the message-id will be unchanged from
        the original bounce notification, and so even if it happens to be
        in the correct format, e.g.\ if it was generated by Postfix on this
        or another server, it will not match the queueid of the mail.

\end{enumerate}

Once a mail has been identified as a bounce notification, the unknown
origin flag is cleared and the mail can be entered in the database.

There is a small chance that a mail will be incorrectly identified as a
bounce notification, as the heuristics used may be too broad.  For this to
occur the following conditions would have to be met:

\begin{enumerate}

    \item The mail must have been generated internally by Postfix.

    \item The sender address must be \verb!<>!.\glsadd{<>}

    \item The message-id must have the correct format and match the queueid
        of the mail.  While a mail sent from elsewhere could easily have
        the correct message-id format, the chance that the queueid in the
        message-id would match the queueid of the mail is extremely small.

\end{enumerate}

If a mail is misclassified as a bounce message it will almost certainly
have been generated internally by Postfix; arguably misclassification in
this case is a benefit rather than a drawback, as other mails generated
internally by Postfix will be handled correctly.  Postfix 2.3 (and
hopefully subsequent versions) log the creation of a bounce message.

This check is performed during the COMMIT action.

\subsection{Aborted Delivery Attempts}

\label{aborted delivery attempts}

Some mail clients behave strangely during the \acronym{SMTP} dialogue: the
client aborts the first delivery attempt after the first recipient is
accepted, then makes a second delivery attempt for the same recipient that
it continues with until delivery is complete.  Microsoft Outlook is one
client that behaves in this fashion; other clients may act in a similar
way.  An example dialogue exhibiting this behaviour is presented below
(lines starting with a three digit number are sent by the server, the other
lines are sent by the client):

\begin{verbatim}
220 smtp.example.com ESMTP
EHLO client.example.com
250-smtp.example.com
250-PIPELINING
250-SIZE 15240000
250-ENHANCEDSTATUSCODES
250-8BITMIME
250 DSN
MAIL FROM: <sender@example.com>
250 2.1.0 Ok
RCPT TO: <recipient@example.net>
250 2.1.5 Ok
RSET
250 2.0.0 Ok
RSET
250 2.0.0 Ok
MAIL FROM: <sender@example.com>
250 2.1.0 Ok
RCPT TO: <recipient@example.net>
250 2.1.5 Ok
DATA
354 End data with <CR><LF>.<CR><LF>
The mail transfer is not shown.
250 2.0.0 Ok: queued as 880223FA69
QUIT
221 2.0.0 Bye
\end{verbatim}

Once again Postfix does not log a message making the client's behaviour
clear, so once again heuristics are required to identify when this
behaviour occurs.  A list of all mails accepted during a connection is
saved in the connection's state table entry, and the accepted mails are
examined when the disconnection action is invoked.  Each mail is checked
for the following:

\begin{itemize}

    \item Was the second result processed by the \texttt{CLONE} action?
        The first two \daemon{smtpd} log lines will be a connection log
        line and a mail acceptance log line.

    \item Is \daemon{smtpd} the only Postfix component that produced a log
        line for this mail?  Every mail which passes normally through
        Postfix will have a \daemon{cleanup} line, and later a
        \daemon{qmgr} log line; lack of a \daemon{cleanup} line is a sure
        sign the mail did not make it too far.

    \item Can the mail be found in the state tables?  If not it cannot be
        an aborted delivery attempt.

    \item If there are third and subsequent results, were those log lines
        processed by the \texttt{SAVE\_DATA} action?  If there are any log
        lines after the first two they should be informational only.

\end{itemize}

If all the checks above are successful then the mail is assumed to be an
aborted delivery attempt and is discarded.  There will be no further
entries logged for such mails, so without identifying and discarding them
they accumulate in the state table and will cause clashes if the queueid is
reused.  The mail cannot be entered in the database as the only data
available is the client hostname and \acronym{IP} address, but the database
schema requires many more fields be populated (see \sectionref{connections
table} and \sectionref{results table}).  These heuristics are quite
restrictive, and it appears there is little scope for false positives; if
there are any false positives there will also be warnings when the next log
line for that mail is parsed.  False negatives are less likely to be
detected: there may be queueid clashes (and warnings) if mails remain in
the state tables after they should have been removed, otherwise the only
way to detect false negatives is to examine the state tables after each
parsing run.

This check is performed in the DISCONNECT action; it requires support in
the CLONE action where the list of accepted mails is maintained.

\subsection{Further Aborted Delivery Attempts}

Some mail clients disconnect abruptly if a second or subsequent recipient
is rejected; they may also disconnect after other errors, but such
disconnections are either unimportant or are handled elsewhere in the
parser, e.g.\ \sectionref{timeouts during data phase}.  Sadly, Postfix does
not log a message saying the mail has been discarded, as should be expected
by now.  The checks to identify this happening are:

\begin{itemize}

    \item Is the mail missing its \daemon{cleanup} log line?  Every mail
        which passes normally through Postfix will have a \daemon{cleanup}
        line; lack of one is a sure sign the mail did not make it too far.

    \item Were there three or more \daemon{smtpd} log lines for the mail?
        There should be a connection log line and a mail acceptance log
        line, followed by one or more delivery rejected log lines.

\end{itemize}

If both checks are successful then the mail is assumed to have been
discarded when the client disconnected.  There will be no further log lines
for such mails, so without identifying and entering them in the database
immediately they accumulate in the state table and will cause clashes if
the queueid is reused.

These checks are made during the DISCONNECT action.

\subsection{Timeouts During DATA Phase}

\label{timeouts during data phase}

The DATA phase of the \acronym{SMTP} conversation is where the headers and
body of the mail are transferred.  Sometimes there is a timeout or the
connection is lost during the DATA phase;\footnote{For the sake of brevity
\textit{timeout\/} will be used throughout this section, but everything
applies equally to lost connections.} when this occurs Postfix will discard
the mail and \parsername{} needs to discard the data associated with that
mail.  It seems more intuitive to save the mail's data to the database, but
if a timeout occurs there is no data available to save; the timeout is
recorded and saved with the connection instead.  To deal properly with
timeouts the TIMEOUT action does the following:

\begin{enumerate}

    \item Record the timeout and captured data in the connection's results.

    \item If no mails have been accepted yet nothing needs to be done; the
        TIMEOUT action ends.

    \item If one or more recipients have been accepted, Postfix will have
        allocated a queueid for the incoming mail, and there may be a mail
        in the state tables that needs to be discarded.  The timeout may
        have interrupted transfer of an accepted delivery attempt, or it
        may have occurred after a mail delivery attempt was rejected.

        XXX THE REMAINDER OF THIS DOES NOT MAKE SENSE\@; CHECK AND CORRECT
        IT\@.  To distinguish between the two cases the action compares the
        timestamp of the last accepted mail against the timestamp of the
        last log line from \daemon{smtpd} for that connection; this depends
        on the CLONE action keeping a list of all mails accepted for each
        connection.  If the mail acceptance timestamp is later than the
        timeout timestamp, the timeout applies to the just-accepted mail,
        which will be removed from the state tables, without entering it in
        the database.  If the \daemon{smtpd} timestamp is later there was a
        rejection between the accepted mail and the timeout: the action
        assumes that the timeout applies to the rejected delivery attempt
        and finishes.  This assumption is not necessarily correct, because
        Postfix may have accepted an earlier recipient and rejected a later
        one, in which case the timeout applies to the accepted mail, which
        should be discarded; this has not been a problem in practice.
        Processing timeouts is further complicated by the presence of out
        of order \daemon{cleanup} log lines: see \sectionref{discarding
        cleanup log lines} for details.

\end{enumerate}

This complication is dealt with by the TIMEOUT action, with help from the
CLONE action.

\subsection{Discarding Cleanup Log Lines}

\label{discarding cleanup log lines}

The author has only observed this complication occurring after a timeout,
though there may be other circumstances that trigger it.  Sometimes the
\daemon{cleanup} log line for a mail being accepted is logged after the
timeout log line, by which time \parsername{} has discarded the mail;
parsing the \daemon{cleanup} log line causes the CLEANUP\_PROCESSING action
to create a new state table entry, to help deal with re-injected mails
(\sectionref{Re-injected mails}).  This is incorrect because the log line
actually belongs to the mail that has just been discarded; if the queueid
is reused there will be a queueid clash, otherwise the new mail will just
remain in the state tables.

During the TIMEOUT action, if the \daemon{cleanup} line is still pending,
the action updates the \texttt{timeout\_queueids} state table, adding the
queueid and the timestamp from the log line.  When the next
\daemon{cleanup} line is parsed for that queueid, the list will be checked
and if the log line meets the following requirements it will be deemed part
of the mail where the timeout occurred and discarded.

\begin{itemize}

    \item The queueid must not have been reused yet, i.e.\ it does not have
        a mail in the state tables.

    \item The timestamp of the \daemon{cleanup} log line must be within ten
        minutes of the mail acceptance timestamp.  Timeouts happen after
        five minutes, but some data may have been transferred slowly, and
        empirical evidence shows that ten minutes is not unreasonable;
        hopefully it is a good compromise between false positives (log
        lines incorrectly discarded) and false negatives (new state table
        entries incorrectly created).

\end{itemize}

The \daemon{cleanup} log line must meet the criteria above for it to be
discarded because some, but not all, connections where a timeout occurs
will have an associated \daemon{cleanup} log line; if the
CLEANUP\_PROCESSING action blindly discarded the next \daemon{cleanup} log
line after a timeout it would sometimes be mistaken.  When the next
\daemon{pickup} log line containing that queueid is processed the queueid
will be removed from the cache of timeout queueids, regardless of whether
it meets the criteria above.

This complication is handled by the CLEANUP\_PROCESSING and TIMEOUT
actions.

\subsection{Pickup Logging After Cleanup}

\label{pickup logging after cleanup}

When mail is submitted locally, \daemon{pickup} accepts the new mail and
generates a log line showing the source.  Occasionally this log line will
occur later in the log file than the \daemon{cleanup} log line, so the
PICKUP action will find that a state table entry exists for that queueid.
Normally if the queueid given in the PICKUP line exists in the state tables
a warning is generated by the \daemon{pickup} action, but if the following
conditions are met it is assumed that the log lines were out of order:

\begin{itemize}

    \item The only program which has logged anything thus far for the mail
        is \daemon{cleanup}.

    \item There is less than a five second difference between the
        timestamps of the \daemon{cleanup} and \daemon{pickup} log lines.

\end{itemize}

As always with heuristics there may be circumstances in which these
heuristics match incorrectly, but none have been identified so far.  This
complication seems to occur during periods of particularly heavy load, so
is most likely caused by process scheduling vagaries.

This complication is dealt with during the PICKUP action.

\subsection{Smtpd Stops Logging}

\label{smtpd stops logging}

Occasionally a \daemon{smtpd} will just stop logging, without an
immediately obvious reason.  After poring over log files for some time
several reasons have been found for this infrequent occurrence:

\begin{enumerate}

    \item Postfix is stopped or its configuration is reloaded.  When this
        happens all \daemon{smtpd} processes exit, so all entries in the
        connections state table must be cleaned up, entered in the database
        if there is sufficient data, and deleted.

    \item Sometimes a \daemon{smtpd} is killed by a signal (sent by Postfix
        for some reason, by the administrator, or by the OS), so its active
        connection must be cleaned up, entered in the database if there is
        sufficient data, and deleted from the connections state table.

    \item Occasionally a \daemon{smtpd} will exit uncleanly, so the active
        connection must be cleaned up, entered in the database if there is
        sufficient data, and deleted from the connections state table.

    \item Every Postfix process uses a watchdog which kills the process if
        it is not reset for a considerable period of time (five hours by
        default).  This safeguard prevents Postfix processes from running
        indefinitely and consuming resources if a failure causes them to
        enter a stuck state.

\end{enumerate}

The circumstances above account for all occasions identified thus far where
a \daemon{smtpd} suddenly stops logging.  In addition to removing an active
connection the last accepted mail may need to be discarded, as detailed in
\sectionref{timeouts during data phase}; otherwise the queueid state table
is untouched.

These occurrences are handled by the three actions POSTFIX\_RELOAD,
SMTPD\_DIED, and SMTPD\_WATCHDOG\@.

\newpage{} %XXX VARIOREF PROBLEM

\subsection{Out of Order Log Lines}

\label{out of order log lines}

Occasionally a log file will have out of order log lines which cannot be
dealt with by the techniques described in \sectionref{tracking re-injected
mail}, \sectionref{discarding cleanup log lines} or \sectionref{pickup
logging after cleanup}.  In the \numberOFlogFILES{} log files used for
testing this occurs only five times in 60,721,709 log lines, but for parser
correctness it must be dealt with.  The five occurrences have the same
characteristics: the \daemon{local} log line showing delivery to a local
mailbox occurs after the \daemon{qmgr} log line showing removal of the mail
from the queue because delivery is completed.  This causes problems: the
data in the state tables for the mail is not complete, so entering it into
the database fails; a new mail is created when the \daemon{local} line is
parsed and remains in the state tables; four warnings are issued per pair
of out of order log lines.

The COMMIT action examines the list of programs that have produced a log
lines for each mail, comparing the list against a table of known-good
program combinations.  If the mail's combination is found in the table the
mail can be entered in the database; if the combination is not found entry
must be postponed and the mail flagged for later entry.  The SAVE\_DATA
action checks for that flag; if the additional log lines have caused the
mail to reach a valid combination then entry in the database will proceed,
otherwise it must be postponed once more.

The list of valid combinations is explained below.  Every mail will
additionally have log line from \daemon{cleanup} and \daemon{qmgr}; any
mail may also have log line from \daemon{bounce}, \daemon{postsuper}, or
both.

% This will put the text on the line following the item name, if the
% enumitem package is loaded.
%\begin{description}[style=nextline]
\begin{description}

    \item [\daemon{local}:] Local delivery of a bounce notification, or
        local delivery of a re-injected mail.

    \item [\daemon{local}, \daemon{pickup}:] Mail submitted locally on the
        server, delivered locally on the server.

    \item [\daemon{local}, \daemon{pickup}, \daemon{smtp}:] Mail submitted
        locally \newline{} on the server, for both local and remote
        delivery.

    \item [\daemon{local}, \daemon{smtp}, \daemon{smtpd}:] Mail accepted
        from a remote client, for both local and remote delivery.

    \item [\daemon{local}, \daemon{smtpd}:] Mail accepted from a remote
        client, for local delivery only.

    \item [\daemon{pickup}, \daemon{smtp}:] Mail submitted locally on the
        server, for remote delivery only.

    \item [\daemon{smtp}:] Remote delivery of either a re-injected mail or
        a bounce notification.

    \item [\daemon{smtp}, \daemon{smtpd}:] Mail accepted from a remote
        client, then remotely delivered (typically relaying mail for
        clients on the local network to addresses outside the local
        network).

    \item [\daemon{smtpd}, \daemon{postsuper}:] Mail accepted from a remote
        client, then deleted by the administrator before any delivery
        attempt was made (the unwanted mail is typically due to a mail loop
        or joe~job\glsadd{Joe-job}).  Notice that \daemon{postsuper} is
        required, not optional, for this combination.

\end{description}

This check applies to accepted mails only, not to rejected mails.  This
check is performed during the COMMIT action.

\subsection{Yet More Aborted Delivery Attempts}

\label{yet more aborted delivery attempts}

The aborted delivery attempts described in \sectionref{aborted delivery
attempts} occur frequently, but the aborted delivery attempts described in
this section only occur four times in the \numberOFlogFILES{} log files
used for testing.  The symptoms are the same as in \sectionref{aborted
delivery attempts}, except that there \textit{is\/} a \daemon{cleanup} log
line; there is nothing in the log file to explain why there are no further
log lines.  The only way to detect these mails is to periodically scan all
mails in the state tables, deleting any mails displaying the following
characteristics:

\begin{itemize}

    \item The timestamp of the last log line for the mail must be 12 hours
        or more earlier than the last log line parsed from the current log
        file.

    \item There must be exactly two \daemon{smtpd} and one \daemon{cleanup}
        log lines for the mail, with no additional log lines.

\end{itemize}

12 hours is a somewhat arbitrary time period, but it is far longer than
Postfix would delay delivery of a mail in the queue (unless Postfix is not
running for an extended period of time).  The state tables are scanned for
mails matching the characteristics above each time the end of a log file is
reached, and matching mails are deleted.

\subsection{Mail Deleted Before Delivery is Attempted}

\label{Mail deleted before delivery is attempted}

Postfix logs the recipient address when delivery of a mail is attempted, so
if delivery has yet to be attempted the parser cannot determine the
recipient address or addresses.  This is a problem when mail is arriving
faster than Postfix can attempt delivery, and the administrator deletes
some of the mail (because it is the result of a mail loop\glsadd{mail
loop}, mail bomb\glsadd{mail bomb}, or joe~job\glsadd{Joe-job}) before
Postfix has had a chance to try to deliver it.  In this case the recipient
address will not have been logged, so a dummy recipient address needs to be
added, as every mail is required by the database schema
(\sectionref{results table}) to have at least one recipient.  Typically
when this complication occurs there are many instances of it, closely
grouped.

This lack of information cannot easily be overcome: it is trivial to log a
warning for every recipient accepted, but when the warning for the first
recipient is logged Postfix will not yet have allocated a queue file and
queueid for the mail, so the warning will be associated with the connection
rather than the accepted mail.  A queue file and queueid will be allocated
after Postfix accepts the MAIL FROM command if
\texttt{smtpd\_delay\_open\_until\_valid\_rcpt} is set to ``no'', but that
setting will cause disk IO for almost every delivery attempt, instead of
just for delivery attempts where recipients are accepted, and consequently
a drastic reduction in the performance of the mail server.

The DELETE action is responsible for handling this complication.

\subsection{Bounce Notification Mails Are Delivered Before Their Creation
Is Logged}

\label{Bounce notification mails delivered before their creation is logged}

This is yet another complication that only occurs during periods of
extremely high load, when process scheduling and even hard disk access
times cause strange behaviour.  In this complication bounce notification
mails are created, delivered, and deleted from the queue, \textit{before\/}
the log line from \daemon{bounce} that explains their source is logged.  To
deal with this the COMMIT action maintains a cache of recently committed
bounce notification mails, which the BOUNCE action subsequently checks if
the bounce mail is not already in the state tables.  If the queueid exists
in the cache, and its start time is less than ten seconds before the
timestamp of the bounce log line, it is assumed that the bounce
notification mail has already been processed and the BOUNCE action does not
create one.  If the queueid exists it is removed from the cache, because it
has either just been used or it is too old to use in future.  Whether the
BOUNCE action creates a mail or finds an existing mail in the state tables,
it flags the mail as having been seen by the BOUNCE action; if this flag is
present the COMMIT action will not add the mail to the cache of recent
bounce notification mails.  This is not required to correctly deal
with the complication, but is an optimisation to reduce the parser's memory
usage; on the occasions the author has observed this complication occurring
there have been a huge number of bounce notification mails generated --- if
every bounce notification mail was cached it would dramatically increase
the memory requirements of the parser.  The cache of bounce notification
mails will be pruned whenever the parser's state is saved, though if the
size of the cache ever becomes a problem it could be pruned periodically to
keep the size in check.

\subsection{Mails Deleted During Delivery}

\label{Mails deleted during delivery}

The administrator can delete mails using \daemon{postsuper}; occasionally
mails that are in the process of being delivered will be deleted.  This
results in the log lines from the delivery agent (\daemon{local},
\daemon{virtual} or \daemon{smtp}) appearing in the log file
\textit{after\/} the mail has been removed from the state tables and saved
in the database.  The DELETE action adds deleted mails to a cache, which is
checked by the SAVE\_DATA action, and the current log line discarded if the
following conditions are met:

\begin{enumerate}

    \item The queueid is not found in the state tables.

    \item The queueid is found in the cache of deleted mails.

    \item The timestamp of the log line is within 5 minutes of the final
        timestamp of the mail.

\end{enumerate}

Sadly this solution involves discarding some data, but the complication
only arises eight times in quick succession in one log file, which is not
in the \numberOFlogFILES{} log files used for testing; if this complication
occurred more frequently it might be desirable to find the mail in the
database and add the log line to it.