### 2009-03-26 doc/logparser-implementation.tex

\chapter{Postfix Parser Implementation}

\label{Postfix Parser Implementation}

XXX WRITE AN INTRODUCTION\@.  MOST OF THIS IS CRAP, BUT I MIGHT BE ABLE TO
RESCUE SOME OF IT\@.  IT SHOULD HAVE A BLURB, FOLLOWED BY A PREVIEW OF THE
CHAPTER\@.

The parser deals with all the complications of parsing, the eccentricities
and oddities of Postfix log files, presenting the resulting data in a
normalised, simple to use representation.  The parser's task is to follow
the journey each mail takes through Postfix, combining the data captured by
rules into a coherent whole, saving it in a useful and consistent form, and
performing housekeeping duties.

The intermingling of log entries from different mails immediately rules out
the possibility of handling each mail in isolation; the parser must be able
to handle multiple mails in parallel, each potentially at a different stage
in its journey, without any interference between mails --- except in the
minority of cases where intra-mail interference is required.  The best way
to implement this is to maintain state information for every unfinished
mail and manipulate the appropriate mail correctly for each log line
encountered.

There is a similarity between this design and the event-driven programming
paradigm commonly used in GUI programs, where one part of the program
responds to events (mouse clicks in a GUI program, log lines being
recognised in \parsername{}) and invokes the correct action.

\section{Assumptions}

\parsername{} makes a small number of (hopefully safe and reasonable)
assumptions:

\begin{itemize}

    \item The log files are whole and complete: nothing has been removed,
        either deliberately or accidentally (e.g.\ log file rotation gone
        awry, file system filling up, logging system unable to cope with
        the volume of log messages).  On a well run system it is extremely
        unlikely that any of these problems will arise, though it is of
        course possible, particularly when suffering from a deluge of spam
        or a mail loop.

    \item Postfix logs enough information to make it possible to accurately
        reconstruct the actions it has taken.  Heuristics are used in
        several places when parsing; see \sectionref{identifying bounce
        notifications}, \sectionref{aborted delivery attempts}, and
        \sectionref{pickup logging after cleanup} for details.  At least
        one difficulty encountered while writing \parsername{}
        (\sectionref{yet more aborted delivery attempts}) could not be
        solved using the data in the log files, and requires a brute force
        approach.

    \item The Postfix queue has not been tampered with, causing unexplained
        appearance or disappearance of mail.  This may happen if the
        administrator deletes mail from the queue without using
        \daemon{postsuper}, or if there is filesystem corruption.

\end{itemize}

In some ways this task is similar to reverse engineering or replicating a
black box system based solely on its inputs and outputs.  Thus far
analysis of the log files has been sufficient to reconstruct Postfix's
behaviour, but for more difficult programs the techniques described in
\cite{black-box-error-reporting} may be useful.  There are some advantages
to treating Postfix as a black box during parser development:

\begin{itemize}

    \item Reading and understanding the source code would require a
        significant investment of time: Postfix 2.5.5 has 17MB of source
        code.

    \item \parsername{} is developed using real world log files rather than
        the idealised log files someone would naturally envisage when
        reading the source code, which cannot accurately communicate the
        variety of orderings in which log lines are found in log files.

    \item The parser acts as a second source of information about Postfix's
        operation, based on empirical evidence.  A separate project could
        compare the empirical knowledge inherent in \parsername{} with
        Postfix's documentation and source code to see how closely the two
        agree.

\end{itemize}

\section{Parser Flow Chart}

\label{flow chart}

A picture is said to be worth a thousand words: the flow chart in
\figureref{flow chart image} shows the most common paths a connection or
mail can take through \parsername{}; the difficulties described in
\sectionref{complications} are excluded for the sake of clarity.  The flow
chart is intended to be a graphical overview of how a mail progresses
through both Postfix and \parsername{}, providing an overall context into
which the detailed descriptions in the remainder of this chapter will fit,
in particular the actions (\sectionref{actions in implementation}) and
complications (\sectionref{complications}).

\showgraph{build/logparser-flow-chart-part-1}{Parser flow chart}{flow chart
image}

Everything starts off with a mail entering the system, whether by local
submission via \daemon{postdrop}, by \acronym{SMTP}, by re-injection due to
forwarding, or internally generated by Postfix.  Local submission is the
simplest case: a queueid is assigned immediately \flowchart{PICKUP}{2}, and
the mail moves on to the delivery stage.  Re-injection due to forwarding
sadly lacks explicit log lines of its own; it is explained fully in
\sectionref{Re-injected mails}.  Internally generated mails lack any
explicit origin in Postfix 2.2.x and must be detected using heuristics as
described in \sectionref{identifying bounce notifications}; later versions
of Postfix do provide log lines for internally generated mails.  Bounce
notifications are the primary example of internally generated mails, though
there are other types, e.g.\ Postfix may generate mails to the
administrator when it encounters configuration errors, but such mails are
presumably rare.  XXX WOULD THIS PARAGRAPH BE BETTER AS A LIST\@?

\acronym{SMTP} is more complicated than the others:

\begin{enumerate}

    \item First there is a connection from the remote client
        \flowchart{CONNECT}{1}.

    \item This is followed by rejection of sender address, recipient
        addresses, client \acronym{IP} address or hostname, HELO hostname,
        etc.\ \flowchart{DELIVERY\_REJECTED}{4}; acceptance of one or more
        mails \flowchart{CLONE}{5}; or some interleaving of both.

    \item The client disconnects \flowchart{DISCONNECT}{6}.  If Postfix has
        rejected any \acronym{SMTP} commands the resulting data will be
        saved to the database; if there were no rejections there will not
        be any data to save.  Any mails accepted will already have been
        cloned so their data is in another data structure, nd will be
        delivered in the same way as mails that entered the system by any
        other route.

\end{enumerate}

The obvious counterpart to mail entering the system is mail leaving the
system, whether by deletion, bouncing, local delivery, or remote delivery.
All four are handled in exactly the same way:

\begin{enumerate}

    \item Postfix will log the sender and recipient addresses separately
        \flowchart{SAVE\_DATA}{8}.

    \item Sometimes mail is re-injected and the child mail needs to be
        tracked by the parent mail \flowchart{TRACK}{9}; the handling of
        re-injected mails is described in \sectionref{tracking re-injected
        mail}.

    \item Eventually the mail will be delivered, bounced, or deleted by the
        administrator \flowchart{COMMIT}{11}.  This is the last log line
        for this particular mail, though it may be indirectly referred to
        if it was re-injected.  The mail is cleaned up and entered in the
        database, then deleted from the state tables.

\end{enumerate}

It should be emphasised that the mail delivery process above happens
whether the mail is delivered to a mailbox, piped to a command, delivered
to a remote server, bounced (due to a mail loop, delivery failure, or five
day timeout), or deleted by the administrator, \textit{unless\/} the mail
is either parent or child of re-injection, as explained in
\sectionref{tracking re-injected mail}.

\section{Database}

\label{database}

A database is used to store both the rules and the data gleaned by parsing
Postfix log files.  Understanding the database schema is important in
understanding the actions of the parser, and essential to developing
further applications which utilise the data gathered; \sectionref{database
as API} describes how the database schema functions as an \acronym{API}.

The database schema can be conceptually divided in two: the rules used to
recognise log lines, and the data saved from the parsing of log files.
Each rule has a regex to recognise log lines and capture data from them,
and specifies the action to be invoked when a log line is recognised; they
also have several fields that aid the user in understanding the meaning of
the log lines recognised by each rule.  The rules are described in detail
in \sectionref{rules in implementation}, but the rules table is documented
in \sectionref{rules table} with the rest of the database schema.

The data saved from parsing the log files is also divided into two tables:
connections and results.  The connections table contains an entry for every
mail accepted and every connection where there was a rejection; the
individual fields will be described in \sectionref{connections table}.  The
results table will have at least one entry for each entry in the
connections table; its fields will be covered in detail in
\sectionref{results table}.  A diagram of the database schema is provided
in \figureref{Diagram of the Database Schema picture}, to complement the
in-depth descriptions of each table.

An important but easily overlooked benefit of storing the rules in the
database is the link between rules and results: if more information is
required when examining a result, the rule that produced the result is
available for inspection because each result references the rule that
created it.  There is no ambiguity about which rule resulted in a
particular result, eliminating one potential source of confusion.

A clear, comprehensible schema is essential when using the data extracted
from log lines; it is more important when using the data than when storing
it, because storing the data is a write-once operation, whereas utilising
the data requires frequent searching, sorting, and manipulation of the data
to produce customised reports or statistics.

\subsection{Using A Database To Provide An Application Programming Interface}

\label{database as API}

The database populated by \parsername{} provides a simple interface to
Postfix log files.  Although the interface is a database schema rather than
a set of functions in a library, it gives the same benefits as any other
\acronym{API}: it provides a stable interface between the user and the
creator of the data, allowing code on either side of the interface to be
changed without adverse effects, as long as the interface is adhered to.
Programs that use the database can range from the simple examples in
\sectionref{motivation} to far more complex data mining tools.

Using a database simplifies writing programs that need to interact with the
data in several ways:

\begin{enumerate}

    \item Most programming languages have facilities for database access,
        allowing a developer to write programs that use the gathered data
        in their preferred programming language, rather than being
        restricted to the language the parser is written in.

    \item Databases provide complex querying and sorting functionality for
        the user without requiring large amounts of programming.  All
        databases provide a program, of varying complexity and
        sophistication, that can be used for ad hoc queries with minimal
        investment of time.

    \item Databases are easily extensible, e.g.:

        \begin{itemize}

            \item New columns can be added to the tables used by the
                program, using DEFAULT clauses or TRIGGERS to populate
                them.

            \item A VIEW gives a custom arrangement of data with minimal
                effort.

            \item Some databases support granting access on a fine-grained
                basis, e.g.\ allowing the finance department to produce
                invoices, the helpdesk to run limited queries as part of
                dealing with support calls, and the administrators to have
                full access to the data.

            \item Triggers can be written to perform actions when certain
                events occur.  In pseudo-\acronym{SQL}\@:

\begin{verbatim}
CREATE TRIGGER ON INSERT INTO results
    WHERE sender = "boss@example.com"
        AND action = "DELIVERY_REJECTED"
    SEND PANIC EMAIL TO "postmaster@example.com";
\end{verbatim}

            \item Other tables can be added to the database, e.g.\ to cache
                historical or computed data, or to link in data from other
                sources.

        \end{itemize}

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn, e.g.\
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
        Although every vendor implements a different dialect of
        \acronym{SQL}, the basics are the same everywhere, analogous to the
        overall similarity and minor differences amongst Irish English,
        British English, American English, and Australian English.
        Depending on the database in use there may be tools available that
        reduce or remove the requirement to know \acronym{SQL}.

\end{enumerate}

Storing the results in a database will also increase the efficiency of
using those results, because the log files need only be parsed once rather
than each time the data is used; indeed the results may be used by someone
with no access to the original log files.

\subsection{Rules Table}

\label{rules table}

\newlength{\belowcaptionskipORIG}
\setlength{\belowcaptionskipORIG}{\belowcaptionskip}
\setlength{\belowcaptionskip}{10pt}
\showgraph{database-schema}{Diagram of the Database Schema}{Diagram
of the Database Schema picture}
\setlength{\belowcaptionskip}{\belowcaptionskipORIG}

Rules are discussed in detail in \sectionref{rules in implementation}, but
the structure of the rules table is documented here along with the other
tables in the database.  Rules are created by the user, not the parser, and
will not be modified by \parsername{} (except for the hits and hits\_total
fields).  Rules classify the individual log lines, capturing data to be
saved in the connections and results tables, and specifying the action to
take for each recognised log line.

Each rule some or all of these fields; all fields are required unless
otherwise stated.

\begin{boldeqlist}

    \item [id] A unique identifier for each rule that other tables can use
        when referring to a specific rule.

    \item [name] A short name for the rule.

    \item [description] Something must have occurred to cause Postfix to
        log each line, this field should describe the event causing the log
        lines this rule recognises.

    \item [restriction\_name] The name of the Postfix restriction that
        caused the mail to be rejected.  Only applicable to rules that
        recognise rejection log lines, i.e.\ rules that have an action of
        \texttt{DELIVERY\_REJECTED}; other rules should not set this field.

    \item [program] The Postfix component (e.g.\ \daemon{smtpd}) whose log
        lines the rule applies to; see \sectionref{rule conditions in
        implementation} for full details of how this attribute is used.

    \item [regex] The regex to recognise log lines with, as documented in
        \sectionref{regex components}.

    \item [connection\_data] Sometimes rules need to provide data that is
        not present in the log line: e.g.\ setting \texttt{client\_ip} when
        a mail is being delivered to another server.  Any field in the
        connections table can be set in this way.  The format is:
        \newline{} \tab{} \texttt{ client\_hostname = localhost,}
        \newline{} \tab{} \tab{} \texttt{client\_ip = 127.0.0.1} \newline{}
        i.e.\ semi-colon or comma separated assignment statements.  Commas
        and semi-colons cannot be escaped and thus cannot be included in
        data, this feature is intended for use with small amounts of data,
        so dealing with escape sequences was deemed unnecessary.  This
        field is optional.

    \item [result\_data] The result table equivalent of
        \texttt{connection\_data}, also optional.

    \item [action] The action that will be invoked when this rule
        recognises a log line; a full list of actions and the parameters
        they are invoked with can be found in \sectionref{actions in detail
        in implementation}.

    \item [hits] This counter is maintained for every rule and incremented
        each time the rule successfully recognises a log line.  At the
        start of each run \parsername{} sorts the rules by hits, and at the end of the run it updates
        every rule's hits field in the database.  Assuming that the
        distribution of log lines is reasonably consistent across log
        files, ordering rules by their recognition frequency will reduce
        the parser's execution time.  Rule ordering for efficiency is
        discussed in \sectionref{rule ordering for efficiency}.  This field
        will be set by the parser rather than the rule author.

    \item [hits\_total] The total number of hits for this rule over all
        runs of the parser; hits is reset to zero each time the parser is
        run, but hits\_total is not.  This field will be set by the parser
        rather than the rule author.

    \item [priority] This is the user-configurable companion to hits: when
        the list of rules is sorted, priority overrides hits.  This allows
        more specific rules to take precedence over more general rules, as
        described in \sectionref{rules in architecture}.  This field is
        optional.

\end{boldeqlist}

\subsection{Connections Table}

\label{connections table}

Every accepted mail and every connection where there was a rejection will
have a single entry in the connections table containing all of the
following fields:

\begin{boldeqlist}

    \item [id] This field uniquely identifies the row.

    \item [server\_ip] The \acronym{IP} address of the server: the local
        address when receiving mail, the remote address when sending mail.

    \item [server\_hostname] The hostname of the server, it will be
        \texttt{unknown} if the \acronym{IP} address could not be resolved
        to a hostname via DNS\@.

    \item [client\_ip] The client \acronym{IP} address: the remote address
        when receiving mail, the local address when sending mail.

    \item [client\_hostname] The hostname of the client, it will be
        \texttt{unknown} if the \acronym{IP} address could not be resolved
        to a hostname via DNS\@.

    \item [helo] The hostname used in the HELO command.  The HELO hostname
        occasionally changes during a connection, presumably because spam
        or virus senders think it is a good idea.  By default Postfix only
        logs the HELO hostname when it rejects an \acronym{SMTP} command,
        but it is quite simple to rectify this as described in
        \sectionref{logging helo}.

    \item [queueid] The queueid of the mail if the connection represents an
        accepted mail, or \texttt{NOQUEUE} if not.

    \item [start] The timestamp of the first log line.

    \item [end] The timestamp of the last log line.

\end{boldeqlist}

\subsection{Results Table}

\label{results table}

Every recognised log line will have a row in the results table, associated
with a single connection; a single connection will have at least one result
associated with it, but will usually have several more, and may have
hundreds.

\begin{boldeqlist}

    \item [connection\_id] The id of the row in the connections table this
        result is associated with.

    \item [rule\_id] The id of the rule in the rules table that recognised
        the log line and created this result.

    \item [id] A unique identifier for this result.

    \item [warning] Postfix can be configured to log a warning instead of
        enforcing a restriction that would reject an \acronym{SMTP} command
        --- a mechanism that is quite useful for testing new restrictions.
        This field will be false for a real rejection, or true if the log
        line was a warning.  This field should be ignored if the result is
        not a rejection, i.e.\ the action field of the associated rule is
        not \texttt{DELIVERY\_REJECTED}.

    \item [smtp\_code] The \acronym{SMTP} code associated with the log
        line.  In general an \acronym{SMTP} code is only present in a log
        line representing a rejection or final delivery; results whose log
        line did not contain an \acronym{SMTP} code will duplicate the
        \acronym{SMTP} code of other results in that connection.  Some
        final delivery log lines do not contain an \acronym{SMTP} code: in
        those cases the \acronym{SMTP} code is specified by the rule's
        \texttt{result\_data} field, based on the success or failure
        represented by the log line.  XXX GIVE AN EXAMPLE LOG LINE\@.

    \item [enhanced\_status\_code] The enhanced status code~\cite{RFC3463}
        is similar to the \acronym{SMTP} code, but is intended to be
        interpreted by mail clients so that error messages can be clearly
        conveyed to the user.  Enhanced status code support was added to
        Postfix in version 2.3; log files from previous versions will not
        contain any enhanced status codes.

    \item [sender] The sender's email address.  This may change during a
        connection if the client uses different sender addresses for
        multiple rejected delivery attempts; however there should not be
        different sender addresses in the results for one accepted mail.
        Mails sent over the same connection can be distinguished by their
        queueids.

    \item [recipient] The recipient address; there may be multiple
        recipient addresses per mail or connection.

    \item [size] The size of the mail, present for delivered mails only.

    \item [delay] How long the mail was delayed while it was being
        delivered.  This will only be present for delivered mails.

    \item [delays] More detailed information about how long the mail was
        delayed while it was being delivered, again only present for
        delivered mails.

    \item [message\_id] The message-id of the accepted mail, again present
        for delivered mails only.

    \item [data] A field available to store a piece of captured data that
        does not have a specific field, e.g.\ the rejection message from a
        \acronym{DNSBL}\@.

    \item [timestamp] The time at which the log line was logged.

\end{boldeqlist}

\section{Framework}

The role of the framework in this architecture was described in detail in
\sectionref{framework in architecture}; this section is concerned with the
implementation of the framework within \parsername{}.  The framework
manages the parsing process, taking care of the drudgery and boring tasks,
provides services to the actions, and implements some efficiency measures.

Each time the parser is run the framework performs some initialisation
tasks, setting up data structures that will later be used by actions,
including several state tables.  Data about the connections and mails being
processed is held in \texttt{connections} and \texttt{queueids}
respectively.  The other state tables used when working around
complications in parsing: \texttt{timeout\_queueids} is used when dealing
with connections that time out during the DATA phase (\sectionref{timeouts
during data phase}); \texttt{bounce\_queueids} is part of the solution to
bounce notification mails being delivered before their creation is logged
(\sectionref{Bounce notification mails delivered before their creation is
logged}); and \texttt{postsuper\_deleted\_queueids} caches information
about mails recently deleted by the administrator so that subsequent log
lines processed by the SAVE\_DATA action can be discarded (\sectionref{Mail
deleted before delivery is attempted}).

When the framework loads the ruleset, it verifies each of the rules,
checking:

\begin{itemize}

    \squeezeitems{}

    \item That the specified action is registered with the framework.

    \item The regex is valid and compiling it for efficiency
        (\sectionref{Caching compiled regexes}).

    \item For overlap between the data captured by the regex and additional
        data specified in either \texttt{connection\_data} or
        \texttt{result\_data}.

    \item That \texttt{connection\_data}, \texttt{result\_data}, and the
        regex captures specify valid fields to save data to.

\end{itemize}

If there are state tables from a previous \parsername{} run to be loaded,
they will be loaded now; the framework supports saving state at any time,
without adversely affecting the parsing process.  The need to track
re-injected mail (\sectionref{Re-injected mails}) complicates the process
of saving state, as the relationships between mails must be maintained;
loading state is also complicated by dealing with aborted delivery attempts
(\sectionref{aborted delivery attempts}), because a separate set of
relationships between connections and mails must be re-created when the
state tables are restored.

The framework is now ready to begin parsing.  For each log line it will use
those rules whose \texttt{program} field equals the program in the log line
(see \sectionref{rule conditions in implementation}), falling back to
generic rules if necessary, and warning if the log line is unrecognised.
The rules are sorted for increased efficiency; see \sectionref{rule
ordering for efficiency} for details.  The repetitive nature of log files
gives them high compression ratios; the framework supports reading
compressed log files, simplifying the process of parsing them because they
do not have to be uncompressed to a temporary file before parsing begins.
The framework also displays a progress bar to show how far parsing has
progressed through the log file, and to give the user an indication of how
long the remainder of the parsing process is likely to take.  The progress
bar is not as reliable when parsing compressed log files, because a
compressed block may uncompress to a variable number of log lines; there is
also variation between the recognition and processing time of individual
log lines, but overall the progress bar is a useful addition.

The framework collects data used when evaluating \parsernames{} efficiency
for the results chapter (\sectionref{Results}); some of the techniques used
to improve parsing speed can be turned off or altered to measure the effect
they have.  Three sets of data are collected:

\begin{enumerate}

    \item How long it takes to parse each log file.

    \item The number of rules tried when recognising log lines; the
        framework may need to try multiple rules for a given log line
        before it finds one that recognises it.

    \item How many log lines did the framework attempt to recognise, how
        many were skipped because there were no rules whose condition was
        equal to that log line's program, how many were successfully
        recognised, and how many were not recognised.

\end{enumerate}

There are five ways the framework can adapt its behaviour to demonstrate
how effective the optimisations it uses are, and how efficient
\parsername{} is:

\begin{enumerate}

    \item The rule ordering used can be changed from optimal (the default,
        most efficient) to shuffled (intended to represent an unordered
        ruleset) or reverse (reverse of optimal, least efficient).  The
        results of this are shown in \sectionref{rule ordering for
        efficiency}.

    \item The framework can record which rule recognised each log line, and
        then on a subsequent run consult that information so that it uses
        the correct rule for each log line.  This gives the best possible
        running time, because only one rule is used for each log line.  A
        variation of this is to use the correct rule last, to get the worst
        possible running time.  \sectionref{perfect rule ordering}
        discusses this in detail.

    \item Each regex is compiled and the result cached when the rules are
        loaded; this can be disabled and the regex compiled every time it
        is used when recognising log lines.  The effect of this
        optimisation is shown in \sectionref{Caching compiled regexes}.

    \item Normally when a log line has been recognised the framework
        invokes the action specified by the rule.  Invoking the action can
        be skipped if desired, so the timing information shows how long is
        spent recognising log lines only; this time can be subtracted from
        the time taken by a normal run to show how long is spent processing
        log lines.  This data is analysed in \sectionref{recognising vs
        processing}.

    \item The framework can skip inserting data into the database, because
        that dominates the execution time of the parser; the evaluation
        chapter measures the speed of \parsername{}, not the speed of the
        database or the disks it resides on.

\end{enumerate}

The framework also provides several debugging options, to aid in writing or
correcting rules, or figuring out why the parser is not behaving as
expected.  In increasing order of severity they are:

\begin{enumerate}

    \item The regex from the recognising rule and the log line it
        recognised can be printed, so that the user can verify that the
        correct rule recognised each log line.  A possible variant of this
        would be to mark specific rules so only their regexes and the log
        lines they recognise would be printed.

    \item Each result can be extended with extra debugging information,
        which is useful when a warning dumps a mail's data for inspection.
        The information currently added is: the log line's timestamp in
        human readable form; the entire log line; the name of the log file
        and the number of the log line within it.  This information is not
        stored in the database.

    \item Every connection and result added to the database can be dumped
        in a human readable form.  This will result in a huge amount of
        debugging information, so it is only useful for small log file
        snippets, because otherwise the amount of information is
        overwhelming.

\end{enumerate}

\section{Actions}

\label{actions in implementation}

Actions are the component of this architecture responsible for processing
all of the inputs recognised by rules; in \parsername{} they reconstruct
the journey each mail takes through Postfix, dealing with all the
complications and difficulties that arise.  \parsername{} has
\numberOFactions{} actions and \numberOFrules{} rules, with an uneven
distribution of rules to actions as shown in \graphref{Distribution of
rules per action}.  Unsurprisingly, the action with the most associated
rules is \texttt{DELIVERY\_REJECTED}, the action that processes Postfix
rejecting a mail delivery attempt.  The next most common action is, perhaps
surprisingly, \texttt{UNINTERESTING}: this action does nothing when
invoked, allowing uninteresting log lines to be parsed without any effects
(it does not imply that the input is ungrammatical or unparsed).  Generally
rules specifying the \texttt{UNINTERESTING} action recognise log lines that
are not associated with a specific mail, e.g.\ notices about configuration
files changing; these log lines are recognised and processed so that the
framework can warn about unrecognised log lines, informing the user that
they need to add to or extend their ruleset.  Most of the remaining actions
have only one or two associated rules: some are specialised actions
required to address a deficiency in the log files, or a complication that
arises during parsing; other actions will only ever have one log line
variant, e.g.\ all log lines showing that a remote client has connected are
recognised by a single rule and handled by the \texttt{CONNECT} action.
\parsernames{} actions are described in detail in \sectionref{actions in
detail in implementation}, and \sectionref{adding new actions in
implementation} covers the process of adding a new action.

\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action}

\subsection{Actions in the \parsernamelong{}}

\label{actions in detail in implementation}

This section explains the actions found in \parsername{}; it may help to
revisit the flow chart in \sectionref{flow chart} to see how a mail passes
from one action to another as its log lines are recognised.  The words
\textit{mail\/} and \textit{connection\/} are used in the actions
descriptions below because they are less unwieldy and more specific than
\textit{state table entry\/}; a connection becomes a mail during the
\texttt{CLONE} action, which processes Postfix accepting a delivery
attempt, and the data structure moves from the connections state table to
the queueids state table.

The complications and difficulties that arose when parsing real-world log
files are documented in \sectionref{complications}; some action
descriptions refer to specific difficulties they address.  The
complications are documented in a separate section to avoid overwhelming
the action descriptions.

If there is enough information in the log line to identify the correct
connection or mail, each action will save all the data captured by the
recognising rule's regex.  Each action is passed the same arguments:

\begin{boldeqlist}

    \squeezeitems{}

    \item [rule] The recognising rule.

    \item [data] The data captured from the log line by the rule's regex.

    \item [line] The log line, separated into fields:

        \begin{boldeqlist}

            \squeezeitems{}

            \item [timestamp] The time the line was logged at.

            \item [program] The name of the program that generated the log
                line.

            \item [pid] The \acronym{pid} of the process that logged the
                line.

            \item [host] The hostname of the server the line was logged on.

            \item [text] The remainder of the log line, i.e.\ the message
                logged by the program and recognised by the rule.

        \end{boldeqlist}

\end{boldeqlist}

\begin{description}

    \item [BOUNCE\_CREATED] Postfix 2.3 and subsequent versions log the
        creation of bounce messages.  This action creates a new mail if
        necessary; if the mail already exists the unknown origin flag will
        be removed from it.  The action also marks the mail as a bounce
        notification.  To deal with complication \sectionref{Bounce
        notification mails delivered before their creation is logged} this
        action checks a cache of recent bounce mails, to avoid creating
        bogus bounce mails when log lines are out of order.

    \item [CLEANUP\_PROCESSING] \daemon{cleanup} processes every mail that
        passes through Postfix; details of what it does are available in
        \sectionref{Postfix Daemons}.  This action saves all data captured
        by the rule's regex if the log line has not come after a timeout
        (see \sectionref{discarding cleanup log lines}); it also creates
        the mail if necessary, setting its unknown origin flag (see
        \sectionref{pickup logging after cleanup}).

    \item [CLONE] Multiple mails may be accepted on a single connection, so
        each time a mail is accepted the connection's state table entry
        must be cloned and saved in the state tables under its queueid; if
        the original data structure was used then second and subsequent
        mails would overwrite one another's data.

    \item [COMMIT] Enter the data from the mail into the database.  Entry
        will be postponed if the mail is a child waiting to be tracked
        (\sectionref{Re-injected mails}).  Once entered in the database,
        the mail will be usually be deleted from the state tables, but
        deletion will be postponed if the mail is the parent of re-injected
        mail (\sectionref{Re-injected mails}).

    \item [CONNECT] Handle a remote client connecting: create a new
        connection, indexed by \daemon{smtpd} \acronym{pid}.  If a
        connection already exists it is treated as a symptom of a bug in
        \parsername{}, and the action will issue a warning containing the
        full contents of the existing connection plus the log line that has
        just been parsed.

    \item [DELETE] Deals with mail deleted using Postfix's administrative
        command \daemon{postsuper}.  This action adds a dummy recipient
        address if required (see \sectionref{Mail deleted before delivery
        is attempted}), then invokes the COMMIT action to save the mail to
        the database.

    \item [DELIVERY\_REJECTED] Deals with Postfix rejecting an
        \acronym{SMTP} command from the remote client.  This is the action
        most frequently specified by rules, because so many different
        restrictions are used to reject delivery attempts.  This action is
        quite simple: if the log line contains a queueid, save the data
        captured by the rule's regex to the mail identified by that
        queueid; otherwise save it to the connection identified by the
        \acronym{pid} in the log line.  No further processing of the log
        line is required.

    \item [DISCONNECT] Invoked when the remote client disconnects, it
        enters the connection in the database if it has any useful data,
        performs any required cleanup, and deletes the connection from the
        state tables.  This action deals with aborted delivery attempts
        (\sectionref{aborted delivery attempts}).

    \item [EXPIRY] If Postfix has not managed to deliver a mail after
        trying for five days, it will give up and return the mail to the
        sender.  When this happens the mail will not have a combination of
        Postfix programs which passes the valid combinations check,
        implemented to deal with the complication described in
        \sectionref{out of order log lines}; this action tags the mail as
        having expired, so the COMMIT action will skip the valid
        combinations check.

    \item [MAIL\_BOUNCED] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish delivery attempts that bounce from other
        delivery attempts.

    \item [MAIL\_DISCARDED] Sometimes mail is discarded, e.g.\ mail
        submitted locally that is larger than the limit configured by the
        administrator.  This action is used for those mails; it invokes the
        \texttt{COMMIT} action, but is a separate action to simplify
        further analysis.

    \item [MAIL\_QUEUED] This action represents Postfix picking a mail from
        the queue to deliver.  This action needs to deal with out of order
        log lines when mail is re-injected for forwarding; see
        \sectionref{Re-injected mails} for details.

    \item [MAIL\_SENT] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish successful delivery attempts from other
        delivery attempts.

    \item [MAIL\_TOO\_LARGE] When a client tries to send a mail larger than
        the local server accepts, the mail will be discarded and the client
        informed of the problem.  The mail may have been accepted and
        partially transferred, in which case there's a data structure to
        dispose of.  See \sectionref{timeouts during data phase} for full
        details; although that describes timeouts the processing is the
        same for mails that are too large.

    \item [PICKUP] The PICKUP action corresponds to the \daemon{pickup}
        service dealing with a locally submitted mail.  A new mail will be
        created, although out of order log entries may have caused it to
        already exist, as documented in \sectionref{pickup logging after
        cleanup}.

    \item [POSTFIX\_RELOAD] When Postfix stops or reloads its
        configuration, it kills all \daemon{smtpd} processes, requiring all
        connections to be cleaned up, entered in the database, and deleted
        from the state tables.  Postfix probably kills all the other
        components too, but \parsername{} is only affected by
        \daemon{smtpd} processes exiting.

    \item [SAVE\_DATA] Every action that can locate the correct entry in
        the state tables saves any data captured by the recognising rule's
        regex to it.  The \texttt{SAVE\_DATA} saves data in this way but
        does not do anything else; it is invoked for log lines that contain
        useful data but do not require any further processing.

    \item [SMTPD\_DIED] Sometimes a \daemon{smtpd} dies, is killed by a
        signal, or exits unsuccessfully; the associated connection must be
        cleaned up, entered in the database if it has sufficient data, and
        deleted from the state tables.  In some cases the connection will
        not have enough data to satisfy the database schema, so it cannot
        be entered into the database; unfortunately this means that some
        data is lost.

    \item [SMTPD\_WATCHDOG] \daemon{smtpd} processes have a watchdog timer
        to deal with unusual situations; after five hours the timer will
        expire and the \daemon{smtpd} will exit.  This occurs very
        infrequently, as there are many other timeouts that should occur in
        the intervening hours, e.g.\ timeouts for DNS requests or timeouts
        reading data from the client.  The active connection for that
        \daemon{smtpd} must be cleaned up, entered in the database, and
        deleted from the state tables.

    \item [TIMEOUT] The connection with the remote client timed out, so the
        mail currently being transferred must be discarded.  The mail may
        have been accepted, in which case there's a data structure to
        dispose of.  See \sectionref{timeouts during data phase} for full
        details.

    \item [TRACK] Track a mail when it is re-injected for forwarding to
        another mail server; this happens when a local address is aliased
        to a remote address (see \sectionref{tracking re-injected mail} for
        a full explanation).  \texttt{TRACK} will be called when dealing
        with the parent mail, and will create the child mail if necessary.
        \texttt{TRACK} checks if the child has already been tracked, either
        by this parent or by another parent, and issues appropriate
        warnings in either case.

    \item [UNINTERESTING] This action just returns successfully: it is used
        when a log line needs to be parsed for completeness, but does not
        either provide any useful data to be saved or require anything to
        be done.

\end{description}

\subsection{Adding New Actions}

\label{adding new actions in implementation}

Adding new actions is not as easy as adding new rules, though care has been
taken in the architecture and implementation to make adding new actions as
painless as possible.  The implementer writes a subroutine that accepts the
standard arguments given to actions, and registers it with the framework.
The new action must be registered before the rules are loaded, because it
is an error for a rule to specify an unregistered action; this helps catch
mistakes made when adding new rules.  No other work is required from the
implementer to integrate the action into the parser; all of their attention
and effort can be focused on correctly implementing their new action.  The
action may need to extend the list of valid combinations described in
\sectionref{out of order log lines} if the new action creates a different
set of acceptable programs, but this would only be necessary if the new
action processes log lines from Postfix components \parsername{} currently
does not have rules for, e.g.\ \daemon{virtual} or \daemon{lmtp}.

\section{Rules}

\label{rules in implementation}

The rules are responsible for classifying each log line and specifying the
correct action to be invoked.  In any parser implemented using this
architecture, the rules will be the most visible component and most likely
to be modified by users.  Rules need to be as simple as possible so that
users can easily modify or add to them, but the implementation must balance
that simplicity with the need to provide enough flexibility and power to
successfully parse inputs.

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that description will not be repeated here, but
should be fleshed out by the example rule in \sectionref{example rule in
implementation}.  The process of creating new rules from unparsed log lines
is dealt with in \sectionref{creating new rules in implementation}; the
implementation of the utility supplied with \parsername{} to create regexes
from unparsed log lines is also described in that section.  The rule
conditions used in \parsername{} are explained next, and the final portion
of this section provides some suggestions for how to detect overlapping
rules.

\subsection{Example Rule}

\label{example rule in implementation}

The example rule in \tableref{Example rule in implementation table}
recognises the log line resulting from Postfix rejecting a delivery attempt
because the domain in the sender address does not have an A, AAAA, or MX
DNS entry, i.e.\ mail could not be delivered to any address in the sender's
domain (for full details see~\cite{reject-unknown-sender-domain}).

Example log line recognised by this rule:

% RFC 3330 says that 192.0.2.0/24 is reserved for example use.

\begin{verbatim}
NOQUEUE: reject: RCPT from smtp.example.com[192.0.2.1]:
  550 5.1.8 <foo@example.com>:
  Sender address rejected: Domain not found;
  from=<foo@example.com> to=<info@example.net>
  proto=SMTP helo=<smtp.example.com>
\end{verbatim}

% do not reformat this!
\begin{table}[ht]

    \caption{Example rule}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}

    \textbf{Field}      & \textbf{Value}                                    \\
    name                & Unknown sender domain                             \\
    description         & We do not accept mail from unknown domains        \\
    restriction\_name   & reject\_unknown\_sender\_domain                   \\
    program             & \daemon{smtpd}                                    \\
    regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                        & \verb!Sender address rejected: Domain not found;! \\
                        & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                        & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
    result\_data        &                                                   \\
    connection\_data    &                                                   \\
    action              & DELIVERY\_REJECTED                                \\
    hits                & 0                                                 \\
    hits\_total         & 0                                                 \\
    priority            & 0                                                 \\
    %cluster\_group      & 400                                               \\

    \end{tabular}

\end{table}

The fields in \tableref{Example rule in implementation table} are used as
follows:

\begin{description}

    \item [name, description, and restriction\_name:] are not used by the
        parser, they serve to document the rule for the user's benefit.

    \item [program and regex:] The program is used to restrict the log
        lines this rule will be used to recognise; see \sectionref{rule
        conditions in implementation} for details.  The regex does the
        actual recognition of the log lines, and data captured by the regex
        (sender, recipient, etc.) will be automatically saved to the
        results and connections tables.

    \item [action:] The action to be invoked when the rule recognises a log
        line.  See \sectionref{actions in detail in implementation} for
        details of the actions implemented by \parsername{}, and
        \sectionref{actions in architecture} for the role of actions in the
        parser architecture.

    \item [result\_data and connection\_data:] are used to provide data not
        present in the log line, and are unused in this rule.

    \item [hits, hits\_total, and priority:] hits and priority are used
        when ordering the rules for more efficient parsing (see
        \sectionref{rule ordering for efficiency}).  At the end of each
        parsing run hits is set to the number of log lines recognised by
        the rule.  Hits\_total is the sum of hits over every parsing run,
        but is otherwise unused by the parser.

\end{description}

\subsection{Creating New Rules}

\label{creating new rules in implementation}

The log files produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different
\acronym{DNSBL} services, policy servers, or custom rejection messages.  To
facilitate parsing new log lines, the architecture separates rules from
actions: adding new actions can be difficult, but adding new rules to
recognise new log lines is trivial, and occurs much more frequently.

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and regex.  The name and description fields should be
set based on the meaning of the log line, to help others understand which
log lines this rule will recognise; the program will be obvious from the
unrecognised log lines.  The action depends on the what the log line
represents, e.g.\ a delivery rejection, a mail being delivered, some useful
information, or something else; examine the list of actions in
\sectionref{actions in detail in implementation} to determine the correct
one.  The regex needs to be constructed based on the log line, but see
below for a tool to ease the process.

There are other attributes that may be required to make a rule work
correctly: connection\_data, result\_data, priority, or restriction\_name.
In general it will only become clear that connection\_data or result\_data
are required when \parsername{} warns about an entry in the connections or
results tables that is missing some required fields, because values for the
missing fields are not present in any of the log lines.  E.g.\ the rule
recognising the \daemon{pickup} component processing a mail sets the
client\_hostname to \texttt{localhost} and the client\_ip to
\texttt{127.0.0.1}, because the mail originates on the local server.  If
the new rule deliberately uses the architecture's overlapping rules feature
it will be obvious that the priority field needs to be set, on this rule
and possibly others; the priority field may be needed on unintentionally
overlapping rules too, but that is more difficult to determine.  Finally
the restriction\_name field should be set when the rule's action is
\texttt{DELIVERY\_REJECTED}; hopefully the name of the restriction will be
clear from the content of the log line.

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: programs generally create log lines by substituting variable words
into a fixed pattern, and analysis of the frequency with which each word
occurs can be used to determine whether the word is variable or part of the
fixed pattern.  This classification can be used to group similar log lines
and generate a regex to match each group of log lines.  There are 4 steps
in the algorithm:

\begin{description}

    \item [Pre-process the file.]  The new algorithm leverages the
        knowledge gained when writing rules and developing \parsername{};
        it performs a large number of substitutions on the input log lines,
        replacing commonly occurring variable terms (e.g.\ email addresses,
        \acronym{IP} addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces a large number of variable words
        with fixed words, improving the subsequent classification of words
        as fixed or variable.  The altered log lines are written to a
        temporary file, which subsequent stages will use instead of the
        original input file.

        In the original algorithm the purpose of the pre-processing stage
        was to reduce the memory consumption of the program.  In the first
        pass it generates a hash~\cite{hash-functions}\glsadd{hash} (from a
        small range of values) for each word of each log line, incrementing
        a counter for each hash.  The counters will be used in the next
        pass to filter out words: if the word's hash does not have a high
        frequency, the word itself cannot have a high frequency, so there
        is no need to maintain a counter for it, reducing the number of
        counters required and thus the program's memory consumption.

    \item [Calculate word frequencies.]  The position of words within a log
        line is important: a word occurring in two log lines does not
        indicate similarity unless it occupies the same position in both
        log lines.  If a variable term substituted into a log line contains
        spaces, it will appear to the algorithm as more than one word.
        This will alter the position of subsequent words, so a word
        occurring in different positions in two log lines \textit{may\/}
        indicate similarity, but the algorithm does not attempt to deal
        with this possibility.  The algorithm maintains a counter for each
        \texttt{(word, word's position within the log line)} tuple,
        incrementing it each time that word occurs in that position.

        The original algorithm hashes each word and checks that result's
        counter from the previous pass; if that counter has a high
        frequency a separate counter will be maintained for this
        \texttt{(word, word's position within the log line)} tuple.  This
        reduces the number of counters maintained by the algorithm in this
        step, reducing the memory requirements of the algorithm at the cost
        of increased CPU usage.

        As time goes by the amount of memory typically available to a
        program or algorithm increases, and the need to reduce memory
        requirements correspondingly decreases, so the modified algorithm
        omits the hashing step and maintains counters for all tokens.  Most
        of the infrequently occurring words will have been substituted with
        keywords during the first step, vastly reducing the number of
        tuples to maintain counters for; the original algorithm does not
        have the detailed knowledge leveraged by the new algorithm, because
        it is a generic tool.

    \item [Classify words based on their frequency.]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that will hopefully be
        noticed by the user when reviewing the new regexes, and will be
        obvious when the new rules do not recognise some log lines they are
        expected to.  Variable terms are replaced by \texttt{.+} to match
        one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes are combined, and their counters added together.  If a
        regex's counter is lower than the threshold supplied by the user
        the regex is discarded; this second threshold is independent of the
        threshold used to differentiate between fixed and variable words,
        but once again 1\% of log lines is a good starting point.  The new
        regexes are displayed for the user to add to the database, either
        as new rules or merged into the regexes of existing rules; the
        counter for each regex is also displayed, giving the user an
        indication of how many of the log lines that regex should match.
        Discarding regexes will result in some of the log lines not being
        matched; when the rules have been augmented with the new regexes,
        the original log files should be parsed again, and any remaining
        unparsed log lines used as input to this utility.

\end{description}

XXX MERGE logs2regexes WITH check-regexes.  CHECK IF SLCT DOES THIS STEP\@.

A second utility is also provided that reads a list of new regexes and the
input given to the first utility.  It tries to match each input log line
with each regex, counting the number of log lines that match each regex,
warning the user if an input log line is matched by more than one regex,
and additionally warning if an input log line is not matched by any regex.
It displays a summary of how many input log lines each regex matched,
comparing it to the expected number of matches; this provides the user with
an easy method of checking if the regexes produced by the first utility are
correctly matching the input log lines they are based upon.

These utilities are not intended to create perfect regexes, but they
greatly reduce the effort required to parse new or different log lines.

\subsection{Rule Conditions}

\label{rule conditions in implementation}

Rule conditions as part of the architecture have already been documented in
\sectionref{rule conditions in architecture}, and they can be very complex
and difficult to evaluate.  In contrast, \parsername{} uses very simple
rule conditions: each rule has a \texttt{program} attribute that specifies
the Postfix component whose log lines it recognises, and each log line
contains the name of the Postfix component that produced it; when trying to
recognise log lines, the framework will only use rules where the two are
equal.  This avoids needlessly trying rules that will not recognise the log
line, or worse, might recognise it unintentionally.  In addition the
framework supports generic rules, whose program attribute is ``*''; these
will be used if none of the program-specific rules recognise the log line.
If none of the rules are successful the framework will warn the user.

\subsection{Regex Components}

\label{regex components}

Each rule's regex will have keywords expanded when the rules are loaded,
for several reasons:

\begin{itemize}

    \item It simplifies both reading and writing of regexes and makes each
        regex largely self-documenting.  E.g.\ the meaning of
        \_\_CLIENT\_HOSTNAME\_\_ is immediately clear, whereas
        \verb!(?:unknown|(?:[-.\w]+))! needs to be deciphered each time it
        is encountered.

    \item It avoids needless repetition of complex regex components, and
        allows the components to be corrected or improved in one location.
        E.g.\ \_\_SENDER\_\_ is used in 68 rules; if a mistake is
        discovered it only needs to be corrected in one place.

    \item It enables automatic extraction and saving of captured data.  The
        regex snippets use Perl 5.10's named capture buffers~\cite{perlre}
        to capture data, so the mapping between captures and fields does
        not need to be explicitly specified by the rule.

\end{itemize}

For efficiency the keywords are expanded and every rule's regex is compiled
before attempting to parse the log file, otherwise every regex would be
recompiled each time it was used, resulting in a large, data dependent
slowdown, as shown in \sectionref{Caching compiled regexes}.  Most of the
keywords are named after the fields in the connections or results tables
they populate: \_\_CLIENT\_HOSTNAME\_\_, \_\_CLIENT\_IP\_\_, \_\_DELAY\_\_,
\_\_DELAYS\_\_, \_\_ENHANCED\_STATUS\_CODE\_\_, \_\_HELO\_\_,
\_\_MESSAGE\_ID\_\_, \_\_QUEUEID\_\_, \_\_RECIPIENT\_\_, \_\_SENDER\_\_,
\_\_SERVER\_HOSTNAME\_\_, \_\_SERVER\_IP\_\_, \_\_SIZE\_\_, and
\_\_SMTP\_CODE\_\_.

Some of the keywords need more explanation:

\begin{eqlist}

    \squeezeitems{}

    \item [\_\_CHILD\_\_]  The queueid of a child mail; see
        \sectionref{Re-injected mails}.

    \item [\_\_COMMAND\_\_]  All \acronym{SMTP} commands.

    \item [\_\_CONN\_USE\_\_]  How many times the connection was reused;
        Postfix tries to reuse connections whenever possible to reduce the
        load on both the sending and receiving servers.

    \item [\_\_DATA\_\_]  This snippet is special: it matches a zero-length
        string, so it needs to be followed by a pattern that matches
        appropriately, but it captures whatever is matched so that it can
        be saved.  E.g.\ \verb!/(__DATA__connection (?:refused|reset!
        \newline{} \tab{}\verb!by peer))/! matches either ``connection
        refused'' or ``connection reset by peer'' and causes it to be saved
        to the data field of that log line's result.

    \item [\_\_PID\_\_]  The \acronym{pid} of a \daemon{smtpd} process that
        dies or is killed; see the SMTPD\_DIED action in
        \sectionref{actions in detail in implementation}.

    \item [\_\_RESTRICTION\_START\_\_]  Matches the standard information
        Postfix includes at the start of almost all log lines resulting
        from rejecting a delivery attempt.

    \item [\_\_SHORT\_CMD\_\_]  Postfix sometimes logs \acronym{SMTP}
        commands in a short, single word form; this snippet matches all of
        those, except \texttt{DATA}, which typically has a more specific
        rule.  Priorities could have been used instead of excluding
        \texttt{DATA}.

\end{eqlist}

\subsection{Overlapping Rules}

\label{overlapping rules in implementation}

The advantages and difficulties of overlapping rules have already been
addressed in \sectionref{overlapping rules in architecture} and will not be
repeated here.  \parsername{} does not try to detect overlapping rules;
that responsibility is left to the author of the rules.  A mechanism is
provided for ordering overlapping rules: the priority field in each rule;
negative priorities may be useful for catchall rules.

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

\begin{itemize}

    \item Sort rules by program and regex, then visually inspect the list,
        e.g.\ with \acronym{SQL} similar to:    \newline{}
        \verb!SELECT program, regex!            \newline{}
        \verb!    FROM rules!                   \newline{}
        \verb!    ORDER BY program, regex;!     \newline{}
        The rules are sorted first by program, then by regex, because rules
        with overlapping regexes will not overlap if their programs are
        different.  Note that this does not properly deal with generic
        rules whose program is \texttt{*}; those rules will be used on all
        log lines that have not been recognised by program-specific rules.

    \item Compare the results of parsing using different rule orderings, as
        described in \sectionref{rule ordering for efficiency}.  Parse
        several log files using optimal ordering, then dump a textual
        representation of the rules, connections, and results tables.
        Repeat with shuffled and reversed ordering, starting with a fresh
        database.  If there are no overlapping rules the tables from each
        run will be identical; differences indicate overlapping rules.
        The rules that overlap can be determined by examining the
        differences in the tables: each result contains a reference to the
        rule which created it, which will change if that rule overlaps with
        another.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the log files that are recognised by more
        than one rule.

\end{itemize}

\input{logparser-complications}

\section{Limitations and Possible Improvements}

\label{limitations and improvements in implementation}

Every piece of software suffers from some limitations, and there is almost
always room for improvement.  Below are the limitations and possible
improvements that have been identified in \parsername{}.

\subsection{Limitations}

\label{logging helo}

\begin{enumerate}

    \item Each new Postfix release requires writing new rules or modifying
        existing rules to cope with the new or changed log lines.
        Similarly using a new \acronym{DNSBL}, a new policy server, or new
        administrator-defined rejection messages also requires new rules.

    \item The hostname used in the HELO command is not logged if the
        incoming delivery attempt is successful.  Configuring Postfix to do
        this is relatively simple; create a restriction that is guaranteed
        to warn for every accepted mail, as follows:

        \begin{enumerate}

            \item Create \texttt{/etc/postfix/log\_helo.pcre}
                containing:\newline{}
                \tab{}\texttt{/\^/~~~~WARN~Logging~HELO}

            \item Modify \texttt{smtpd\_data\_restrictions} in
                \texttt{/etc/postfix/main.cf} to contain:\newline{}
                \tab{}\texttt{check\_helo\_access~pcre:/etc/postfix/log\_helo.pcre}

        \end{enumerate}

        Although \texttt{smtpd\_helo\_restrictions} seems like the natural
        place to log the HELO hostname, when it is evaluated for the first
        recipient there will not yet be a queueid allocated for the
        delivery attempt, so the log line cannot be associated with the
        correct mail.  There is guaranteed to be a queueid when the DATA
        command has been reached, and thus the queueid will be logged by
        any restrictions taking effect in
        \texttt{smtpd\_data\_restrictions}, and the log line can be
        associated with the correct mail.  There is no difficulty in
        specifying a HELO-based restriction in
        \texttt{smtpd\_data\_restrictions}, Postfix will perform the check
        correctly.

        Logging the HELO hostname in this fashion also partially prevents
        the complication described in \sectionref{Mail deleted before
        delivery is attempted} from occurring, but only when there is a
        single recipient; in that case the recipient address will be logged
        also, but when there are multiple recipients no addresses are
        logged.

    \item \parsername{} does not create separate mails where one or more
        delivery attempts are rejected and subsequently a mail is accepted;
        it will appear in the database as one mail with lots of rejections
        followed by acceptance (this has already been mentioned in
        \sectionref{connection reuse}).  It does not appear to be possible
        to make this distinction given the data Postfix logs, though it
        might be possible to write a policy server to provide additional
        logging.

    \item \parsername{} will not detect that it is parsing the same log file
        twice, resulting in the database containing duplicate entries.

    \item \parsername{} does not distinguish between log files produced by
        different sources when parsing; all results will be saved to the
        same database.  This may be viewed as an advantage, because log
        files from different sources can be combined in the same database,
        or it may be viewed as a limitation because there is no facility to
        distinguish between log files from different sources in the same
        database.  If the results of parsing log files from different
        sources must remain separate, \parsername{} can easily be
        instructed to use a different database.

    \item The solution to complication \sectionref{Mails deleted during
        delivery} involves discarding data.

\end{enumerate}

\subsection{Possible Improvements}

\begin{enumerate}

    \item Investigate and write the policy server referred to in limitation
        3 above.

    \item Improve the solution to complication \sectionref{Mails deleted
        during delivery} so that data is not discarded.

\end{enumerate}

\section{Summary}

This section has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the resemblance between a database schema and an
\acronym{API} is described, followed by a detailed description and a
diagram of the database schema.  The framework is documented next,
including the steps it takes during initialisation, the parsing process,
and the conveniences it offers to users.  The data collected by the
framework is described, as are the ways in which various optimisations can
be disabled to demonstrate their effect, and the debugging options the
framework provides.  The implementation of actions in \parsername{} is
documented, including how frequently each action is specified by rules, and
brief descriptions of why some actions are more popular than others.  All
the actions that are part of \parsername{} are described in detail,
followed by the process of adding a new action to the parser.  The final
component of the architecture, the rules, is also the most visible
component, and its implementation is examined in detail with: an example
rule and an explanation of every field; a description of how to add new
rules and determine the values that should be used for each field; an
explanation of the algorithm used by the utility that creates new regexes
based on unparsed log lines, and how it differs from the original algorithm
it is based on.  How \parsername{} uses rule conditions and overlapping
rules is discussed, accompanied by a description of the regex snippets the
framework provides to simplify the regexes used in rules.  The many
complications and difficulties encountered while writing \parsername{}, and
the solutions developed to overcome them, are documented in depth, also
documented are how some solutions interact and which action or actions the
solution is implemented in.  This chapter concludes with a list of the
limitations identified in \parsername{}, followed by some possible
improvements suggested by the author.