### 2009-03-31 doc/logparser.tex
Added a new action, DELIVERY_ERROR, to save data from delivery errors that     happen before a disconnection. Changed some UNINTERESTING rules to use DELIVERY_ERROR and SAVE_DATA. Removed connection_data fields from UNINTERESTING rules. Added a debug field to rules, so that individual rules can have debugging     information dumped when they recognise a log line. Improved the abstract, though it still needs work. Documented everything in the thesis.

% $Id$
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

% Aligned list environments.
\usepackage{eqlist}
% This voodoo below gives us an eqlist environment with bold items.
\newcommand{\bolditem}[1]{%
    \bfseries#1%
}
\newenvironment{boldeqlist}
    {\begin{eqlist}[\renewcommand{\makelabel}{\bolditem}]}
    {\end{eqlist}}

% Enable customised lists.
\usepackage{enumitem}
% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

% Extra packages recommended by Springer.
%\usepackage{mathptmx}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{makeidx}
%\usepackage{multicol}

% Embed SVN Id etc.
\usepackage{svn}

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Warnings rather than errors when references cross page boundaries.
\vrefwarning{}
% Errors when loops are encountered
\vrefshowerrors{}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

% Sort numbers where there are multiple citations.  Does not appear to have
% the expected effect (probably clashes with hyperref), though it does
% reduce the space between numbers.
\usepackage{cite}
% Check for unused references, and how the citation key (e.g.\ slct-paper)
% in the margin beside the reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used; neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the defaul spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    % XXX REVISIT THE PLACEMENT OF FIGURES AND TABLES, THOUGH I THINK
    % LATEX'S DEFAULTS ARE PROBABLY PRETTY GOOD.
    \begin{figure}%[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[3]{
    \begin{table}%[ht]
        \caption{#2}\label{#3}
        \input{#1}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \hline%
    \noalign{\smallskip}%
}

\newcommand{\tablebottomline}[0]{%
    \noalign{\smallskip}%
    \hline%
}

\newcommand{\tablemiddleline}[0]{%
    \noalign{\smallskip}%
    \hline%
    \noalign{\smallskip}%
}

% A command to format a Postfix daemon's name
\newcommand{\daemon}[1]{%
    \texttt{postfix/#1}%
}

\newcommand{\flowchart}[2]{%
    (action: \action{#1}\@; flowchart:~#2)%
}

\newcommand{\action}[1]{%
    \texttt{#1}%
}

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}

\newcommand{\tab}[0]{%
    \hspace*{1em}%
}

% Constant values.
\newcommand{\numberOFlogFILES}[0]{%
    93%
}

% XXX UPDATE THIS WHEN all-logs PASSES
\newcommand{\numberOFlogFILESall}[0]{%
    522%
}

\newcommand{\numberOFrules}[0]{%
    184%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.50\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.00\%%
}

\newcommand{\numberOFlogLINES}[0]{%
    60,721,709%
}

\newcommand{\numberOFlogLINEShuman}[0]{%
    60.72 million%
}

\newcommand{\numberOFactions}[0]{%
    23%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    16,836%
}

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\acronym{PLP}}
\newcommand{\parsernames}{\acronym['s]{PLP}}
\newcommand{\parsernamelong}{Postfix Log Parser}
\newcommand{\parsernameshort}{PLP}

\begin{document}

\title{Parsing Postfix log files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

\begin{abstract}

    \setcounter{page}{2}
    \phantomsection{}
    \addcontentsline{toc}{chapter}{Abstract}

    XXX REWRITE WHEN THESIS IS FINISHED\@.

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding Postfix log files and reconstructing the actions taken by
    Postfix that generated the log files.  This thesis describes the
    creation of a parser for Postfix log files, documenting the
    architecture developed and the parser implementing it, explaining the
    difficulties encountered and the solutions developed, with reference to
    an implementation that stores data gleaned from the log files in an SQL
    database.  The gathered data can be used to optimise current anti-spam
    measures, to produce statistics showing how effective those measures
    are, or to provide a baseline to test new anti-spam measures against.
    The Postfix log file parser needs to be very precise and strict when
    parsing, yet must allow users to easily adapt or extend it to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  The newly developed
    architecture aims to make the process of parsing new inputs as simple
    as possible, enabling users to trivially add new rules (to parse
    variants of existing inputs) and relatively easily add new actions (to
    process a previously unknown category of inputs).  The architecture
    scales linearly or better as the number of rules and size of input
    increases, making it suitable for parsing large corpora or months of
    accumulated data.

    \SVN$Id$
    \begin{center}
        \SVNId{}
    \end{center}

    XXX FIX THE PROBLEMS THAT AROSE FROM THE LAST all-logs RUN\@.

    XXX RUN ``test-some'' AGAIN TO UPDATE THE DB\@.

    XXX SHOULD I HAVE A ``FUTURE WORK'' SECTION SOMEWHERE IN THE
    CONCLUSION\@?

    XXX MAIL MADS ABOUT EXAMINERS\@.

\end{abstract}

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\setcounter{page}{3}
\tableofcontents
\listoffigures
\listoftables

\newpage

% WSUIPA fonts.
\input{ipamacs}

\input{logparser-introduction}
\input{logparser-background}
\input{logparser-other-parsers-review}
\input{logparser-architecture}
\input{logparser-implementation}
\input{logparser-results}
\input{logparser-conclusion}

\appendix

\bibliographystyle{logparser-bibliography-style}
\bibliography{logparser-bibliography}

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

\end{document}