### 2009-04-12 doc/logparser-background.tex
Ran diction and made changes based on its suggestions.

\chapter{Background}

\label{background}

This section provides background information helpful in understanding the
remainder of this thesis.  It begins with a discussion of the motivation
underlying the project, followed by a brief introduction to \acronym{SMTP},
and finishes with a longer introduction to Postfix, concentrating on
anti-spam restrictions and policy servers.

\section{Motivation}

\label{motivation}

This work is part of a larger project to optimise a mail server's
Postfix-based anti-spam restrictions, generate statistics and graphs, and
provide a platform on which new restrictions can be evaluated to determine
if they are beneficial in the fight against spam.  The program written for
this project, \parsername{}, parses Postfix log files and populates a
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

A short example of the optimisation possible using data from the database
is determining which Postfix restrictions reject the highest number of
mails:

\begin{verbatim}
SELECT name, description, restriction_name, hits_total
    FROM rules
    WHERE action = 'DELIVERY_REJECTED'
    ORDER BY hits_total DESC;
\end{verbatim}

If the database supports sub-selects (where the results of one query are
used as a parameter in another), percentages can be obtained for the top
ten restrictions (sample output shown in \tableref{Sample output from SQL
query}):

\input{build/include-sample-sql-query}

\begin{table}[thbp]
    \caption{Sample output from SQL query, showing the percentage of
    delivery attempts rejected by top ten restrictions}
    \empty{}\label{Sample output from SQL query}
    \begin{tabular}[]{lrr}
        \tabletopline{}%
        Restriction & Number of  & Percentage of    \\
                    & rejections & total rejections \\
        \tablemiddleline{}%
        \input{build/include-sample-sql-output}
        \tablebottomline{}%
    \end{tabular}

    All percentages above are exclusive, i.e.\ the first restriction's
    number of rejections does not include the second restriction's number
    of rejections.

\end{table}

Another example is determining which restrictions are not effective: this
sample query shows which restrictions had fewer than 100 rejections in the
last log file parsed, and the percentage of total rejections each of those
restrictions represents.

\begin{verbatim}
SELECT name, description, restriction_name, hits,
        (hits * 100.0 /
            (SELECT SUM(hits)
                FROM rules
                WHERE action = 'DELIVERY_REJECTED'
            )
        ) || '%' AS percentage
    FROM rules
    WHERE action = 'DELIVERY_REJECTED'
        AND hits < 100
    ORDER BY hits ASC;
\end{verbatim}

The database queries above yield summary statistics about the efficiency of
anti-spam techniques.  These statistics are far less feasible to assess
directly from log files without prior pre-processing into a database in the
fashion proposed, implemented, and tested herein.

\section{Simple Mail Transfer Protocol (SMTP)}

\SMTPglossaryDescription{} Despite the simplicity of the protocol many
virus and spam sending programs fail to implement it properly, so requiring
strict adherence to the protocol specification is beneficial in protecting
against spam and viruses.\footnote{Originally all mail servers adhered to
the principle of \textit{Be liberal in what you accept, and conservative in
what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but unfortunately
that principle was written in a friendlier time.  Given the deluge of spam
that mail servers are subjected to daily, a more appropriate maxim could
be: \textit{Require strict adherence to relevant \acronyms{RFC}; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  It is neither as friendly nor as
catchy, but it more accurately reflects the current situation.} A typical
\acronym{SMTP} conversation resembles the following (the lines starting
with a three digit number are sent by the server, all other lines are sent
by the client):

\begin{verbatim}
220 smtp.example.com ESMTP
HELO client.example.com
250 smtp.example.com
MAIL FROM: <alice@example.com>
250 2.1.0 Ok
RCPT TO: <bob@example.com>
250 2.1.5 Ok
DATA
354 End data with <CR><LF>.<CR><LF>
Message headers and body sent here.
.
250 2.0.0 Ok: queued as D7AFA38BA
QUIT
221 2.0.0 Bye
\end{verbatim}

An example deviation from the protocol:

\begin{verbatim}
220 smtp.example.com ESMTP
HELO client.example.com
250 smtp.example.com
MAIL FROM: Alice in Distribution alice@example.com
501 5.1.7 Bad sender address syntax
RCPT TO: Bob in Sales/Marketing bob@example.com
503 5.5.1 Error: need MAIL command
DATA
503 5.5.1 Error: need RCPT command
Message headers and body sent here.
.
502 5.5.2 Error: command not recognized
QUIT
221 2.0.0 Bye
\end{verbatim}

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, other programs,
particularly newer programs, were written by competent programmers or send
mail using well written programs (e.g.\ Postfix or Sendmail on Unix hosts,
Microsoft Outlook on Windows hosts).  Traditionally a mail server would
have done its best to deal with deficient clients, with the intention of
accepting as much mail sent to its users as possible, e.g.\ by ignoring the
absence of a HELO command, or by accepting sender or recipient addresses
that were not enclosed in \texttt{<>}.

\section{Postfix}

\label{postfix background}

Postfix is a \glsfirst{MTA} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to employ those restrictions which they judge
suitable for their server's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
different levels of stricture and permissiveness.  Administrators can
supply their own rejection messages to make it clear to senders exactly why
their mail was rejected.  Policy servers (\sectionref{policy servers})
provide a simple way to write new restrictions without having to modify
Postfix's source code.  Unfortunately this flexibility has a cost:
complexity in the log files generated.  Although it is relatively simple in
most cases to use standard Unix text processing utilities to determine the
fate of an individual mail, sometimes it can be quite difficult.  For
most mails the journey is simple and brief, but the remainder can have
quite complex journeys (see \sectionref{complications} for details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple components that each perform one of the tasks
required of an \acronym{MTA}, e.g.\ receive mail, send mail, local delivery
of mail;\ full details can be found in~\cite{postfix-overview}.  Postfix's
design is strongly influenced by security concerns: those components that
interact with other hosts are not privileged,\footnote{Privilege means the
power to perform actions that are limited to the administrator, and not
available to ordinary users.} so bugs in those components will not give an
attacker extra privileges; those components that are privileged do not
interact with other hosts, making it much more difficult for an attacker to
exploit any bugs that may exist in those components.

\subsection{Mixing And Matching Postfix Restrictions}

\label{Mixing and matching Postfix restrictions}

Postfix restrictions are documented fully in~\cite{smtpd_access_readme,
smtpd_per_user_control, policy-servers}; the following is only a brief
introduction.

Postfix uses one restriction list (containing zero or more restrictions)
for each stage of the \acronym{SMTP} conversation: client connection, HELO
command, MAIL FROM command, RCPT TO commands, DATA command, and end of
data.  The appropriate restriction list is evaluated for each stage, though
by default the restriction lists for client connection, HELO, and MAIL FROM
commands will not be evaluated until the first RCPT TO command is received,
because some clients do not deal properly with rejections before this
stage; a benefit of this delay is that Postfix has more information
available when logging rejections.

Postfix uses simple lookup tables to make decisions when evaluating some
restrictions, e.g.\newline{}
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}

\begin{eqlist}

    \item [check\_client\_access] The name of the restriction to evaluate.

    \item [cidr] The type of the lookup table.

    \item [/etc/postfix/client\_access] The file containing the lookup
        table.

\end{eqlist}

The restriction \texttt{check\_client\_access} checks if the \acronym{IP}
address of the connected client is found in the specified table and returns
the associated action if found; the method of searching the file is
dependant on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\newline{}
\tab{}\texttt{reject\_rbl\_client dnsbl.example.com}\newline{} checks the
\acronym{DNSBL} \texttt{dnsbl.example.com} and rejects the command if the
client's \acronym{IP} address is listed.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}.  The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  E.g.\ the general rule is that machines
on the local network must authenticate before sending mail, except for the
web server, because the legacy applications running on it lack
authentication support; this configuration example shows how:

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    <other restrictions>
    permit_sasl_authenticated,
    check_client_access /etc/postfix/allow_webserver.cidr,
    <other restrictions>

  /etc/postfix/allow_webserver.cidr:
    192.0.2.80/32   dunno
    192.0.2.0/24    reject "Please authenticate to send mail"
\end{verbatim}

That example also shows that it is trivial to supply a custom rejection
message.  When the result is the name of another restriction Postfix will
evaluate that restriction, allowing restrictions to be chosen based on the
client \acronym{IP} address, client hostname, HELO hostname, sender
address, or recipient address.  E.g.\ the administrator may require that
all clients on the local network have valid DNS entries, to prevent people
sending mail from unknown machines; one example of how to achieve this is:

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    <other restrictions>
    check_client_access /etc/postfix/require_dns_entries.cidr,
    <other restrictions>

  /etc/postfix/require_dns_entries.cidr:
    192.0.2.0/24    reject_unknown_client_hostname
\end{verbatim}

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex user-defined sequences
of lookups, restrictions, and exceptions.  Postfix tries to protect the
administrator from misconfiguration in as far as is reasonable, e.g.\ the
restriction \texttt{check\_helo\_mx\_access} cannot cause a mail to be
accepted, because the parameter it checks (the hostname given in the HELO
command) is under the control of the remote client.  Despite this, it is
possible for the administrator to make catastrophic mistakes, e.g.\
rejecting all mail --- the administrator must be cognisant of the effects
their configuration changes will have.  This is similar to one of UNIX's
design philosophies: \textit{``UNIX was not designed to stop its users from
doing stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

\subsection{Policy Servers}

\label{policy servers}

A policy server~\cite{policy-servers} is an external program consulted by
Postfix to determine the fate of an \acronym{SMTP} command.  The policy
server is given state information by Postfix (sample state information is
shown in \tableref{Example attributes sent to policy servers}) and returns
a result (reject, permit, dunno, a rejection name) as described in
\sectionref{Mixing and matching Postfix restrictions}.  A policy server can
perform more complex checks than those provided by Postfix: a trivial
example is allowing addresses associated with the payroll system to send
mail on the third Tuesday after pay day only, to help prevent problems from
phishing mails using faked sender addresses.  E.g.\ a phishing mail might
claim that the payroll system had a disastrous disk failure, and until the
server is replaced all salary payments will have to be processed manually,
so please reply to this mail with your name, address, and bank account
details; the criminal can then use any details sent to him to help with
identity theft.

\begin{table}[thbp]
    \caption{Example attributes sent to policy servers, taken
    from~\cite{policy-servers}}
    \empty{}\label{Example attributes sent to policy servers}
    \centering{}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Attribute name          & Attribute value           \\
        \tablemiddleline{}%
        request                 & smtpd\_access\_policy     \\
        protocol\_state         & RCPT                      \\
        protocol\_name          & SMTP                      \\
        helo\_name              & some.domain.tld           \\
        queue\_id               & 8045F2AB23                \\
        sender                  & foo@bar.tld               \\
        recipient               & bar@foo.tld               \\
        recipient\_count        & 0                         \\
        client\_address         & 1.2.3.4                   \\
        client\_name            & another.domain.tld        \\
        reverse\_client\_name   & another.domain.tld        \\
        instance                & 123.456.7                 \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

Some widely deployed policy servers:

\begin{itemize}

    \item \acronym{SPF}~\cite{openspf}.  \acronym{SPF} records specify
        which mail servers are allowed to send mail using sender addresses
        from a particular domain.  The intention is to reduce spam from
        faked sender addresses, backscatter~\cite{postfix-backscatter}, and
        Joe~jobs~\glsadd{Joe-job}.  There has been considerable resistance
        to \acronym{SPF} because it breaks or vastly complicates some
        commonly-used features of \acronym{SMTP}, e.g.\ forwarding mail
        from one domain to another when a user moves.

    \item Greylisting~\cite{greylisting} is a technique that temporarily
        rejects a delivery attempt when the triple of \newline{}
        \tab{}\texttt{(sender address, recipient address, remote
        \acronym{IP} address)}\newline{} has not been seen before; on
        second and subsequent delivery attempts from that triple the mail
        will be accepted.  This blocks spam from some senders because
        maintaining a list of failed addresses and retrying after a
        temporary failure is uneconomical for a spam sender, but a
        legitimate mail server must retry deliveries that failed
        temporarily.  Sadly spam senders are using increasingly complex and
        well written programs to distribute spam, frequently using an ISP
        provided \acronym{SMTP} server from a compromised machine on the
        ISP's network.  Greylisting will slowly become less effective as
        spam senders adapt, but it does block a large percentage of spam
        mail at the moment; the most effective restrictions from the
        \numberOFlogFILES{} log files used when generating the results in
        \sectionref{Evaluation} are shown in \tableref{Most effective
        Postfix restrictions}.  The table shows that greylisting is worth
        using at the moment, particularly when you take into account its
        position as the final restriction that a mail must overcome in the
        configuration used on the mail server that generated the log files:
        on that server greylisting only takes effect for mails that have
        passed all other restrictions.  Some problems may be encountered
        when using greylisting: some servers fail to retry after a
        temporary failure, or legitimate mail may be delayed, particularly
        when coming from a pool of servers.

        \showtable{build/include-restriction-summary-table}{Most effective
        Postfix restrictions}{Most effective Postfix restrictions}

    \item Scoring systems such as Policyd-weight~\cite{policyd-weight}
        perform tests on features of the delivery attempt (e.g.\
        \acronym{IP} address, sender address), incrementing or decrementing
        a score based on the results; if the eventual score is higher than
        a threshold the mail is rejected.  The administrator must manually
        whitelist clients if they are to bypass a Postfix restriction,
        whereas using a threshold that requires a delivery attempt to hit
        several scored restrictions will allow mail from clients that would
        fall foul of one binary restriction only.

\end{itemize}

\section{Summary}

This section has provided background information useful in understanding
this thesis, starting with the motivation behind the project, continuing
with an introduction to \acronym{SMTP}, and finishing with Postfix, its
anti-spam restrictions, and its support for policy servers.