### 2009-04-12 doc/logparser-implementation.tex

\chapter{Postfix Parser Implementation}

\label{Postfix Parser Implementation}

This chapter documents the implementation of \parsername{}, based on the
architecture described in \sectionref{parser architecture}.  Any design may
look good when examined in the abstract, but the real test of the design
comes with the first concrete implementation; only then do any practical
difficulties come to light.  Implementing this architecture was relatively
straightforward: the difficulties came not from the architecture and
design, but from anomalies in the log files and Postfix's behaviour, as
described in \sectionref{complications}.  \parsername{} successfully deals
with all the difficulties that were discovered during its development, but
further difficulties may arise during future usage; unfortunately solving
all potential difficulties that may arise is an intractable task, but the
descriptions of the solutions developed thus far (see
\sectionref{complications}) should help if someone needs to develop a
solution of their own.

\parsername{} deals with all the complications of parsing, the
eccentricities and oddities of Postfix log files, presenting the resulting
data in a normalised, simple to use representation.  Unfortunately dealing
with the complications that arise sometimes requires the parser to discard
log lines (e.g.\ \sectionref{aborted delivery attempts}), and, less
frequently, to discard a data structure (e.g.\ \sectionref{timeouts during
data phase}).  \parsername{} can parse log files from Postfix version 2.2
through to version 2.5, and should be capable of parsing log files from
later versions with only minor modifications or an updated ruleset.  The
parser's task is to follow the journey each mail takes through Postfix,
combining the data captured by rules into a coherent whole, saving it in a
useful and consistent form, and performing housekeeping duties.  The
intermingling of log lines from different mails immediately rules out the
possibility of handling each mail in isolation; the parser must handle
multiple mails in parallel, each potentially at a different stage in its
journey, without any interference between mails --- except in the minority
of cases where intra-mail interference is required, e.g.\ mail re-injected
for forwarding (\sectionref{Re-injected mails}).  The best way to deal with
intermingling of log lines is to maintain state information for every
unfinished mail, and to manipulate the appropriate mail correctly for each
log line encountered.

This chapter begins with the assumptions under which \parsername{} was
designed and written, followed by a flowchart showing the most common paths
taken through Postfix and \parsername{}, with a description of the stages
and transitions.  The topic of the next section is the database that
provides storage for the parser: any future work that analyses gathered
data will do so using the database, so the database schema acts as an
\acronym{API}.  A diagram of the database schema is provided, plus
documentation for every table and field.

The next three sections document the implementation of the three components
of the architecture.  First is the framework, including its initialisation
phase, the parsing process, and the conveniences it offers to users; that
section finishes with the features that are important for this thesis but
not strictly necessary for parsing: the performance data collected by the
framework, the optimisations that can be disabled to show their effect, and
the debugging options it provides.  The second component of the
architecture is the actions: their implementation is documented, with a
graph of how often each action is specified by rules, and a description of
why some actions are more popular than others.  Every action that is part
of \parsername{} is documented, and the actions section concludes with the
process of adding a new action, and how to determine the values for each of
the required fields.  Rules are the final component of the architecture,
and also the most visible to advanced users, e.g.\ systems administrators,
because it is likely that they will need to add their own rules to
recognise their log lines.  An example rule used by \parsername{} is shown,
with every field clearly documented; this is followed by a description of
adding new rules, and how to determine the value of each of the fields.
\parsername{} provides a utility to create regexes from unrecognised log
lines; the algorithm it uses is documented, including the differences
between it and the original algorithm.  The rules section finishes with a
discussion of how \parsername{} uses rule conditions and overlapping rules,
plus a description of the regex snippets provided to ease the process of
writing regexes.

On first inspection, Postfix log files look like they will be simple to
parse, but this impression turns out to be incorrect.  The many
complications and difficulties encountered while writing \parsername{}, and
the solutions developed to overcome them, are documented in
\sectionref{complications}.  This chapter finishes with a list of
\parsernames{} limitations, and some possible improvements that could be
implemented.

\section{Assumptions}

\parsername{} makes a small number of (hopefully safe and reasonable)
assumptions:

\begin{itemize}

    \item The log files are whole and complete: nothing has been removed,
        either deliberately or accidentally (e.g.\ log file rotation gone
        awry, file system filling up, logging system unable to cope with
        the volume of log messages).  On a well run system it is extremely
        unlikely that any of these problems will arise, though it is of
        course possible, particularly when suffering from a deluge of spam
        or a mail loop.  When parsing individual log files in isolation,
        it is highly likely that some mails will have log lines in previous
        log files, and others will have log lines in subsequent log files;
        to alleviate this problem \parsername{} supports saving and loading
        its state tables, so they can be preserved between log files.

    \item Postfix logs enough information to make it possible to accurately
        reconstruct the actions it has taken.  Heuristics are used in
        several places when parsing; see \sectionref{identifying bounce
        notifications}, \sectionref{aborted delivery attempts}, and
        \sectionref{pickup logging after cleanup} for details.  At least
        one difficulty encountered while writing \parsername{}
        (\sectionref{yet more aborted delivery attempts}) could not be
        solved using the data in the log files, and requires a brute force
        approach.

    \item The Postfix queue has not been tampered with, causing unexplained
        appearance or disappearance of mail.  This may happen if the
        administrator deletes mail from the queue without using
        \daemon{postsuper}, or if the server suffers from filesystem
        corruption.

\end{itemize}

In some ways this task is similar to reverse engineering or replicating a
black box system based solely on its inputs and outputs.  Thus far,
analysis of the log files has been enough to reconstruct Postfix's
behaviour, but for more difficult programs the techniques described in
\cite{black-box-error-reporting} may be useful.  Some advantages come from
treating Postfix as a black box during parser development:

\begin{itemize}

    \item Reading and understanding the source code would require a
        significant investment of time: Postfix 2.5.5 has 17MB of source
        code.  Each subsequent version would require further work to
        investigate the changes; many of those changes, although they
        improve Postfix's internals, would not have any effect on its
        externally observable behaviour.

    \item \parsername{} is developed using real world log files rather than
        the idealised log files someone would naturally envisage when
        reading the source code, which cannot accurately communicate the
        variety of orderings in which log lines are found in log files.

    \item The parser acts as a second source of information about Postfix's
        operation, based on empirical evidence.  A separate project could
        compare the empirical knowledge inherent in \parsername{} with
        Postfix's documentation and source code to see how closely the two
        agree.

\end{itemize}

\section{Parser Flow Chart}

\label{flow chart}

The flow chart in \figureref{flow chart image} shows the most common paths
a connection or mail can take through \parsername{}; decision boxes and the
difficulties described in \sectionref{complications} are excluded for the
sake of clarity.  The flow chart is intended to be a graphical overview of
how a mail progresses through both Postfix and \parsername{}, providing an
overall context into which the detailed descriptions in the remainder of
this chapter will fit, in particular the actions (\sectionref{actions in
implementation}) and complications (\sectionref{complications}).  The
states and transitions shown in \figureref{flow chart image} will be
explained later in this section.

\showgraph{build/logparser-flow-chart-part-1}{\parsernamelong{} flow
chart}{flow chart image}

Everything starts off with a mail entering the system, whether by local
submission via \daemon{postdrop}, by \acronym{SMTP}, by re-injection
because of forwarding, or generated internally by Postfix.  Local
submission is the simplest of the four: a queueid is assigned immediately
\flowchart{PICKUP}{2}, and the mail moves on to the delivery stage.
Re-injection because of forwarding lacks explicit log lines of its own; it
is explained fully in \sectionref{Re-injected mails}.  Internally generated
mails lack any explicit origin in Postfix 2.2.x and must be detected using
heuristics as described in \sectionref{identifying bounce notifications};
later versions of Postfix do provide log lines for internally generated
mails \flowchart{BOUNCE\_CREATED}{3}.  Bounce notifications are the primary
example of internally generated mails, though other types exist, e.g.\
Postfix may generate mails to the administrator when it encounters
configuration errors.

\acronym{SMTP} is more complicated than the others:

\begin{enumerate}

    \item The remote client connects \flowchart{CONNECT}{1}.

    \item This is followed by rejection of \acronym{SMTP} commands by the
        restrictions configured by the administrator
        \flowchart{DELIVERY\_REJECTED}{4}; acceptance of one or more mails
        \flowchart{CLONE}{5}; failure of the remote client's connection
        \flowchart{DELIVERY\_ERROR}{6}; or some random interleaving of two
        or more of the above.

    \item The client disconnects \flowchart{DISCONNECT, TIMEOUT}{6}, either
        normally or with an error.  If Postfix has rejected any
        \acronym{SMTP} commands the resulting data will be saved to the
        database; if there were no rejections there will not be any data to
        save.  Any mails accepted will already have been cloned so their
        data is in another data structure, and will be delivered in the
        same way as mails that entered the system by any other route.

\end{enumerate}

The obvious counterpart to mail entering the system is mail leaving the
system, whether by deletion, bouncing, local delivery, or remote delivery.
All four are handled in exactly the same way:

\begin{enumerate}

    \item The mail will have one or more delivery attempts
        \flowchart{MAIL\_SENT, SAVE\_DATA}{8}.

    \item Sometimes mail is re-injected and the child mail needs to be
        tracked with the parent mail \flowchart{TRACK}{9}; the handling of
        re-injected mails is described in \sectionref{tracking re-injected
        mail}.

    \item After one or more delivery attempts the mail will be delivered
        \flowchart{MAIL\_SENT}{8}, bounced \flowchart{MAIL\_BOUNCED}{8},
        expired \flowchart{EXPIRY}{8}, or deleted by the administrator
        \flowchart{DELETE}{8}.

    \item This is the last log line for this particular mail, though it may
        be indirectly referred to if it was re-injected.  The mail is
        cleaned up and entered in the database, then deleted from the state
        tables \flowchart{COMMIT}{11}.

\end{enumerate}

It should be emphasised that the mail delivery process above happens
whether the mail is delivered to a mailbox, piped to a command, delivered
to a remote server, bounced (because of a mail loop, delivery failure, or
five day timeout), or deleted by the administrator, \textit{unless\/} the
mail is either parent or child of re-injection, as explained in
\sectionref{tracking re-injected mail}.

\section{Database}

\label{database}

A database is used to store both the rules and the data gleaned by parsing
Postfix log files.  Understanding the database schema is important in
understanding the actions of the parser, and essential to developing
further applications that utilise the data gathered; \sectionref{database
as API} describes how the database schema functions as an \acronym{API}.

The database schema can be conceptually divided in two: the rules used to
recognise log lines, and the data saved from the parsing of log files.
Each rule has a regex to recognise log lines and capture data from them,
and specifies the action to be invoked when a log line is recognised; they
also have several fields that aid the user in understanding the meaning of
the log lines recognised by each rule.  The rules are described in detail
in \sectionref{rules in implementation}, but the rules table is documented
in \sectionref{rules table} with the rest of the database schema.

The data saved from parsing the log files is also divided into two tables:
connections and results.  The connections table contains an entry for every
mail accepted and every connection that rejected an \acronym{SMTP} command;
the individual fields will be described in \sectionref{connections table}.
The results table will have at least one entry for each entry in the
connections table; its fields will be covered in detail in
\sectionref{results table}.  A diagram of the database schema is provided
in \figureref{Diagram of the Database Schema picture}, to complement the
in-depth descriptions of each table.

An important but easily overlooked benefit of storing the rules in the
database is the link between rules and results: if more information is
required when examining a result, the rule that produced the result is
available for inspection because each result references the rule that
created it.  No ambiguity or confusion is possible about which rule
resulted in a particular result, eliminating one potential source of
confusion.

A clear, comprehensible schema is essential when using the data extracted
from log lines; it is more important when using the data than when storing
it, because storing the data is a write-once operation, whereas utilising
the data requires frequent searching, sorting, and manipulation of the data
to produce customised reports or statistics.

\subsection{Using A Database To Provide An Application Programming Interface}

\label{database as API}

The database populated by \parsername{} provides a simple interface to
Postfix log files.  Although the interface is a database schema rather than
a set of functions in a library, it gives the same benefits as any other
\acronym{API}: it provides a stable interface between the user and the
creator of the data, allowing code on either side of the interface to be
changed without adverse effects, as long as the interface is adhered to.
Programs that use the database can range from the simple examples in
\sectionref{motivation} to far more complex data mining tools.

Using a database simplifies writing programs that need to interact with the
data in several ways:

\begin{enumerate}

    \item Most programming languages have facilities for database access,
        allowing a developer to write programs that use the gathered data
        in their preferred programming language, rather than being
        restricted to the language the parser is written in.

    \item Databases provide complex querying and sorting functionality for
        the user without requiring large amounts of programming.  All
        databases provide a program, of varying complexity and
        sophistication, that can be used for ad hoc queries with minimal
        investment of time.

    \item Databases are easily extensible, e.g.:

        \begin{itemize}

            \item New columns can be added to the tables used by the
                program, using DEFAULT clauses or TRIGGERS to populate
                them.

            \item A VIEW gives a custom arrangement of data with minimal
                effort.

            \item Triggers can be written to perform actions when certain
                events occur.  In pseudo-\acronym{SQL}\@:

\begin{verbatim}
CREATE TRIGGER ON INSERT INTO results
    WHERE sender = "boss@example.com"
        AND action = "DELIVERY_REJECTED"
    SEND PANIC EMAIL TO "postmaster@example.com";
\end{verbatim}

            \item Other tables can be added to the database, e.g.\ to cache
                historical or computed data, or to link in data from other
                sources.

        \end{itemize}

    \item Some databases support granting access on a fine-grained basis,
        e.g.\ allowing the finance department to produce invoices, the
        helpdesk to run limited queries as part of dealing with support
        calls, and the administrators to have full access to the data.

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with \acronym{SQL}, lots of
        resources are available from which to learn, e.g.\
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
        Although every vendor implements a different dialect of
        \acronym{SQL}, the basics are the same everywhere.  Depending on
        the database in use there may be tools available that reduce or
        remove the requirement to know \acronym{SQL}.

\end{enumerate}

Storing the results in a database will also increase the efficiency of
using those results, because the log files need only be parsed once rather
than each time the data is used; indeed the results may be used by someone
with no access to the original log files.

\subsection{Rules Table}

\label{rules table}

\newlength{\belowcaptionskipORIG}
\setlength{\belowcaptionskipORIG}{\belowcaptionskip}
\setlength{\belowcaptionskip}{10pt}
\showgraph{build/database-schema}{Diagram of the Database Schema}{Diagram
of the Database Schema picture}
\setlength{\belowcaptionskip}{\belowcaptionskipORIG}

Rules are discussed in detail in \sectionref{rules in implementation}, but
the structure of the rules table is documented here along with the other
tables in the database.  Rules are created by the user, not the parser, and
will not be modified by \parsername{} (except for the hits and hits\_total
fields).  Rules classify the individual log lines, capturing data to be
saved in the connections and results tables, and specifying the action to
take for each recognised log line.

Each rule some or all of these fields; all fields are required unless
otherwise stated.

\begin{boldeqlist}

    \item [id] A unique identifier for each rule that other tables can use
        when referring to a specific rule.

    \item [name] A short name for the rule.

    \item [description] Something must have occurred to cause Postfix to
        log each line, this field should describe the event causing the log
        lines this rule recognises.

    \item [restriction\_name] The name of the Postfix restriction that
        caused the mail to be rejected.  This field is valid only for rules
        that recognise rejection log lines, i.e.\ rules that have an action
        of \action{DELIVERY\_REJECTED}.

    \item [program] The Postfix component (e.g.\ \daemon{smtpd}) whose log
        lines the rule applies to; see \sectionref{rule conditions in
        implementation} for full details of how this attribute is used.

    \item [regex] The regex to recognise log lines with, as documented in
        \sectionref{regex components}.

    \item [connection\_data] Sometimes rules need to provide data that is
        not present in the log line: e.g.\ setting \texttt{client\_ip} when
        a mail is being delivered to another server.  Any field in the
        connections table can be set in this way.  The format is:
        \newline{} \tab{} \texttt{ client\_hostname = localhost,}
        \newline{} \tab{} \tab{} \texttt{client\_ip = 127.0.0.1} \newline{}
        i.e.\ semi-colon or comma separated assignment statements.  Commas
        and semi-colons cannot be escaped and thus cannot be included in
        data, this feature is intended for use with small amounts of data,
        so dealing with escape sequences was deemed unnecessary.  This
        field is optional.

    \item [result\_data] The result table equivalent of
        \texttt{connection\_data}, also optional.

    \item [action] The action that will be invoked when this rule
        recognises a log line; a full list of actions and the parameters
        they are invoked with can be found in \sectionref{actions in detail
        in implementation}.

    \item [hits] This counter is maintained for every rule and incremented
        each time the rule successfully recognises a log line.  At the
        start of each run \parsername{} sorts the rules by hits, and at the end of the run it updates
        every rule's hits field in the database.  Assuming that the
        distribution of log lines is reasonably consistent across log
        files, ordering rules by their recognition frequency will reduce
        the parser's execution time.  Rule ordering for efficiency is
        discussed in \sectionref{rule ordering for efficiency}.  This field
        will be set by the parser rather than the rule author.

    \item [hits\_total] The total number of hits for this rule over all
        runs of the parser; hits is reset to zero each time the parser is
        run, but hits\_total is not.  This field will be set by the parser
        rather than the rule author.

    \item [priority] This is the user-configurable companion to hits: when
        the list of rules is sorted, priority overrides hits.  This allows
        more specific rules to take precedence over more general rules, as
        described in \sectionref{rules in architecture}.  This field is
        optional.

    \item [debug] If this field is true, a warning will be issued with
        information about the rule and the log line every time this rule
        recognises a log line.  This field is optional.

\end{boldeqlist}

\subsection{Connections Table}

\label{connections table}

Every accepted mail and every connection that rejected an \acronym{SMTP}
command will have a single entry in the connections table containing all of
the following fields:

\begin{boldeqlist}

    \item [id] This field uniquely identifies the row.

    \item [server\_ip] The \acronym{IP} address of the server: the local
        address when receiving mail, the remote address when sending mail.

    \item [server\_hostname] The hostname of the server, it will be
        \texttt{unknown} if the \acronym{IP} address could not be resolved
        to a hostname via DNS\@.

    \item [client\_ip] The client \acronym{IP} address: the remote address
        when receiving mail, the local address when sending mail.

    \item [client\_hostname] The hostname of the client, it will be
        \texttt{unknown} if the \acronym{IP} address could not be resolved
        to a hostname via DNS\@.

    \item [helo] The hostname used in the HELO command.  The HELO hostname
        occasionally changes during a connection, presumably because spam
        or virus senders think it is a good idea.  By default Postfix only
        logs the HELO hostname when it rejects an \acronym{SMTP} command,
        but it is quite simple to rectify this as described in
        \sectionref{logging helo}.

    \item [queueid] The queueid of the mail if the connection represents an
        accepted mail, or \texttt{NOQUEUE} if not.

    \item [start] The timestamp of the first log line.

    \item [end] The timestamp of the last log line.

\end{boldeqlist}

\subsection{Results Table}

\label{results table}

Every recognised log line will have a row in the results table, associated
with a single connection; a single connection will have at least one result
associated with it, but will usually have several more, and may have
hundreds.

\begin{boldeqlist}

    \item [connection\_id] The id of the row in the connections table this
        result is associated with.

    \item [rule\_id] The id of the rule in the rules table that recognised
        the log line and created this result.

    \item [id] A unique identifier for this result.

    \item [warning] Administrators can configure Postfix to log a warning
        instead of enforcing a restriction that would reject an
        \acronym{SMTP} command --- a mechanism that is quite useful for
        testing new restrictions.  This field will be false for a real
        rejection, or true if the log line was a warning.  This field
        should be ignored if the result is not a rejection, i.e.\ the
        action field of the associated rule is not
        \action{DELIVERY\_REJECTED}.

    \item [smtp\_code] The \acronym{SMTP} code associated with the log
        line.  In general an \acronym{SMTP} code is only present in a log
        line representing a rejection or final delivery; results whose log
        line did not contain an \acronym{SMTP} code will duplicate the
        \acronym{SMTP} code of other results in that connection.  Some
        final delivery log lines do not contain an \acronym{SMTP} code,
        e.g.\ when Postfix delivers to a command; in those cases the
        \acronym{SMTP} code is specified by the rule's
        \texttt{result\_data} field, based on the success or failure
        represented by the log line.

    \item [enhanced\_status\_code] The enhanced status code~\cite{RFC3463}
        is similar to the \acronym{SMTP} code, but is intended to be
        interpreted by mail clients so that error messages can be clearly
        conveyed to the user.  Enhanced status code support was added to
        Postfix in version 2.3; log lines from previous versions will not
        contain any enhanced status codes.

    \item [sender] The sender's email address.  This may change during a
        connection if the client uses different sender addresses for
        multiple rejected delivery attempts; however there should not be
        different sender addresses in the results for one accepted mail.
        Mails sent over the same connection can be distinguished by their
        queueids.

    \item [recipient] The recipient address; there may be multiple
        recipient addresses per mail or connection.

    \item [size] The size of the mail, present for delivered mails only.

    \item [delay] How long the mail was delayed while it was being
        delivered.  This will only be present for delivered mails.

    \item [delays] More detailed information about how long the mail was
        delayed while it was being delivered, again only present for
        delivered mails.

    \item [message\_id] The message-id of the accepted mail, again present
        for delivered mails only.

    \item [data] A field available to store a piece of captured data that
        does not have a specific field, e.g.\ the rejection message from a
        \acronym{DNSBL}\@.

    \item [timestamp] The time at which the log line was logged.

\end{boldeqlist}

\section{Framework}

The role of the framework in this architecture was described in detail in
\sectionref{framework in architecture}; this section is concerned with the
implementation of the framework within \parsername{}.  The framework
manages the parsing process, taking care of the drudgery and boring tasks,
provides services to the actions, and implements some efficiency measures.

Each time the parser is run the framework performs some initialisation
tasks, setting up data structures that will later be used by actions,
including several state tables.  Data about the connections and mails being
processed is held in \texttt{connections} and \texttt{queueids}
respectively.  The other state tables are used when working around
complications in parsing: \texttt{timeout\_queueids} is used when dealing
with connections that time out during the DATA phase (\sectionref{timeouts
during data phase}); \texttt{bounce\_queueids} is part of the solution to
bounce notification mails being delivered before their creation is logged
(\sectionref{Bounce notification mails delivered before their creation is
logged}); and \texttt{postsuper\_deleted\_queueids} caches information
about mails recently deleted by the administrator so that subsequent log
lines processed by the \action{SAVE\_DATA} action can be discarded
(\sectionref{Mail deleted before delivery is attempted}).

When the framework loads the ruleset, it verifies each of the rules,
checking:

\begin{itemize}

    \squeezeitems{}

    \item That the specified action is registered with the framework.

    \item That the regex is valid.  The regex will also be compiled for
        efficiency (\sectionref{Caching compiled regexes}).

    \item For overlap between the data captured by the regex and additional
        data specified in either \texttt{connection\_data} or
        \texttt{result\_data}.

    \item That \texttt{connection\_data}, \texttt{result\_data}, and the
        regex captures specify valid fields to save data to.

\end{itemize}

State tables from a previous \parsername{} run, if any, will be loaded now;
the framework supports saving state at any time, without adversely
affecting the parsing process.  The need to track re-injected mail
(\sectionref{Re-injected mails}) complicates the process of saving state,
as the relationships between mails must be maintained; loading state is
also complicated by dealing with aborted delivery attempts
(\sectionref{aborted delivery attempts}), because a separate set of
relationships between connections and mails must be re-created when the
state tables are restored.

The framework is now ready to begin parsing.  For each log line it will use
those rules whose \texttt{program} field equals the program in the log line
(previously described in \sectionref{rule conditions in implementation}),
falling back to generic rules if necessary, and warning if the log line is
unrecognised.  The rules are sorted for increased efficiency; see
\sectionref{rule ordering for efficiency} for details.  The repetitive
nature of log files gives them high compression ratios; the framework uses
a Perl module named \texttt{IO::Uncompress::AnyUncompress} to read
compressed log files; this saves users the trouble of uncompressing them to
a temporary file before parsing begins.  When used interactively, the
framework also displays a progress bar to show how far parsing has
progressed through the log file and how long the remainder of the parsing
process is likely to take.  The progress bar is not as reliable when
parsing compressed log files, because a compressed block may uncompress to
a variable number of log lines; variation between the recognition and
processing time of individual log lines also affects its accuracy, but
overall the progress bar is a useful addition.  \parsername{} is primarily
intended for parsing complete log files, but with minor modifications it
could be used to parse a live log file, periodically checking it for new
log lines; this could be very useful for programs that work best with up to
date data, e.g.\ a program for monitoring the health of the mail system.

The framework collects data used when evaluating \parsernames{} efficiency
for the evaluation chapter (\sectionref{Evaluation}); some of the
techniques used to improve parsing speed can be turned off or altered to
measure the effect they have.  The data gathered with each optimisation
disabled will be compared to the data gathered with it enabled, to quantify
the benefit each optimisation provides.  Three sets of data are collected:

\begin{enumerate}

    \item How long it takes to parse each log file (\graphref{parsing time
        vs file size vs number of log lines graph}).

    \item The number of rules tried when recognising log lines
        (\graphref{Mean number of rules tried per log line}); the framework
        may need to try multiple rules for a given log line before it finds
        one that recognises it.

    \item The number of log lines and rules tried for each Postfix
        component (\graphref{Mean number of rules tried per log line for
        each Postfix component}).

\end{enumerate}

The framework has five ways that it can adapt its behaviour to demonstrate
how effective each optimisation is:

\begin{enumerate}

    \item The rule ordering used can be changed from optimal (the default,
        most efficient) to shuffled (intended to represent an unordered
        ruleset) or reverse (reverse of optimal, least efficient).  The
        results of this are shown in \sectionref{rule ordering for
        efficiency}.

    \item The framework can record which rule recognised each log line, and
        then on a subsequent run consult that information so that it uses
        the correct rule for each log line.  This gives the best possible
        running time, because only one rule is used for each log line.  A
        variation of this is to use the correct rule last, to get the worst
        possible running time.  \sectionref{perfect rule ordering}
        discusses this in detail.

    \item Each regex is compiled and the result cached when the rules are
        loaded; this can be disabled and the regex compiled every time it
        is used when recognising log lines.  The effect of this
        optimisation is shown in \sectionref{Caching compiled regexes}.

    \item Normally when a log line has been recognised the framework
        invokes the action specified by the rule.  Invoking the action can
        be skipped if desired, so the timing information shows how long is
        spent recognising log lines only; this time can be subtracted from
        the time taken by a normal run to show how long is spent processing
        log lines.  This data is analysed in \sectionref{recognising vs
        processing}.

    \item The framework can skip inserting data into the database, because
        that dominates the execution time of the parser; the evaluation
        chapter measures the speed of \parsername{}, not the speed of the
        database or the disks it resides on.  The effect on parsing time of
        storing results in the database is described at the beginning of
        \sectionref{Evaluation}.

\end{enumerate}

The framework also provides several debugging options, to aid in writing or
correcting rules, or figuring out why the parser is not behaving as
expected.  In increasing order of severity they are:

\begin{enumerate}

    \item Individual rules can set their debug field to true, and debugging
        information will be printed each time they recognise a log line.

    \item Each result can be extended with extra debugging information,
        which is useful when a warning dumps a mail's data for inspection.
        The information that is added is: the log line's timestamp in human
        readable form; the entire log line; the name of the log file and
        the number of the log line within it.  This information is not
        stored in the database.

    \item Every time a log line is recognised, the recognising rule's regex
        and the log line can be printed, so that the user can verify that
        the correct rule recognised each log line.

    \item Every connection and result added to the database can be dumped
        in a human readable form.  This will result in a huge amount of
        debugging information, so it is only useful for small log file
        snippets, because otherwise the amount of information is
        overwhelming.

\end{enumerate}

A user, typically a mail administrator, would use these options when having
difficulty adding or modifying rules to recognise log lines not recognised
by \parsernames{} \numberOFrules{} rules.  Option 1 is useful for debugging
a single rule, to check if it is recognising log lines it should not;
figuring out that a rule is not recognising log lines that it should
recognise is a more difficult task.  Option 2 is useful when \parsername{}
is warning about parsing problems, because all the information about a mail
will be present in the warning message: the full text of every log line
that is part of the mail; the rule that recognised each log line; the
filename of the log file and the line number within it for each log line.
Option 3 is less useful because of the volume of data it produces, though
for small log file snippets, e.g.\ 1000 log lines, it is possible to
manually verify that the correct rule recognised each log line.  Option 4
is not useful for most users: it is for debugging problems with the data
that is saved to the database, and sufficient safeguards are in place that
there should be plenty of warning messages explaining what is wrong without
using this option.

\section{Actions}

\label{actions in implementation}

Actions are the component of this architecture responsible for processing
all of the inputs recognised by rules; in \parsername{} they reconstruct
the journey each mail takes through Postfix, dealing with all the
complications and difficulties that arise.  \parsername{} has
\numberOFactions{} actions and \numberOFrules{} rules, with an uneven
distribution of rules to actions as shown in \graphref{Distribution of
rules per action}.  Unsurprisingly, the action with the most associated
rules is \action{DELIVERY\_REJECTED}, the action that processes Postfix
rejecting a mail delivery attempt; next is \action{SAVE\_DATA}, the action
that saves useful information without doing any other processing.  The
third most common action is, perhaps surprisingly, \action{UNINTERESTING}:
this action does nothing when invoked, allowing uninteresting log lines to
be parsed without any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\action{UNINTERESTING} action recognise log lines that are not associated
with a specific mail, e.g.\ notices about configuration files changing;
these log lines are recognised and processed so that the framework can warn
about unrecognised log lines, informing the user that they need to add to
or extend their ruleset.  Most of the remaining actions have only one or
two associated rules, because they will only ever have one or two log line
variants, e.g.\ all log lines showing that a remote client has connected
are recognised by a single rule and handled by the \action{CONNECT} action.

No correlation exists between how often an action is specified by rules and
how often an action is invoked because a rule recognises a log line.
\Graphref{Number of times each action was invoked excluding 22 and 62--68}
shows the number of times each action was invoked when parsing the
\numberOFlogFILES{} log files used to generate the statistics in
\sectionref{parser efficiency}, \textit{excluding\/} log files 22 and
62--68, because their content is extremely skewed by two mail loops.  The
action most commonly specified by rules, \action{DELIVERY\_REJECTED}, is
the third most commonly invoked action; the most commonly invoked actions,
\action{CONNECT} and \action{DISCONNECT}, are each specified by only one
rule.  The \action{UNINTERESTING} action, the third most commonly specified
action, is halfway down the graph.

As expected, some actions have been invoked almost exactly the same number
of times: almost every \action{CONNECT} will have a \action{DISCONNECT},
with only a 0.00042\% difference between the number of times the two were
invoked; every mail that enters Postfix's queue will be managed by
\daemon{qmgr} (\action{MAIL\_QUEUED}) and processed by \daemon{cleanup}
(\action{CLEANUP\_PROCESSING}), and again the two have only a 0.01648\%
difference between their invocations.  Confusingly,
\action{CLEANUP\_PROCESSING} is invoked more often than
\action{MAIL\_QUEUED}: occasionally, accepted mail in the process of being
transferred is interrupted by a timeout, and there is a log line from
\daemon{cleanup} but not from \daemon{qmgr}, as described in
\sectionref{discarding cleanup log lines}.

\showgraph{build/graph-action-distribution}{Number of rules specifying each
action}{Distribution of rules per action}

\showgraph{build/graph-number-of-action-invocations-excluding-22-and-62--68}{Number
of times each action was invoked when parsing \numberOFlogFILES{} log
files, excluding log files 22 and 62--68}{Number of times each action was
invoked excluding 22 and 62--68}

\subsection{Actions In The \parsernamelong{}}

\label{actions in detail in implementation}

This section explains the actions found in \parsername{}; it may help to
revisit the flow chart in \sectionref{flow chart} to see how a mail passes
from one action to another as its log lines are recognised.  The words
\textit{mail\/} and \textit{connection\/} are used in the actions
descriptions below because they are less unwieldy and more specific than
\textit{state table entry\/}; a connection becomes a mail during the
\action{CLONE} action, which processes Postfix accepting a delivery
attempt, and the data structure moves from the connections state table to
the queueids state table.

The complications and difficulties that arose when parsing real-world log
files are documented in \sectionref{complications}; some action
descriptions refer to specific difficulties they address.  The
complications are documented in a separate section to avoid overwhelming
the action descriptions.

If the log line has enough information to identify the correct connection
or mail, each action will save all the data captured by the recognising
rule's regex.  Each action is passed the same arguments:

\begin{boldeqlist}

    \squeezeitems{}

    \item [rule] The recognising rule.

    \item [data] The data captured from the log line by the rule's regex.

    \item [line] The log line, separated into fields:

        \begin{boldeqlist}

            \squeezeitems{}

            \item [timestamp] The time the line was logged at.

            \item [program] The name of the program that generated the log
                line.

            \item [pid] The \acronym{pid} of the process that logged the
                line.

            \item [host] The hostname of the server the line was logged on.

            \item [text] The remainder of the log line, i.e.\ the message
                logged by the program and recognised by the rule.

        \end{boldeqlist}

\end{boldeqlist}

\begin{description}

    \item [BOUNCE\_CREATED] Postfix 2.3 and subsequent versions log the
        creation of bounce messages.  This action creates a new mail if
        necessary; if the mail already exists the unknown origin flag will
        be removed from it.  The action also marks the mail as a bounce
        notification.  To deal with complication \sectionref{Bounce
        notification mails delivered before their creation is logged} this
        action checks a cache of recent bounce mails, to avoid creating
        bogus bounce mails when log lines are out of order.

    \item [CLEANUP\_PROCESSING] \daemon{cleanup} processes every mail that
        passes through Postfix; details of what it does are available in
        \sectionref{Postfix Daemons}.  This action saves all data captured
        by the rule's regex if the log line has not come after a timeout
        (see \sectionref{discarding cleanup log lines}); it also creates
        the mail if necessary, setting its unknown origin flag (see
        \sectionref{pickup logging after cleanup}).

    \item [CLONE] Multiple mails may be accepted on a single connection, so
        each time a mail is accepted the connection's state table entry
        must be cloned and saved in the state tables under its queueid; if
        the original data structure was used then second and subsequent
        mails would overwrite one another's data.

    \item [COMMIT] Enter the data from the mail into the database.  Entry
        will be postponed if the mail is a child waiting to be tracked
        (\sectionref{Re-injected mails}).  Once entered in the database,
        the mail will be usually be deleted from the state tables, but
        deletion will be postponed if the mail is the parent of re-injected
        mail (\sectionref{Re-injected mails}).

    \item [CONNECT] Handle a remote client connecting: create a new
        connection, indexed by \daemon{smtpd} \acronym{pid}.  If a
        connection already exists it is treated as a symptom of a bug in
        \parsername{}, and the action will issue a warning containing the
        full contents of the existing connection plus the log line that has
        just been parsed.

    \item [DELETE] Deals with mail deleted using Postfix's administrative
        command \daemon{postsuper}.  This action adds a dummy recipient
        address if required (see \sectionref{Mail deleted before delivery
        is attempted}), then invokes the \action{COMMIT} action to save the
        mail to the database.

    \item [DELIVERY\_ERROR] Process log lines indicating that an error
        occurred and the remote client disconnected; the log lines
        processed by this action will be followed by log lines processed by
        the \action{DISCONNECT} action, so all this action does is save
        data from the log line.

    \item [DELIVERY\_REJECTED] Postfix rejected an \acronym{SMTP} command
        from the remote client.  This is the action most frequently
        specified by rules, because so many different restrictions are used
        to reject delivery attempts.  This action is quite simple: if the
        log line contains a queueid, save the data captured by the rule's
        regex to the mail identified by that queueid; otherwise save it to
        the connection identified by the \acronym{pid} in the log line.

    \item [DISCONNECT] Invoked when the remote client disconnects, it
        enters the connection in the database if it has any useful data,
        performs any required cleanup, and deletes the connection from the
        state tables.  This action deals with aborted delivery attempts
        (\sectionref{aborted delivery attempts}).

    \item [EXPIRY] If Postfix has not managed to deliver a mail after
        trying for five days, it will give up and return the mail to the
        sender.  When this happens the mail will not have a combination of
        Postfix programs that passes the valid combinations check,
        implemented to deal with the complication described in
        \sectionref{out of order log lines}; this action tags the mail as
        having expired, so the \action{COMMIT} action will skip the valid
        combinations check.

    \item [MAIL\_BOUNCED] This action behaves in exactly the same way as
        the \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish delivery attempts that bounce from other
        delivery attempts.

    \item [MAIL\_DISCARDED] Sometimes mail is discarded by Postfix, e.g.\
        mail submitted locally that is larger than the limit configured by
        the administrator.  This action is used for those mails; it invokes
        the \action{COMMIT} action, but is a separate action to simplify
        further analysis.

    \item [MAIL\_QUEUED] This action represents Postfix picking a mail from
        the queue to deliver.  This action needs to deal with out of order
        log lines when mail is re-injected for forwarding; see
        \sectionref{Re-injected mails} for details.

    \item [MAIL\_SENT] This action behaves in exactly the same way as the
        \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish successful delivery attempts from other
        delivery attempts.

    \item [MAIL\_TOO\_LARGE] When a client tries to send a mail larger than
        the local server accepts, the mail will be discarded and the client
        informed of the problem.  The mail may have been accepted and
        partially transferred; if so a data structure must be disposed of.
        See \sectionref{timeouts during data phase} for full details;
        although that describes timeouts the processing is the same for
        mails that are too large.

    \item [PICKUP] The \action{PICKUP} action corresponds to the
        \daemon{pickup} service dealing with a locally submitted mail.  A
        new mail will be created, although out of order log lines may have
        caused it to already exist, as documented in \sectionref{pickup
        logging after cleanup}.

    \item [POSTFIX\_RELOAD] When Postfix stops or reloads its
        configuration, it kills all \daemon{smtpd} processes, requiring all
        connections to be cleaned up, entered in the database, and deleted
        from the state tables.  Postfix probably kills all the other
        components too, but \parsername{} is only affected by
        \daemon{smtpd} processes exiting.

    \item [SAVE\_DATA] Every action that can locate the correct entry in
        the state tables saves any data captured by the recognising rule's
        regex to it.  The \action{SAVE\_DATA} saves data in this way but
        does not do anything else; it is invoked for log lines that contain
        useful data but do not require any further processing.

    \item [SMTPD\_DIED] Sometimes a \daemon{smtpd} dies, is killed by a
        signal, or exits unsuccessfully; the associated connection must be
        cleaned up, entered in the database if it has enough data, and
        deleted from the state tables.  Sometimes the connection will not
        have enough data to satisfy the database schema, so it cannot be
        entered into the database for future analysis; unfortunately this
        means that the small amount of data that has been gathered will be
        lost.

    \item [SMTPD\_WATCHDOG] \daemon{smtpd} processes have a watchdog timer
        to deal with unusual situations; after five hours the timer will
        expire and the \daemon{smtpd} will exit.  This occurs very
        infrequently, as there are many other timeouts that should occur in
        the intervening hours, e.g.\ timeouts for DNS requests or timeouts
        reading data from the client.  The active connection for that
        \daemon{smtpd} must be cleaned up, entered in the database, and
        deleted from the state tables.

    \item [TIMEOUT] The connection with the remote client timed out, so the
        mail being transferred must be discarded.  The mail may have been
        accepted: if so there is a data structure to dispose of.  See
        \sectionref{timeouts during data phase} for full details.

    \item [TRACK] Track a mail when it is re-injected for forwarding to
        another mail server; this happens when a local address is aliased
        to a remote address (see \sectionref{tracking re-injected mail} for
        a full explanation).  \action{TRACK} will be called when dealing
        with the parent mail, and will create the child mail if necessary.
        \action{TRACK} checks if the child has already been tracked, either
        with this parent or with another parent, and issues appropriate
        warnings if so.

    \item [UNINTERESTING] This action just returns successfully: it is used
        when a log line needs to be parsed for completeness, but does not
        either provide any useful data to be saved or require anything to
        be done.

\end{description}

\subsection{Adding New Actions}

\label{adding new actions in implementation}

Adding new actions is not as easy as adding new rules, though care has been
taken in the architecture and implementation to make adding new actions as
painless as possible; one of the few limitations is that \parsername{} is
written in Perl, so new actions must also be written in Perl.  The
implementer writes a subroutine that accepts the standard arguments given
to actions, and registers it with the framework by calling the framework's
\texttt{add\_actions()} subroutine.  The new action must be registered
before the rules are loaded, because it is an error for a rule to specify
an unregistered action; this helps catch mistakes made when adding new
rules.  No other work is required from the implementer to integrate the
action into the parser; all of their attention and effort can be focused on
correctly implementing their new action.  The action may need to extend the
list of valid combinations described in \sectionref{out of order log lines}
if the new action creates a different set of acceptable programs, but this
would only be necessary if the new action processes log lines from Postfix
components that \parsername{} does not have rules for, e.g.\
\daemon{virtual} or \daemon{lmtp}.

\section{Rules}

\label{rules in implementation}

The rules are responsible for classifying each log line and specifying the
correct action to be invoked.  In any parser implemented using this
architecture, the rules will be the most visible component and most likely
to be modified by users.  Rules need to be as simple as possible so that
users can easily modify or add to them, but the implementation must balance
that simplicity with the need to provide enough flexibility and power to
successfully parse inputs.

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that description will not be repeated here, but
should be fleshed out by the example rule in \sectionref{example rule in
implementation}.  The process of creating new rules from unparsed log lines
is dealt with in \sectionref{creating new rules in implementation}; the
implementation of the utility supplied with \parsername{} to create regexes
from unparsed log lines is also described in that section.  The rule
conditions used in \parsername{} are explained next, and the final portion
of this section provides some suggestions for how to detect overlapping
rules.

\subsection{Example Rule}

\label{example rule in implementation}

The example rule in \tableref{Example rule in implementation table}
recognises the log line resulting from Postfix rejecting a delivery attempt
because the domain in the sender address does not have an A, AAAA, or MX
DNS entry, i.e.\ mail could not be delivered to any address in the sender's
domain (for full details see~\cite{reject-unknown-sender-domain}).

Example log line recognised by this rule:

% RFC 3330 says that 192.0.2.0/24 is reserved for example use.

\begin{verbatim}
NOQUEUE: reject: RCPT from smtp.example.com[192.0.2.1]:
  550 5.1.8 <foo@example.com>:
  Sender address rejected: Domain not found;
  from=<foo@example.com> to=<info@example.net>
  proto=SMTP helo=<smtp.example.com>
\end{verbatim}

% do not reformat this!
\begin{table}[thbp]
    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

The fields in \tableref{Example rule in implementation table} are used as
follows:

\begin{description}

    \item [name, description, and restriction\_name:] are not used by the
        parser, they serve to document the rule for the user's benefit.

    \item [program and regex:] The program is used to restrict the log
        lines this rule will be used to recognise; see \sectionref{rule
        conditions in implementation} for details.  The regex does the
        actual recognition of the log lines, and data captured by the regex
        (e.g.\ sender, recipient) will be automatically saved to the
        results and connections tables.

    \item [action:] The action to be invoked when the rule recognises a log
        line.  See \sectionref{actions in detail in implementation} for
        details of the actions implemented by \parsername{}, and
        \sectionref{actions in architecture} for the role of actions in the
        parser architecture.

    \item [result\_data and connection\_data:] are used to provide data not
        present in the log line, and are unused in this rule.

    \item [hits, hits\_total, and priority:] hits and priority are used
        when ordering the rules for more efficient parsing (see
        \sectionref{rule ordering for efficiency}).  At the end of each
        parsing run hits is set to the number of log lines recognised by
        the rule.  Hits\_total is the sum of hits over every parsing run,
        but is otherwise unused by the parser.

    \item [debug] controls the debugging information that can be generated
        when this rule recognises a log line.

\end{description}

\subsection{Creating New Rules}

\label{creating new rules in implementation}

The log files produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions that suits their needs, including using different
\acronym{DNSBL} services, policy servers, or custom rejection messages.  To
ease the process of parsing new log lines, the architecture separates rules
from actions: adding new actions can be difficult, but adding new rules to
recognise new log lines is trivial, and occurs much more frequently.

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and regex.  The name and description fields should be
set based on the meaning of the log line, to help others understand which
log lines this rule will recognise; the program will be obvious from the
unrecognised log lines.  The action depends on the what the log line
represents, e.g.\ a delivery rejection, a mail being delivered, some useful
information, or something else; examine the list of actions in
\sectionref{actions in detail in implementation} to determine the correct
one.  The regex needs to be constructed based on the log line, but see
below for a tool to ease the process.

Other attributes may be required to make a rule work correctly:
connection\_data, result\_data, priority, or restriction\_name.  In general
it will only become clear that connection\_data or result\_data are
required when \parsername{} warns about an entry in the connections or
results tables that is missing some required fields, because values for the
missing fields are not present in any of the log lines.  For example, the
rule recognising the \daemon{pickup} component processing a mail sets the
client\_hostname to \texttt{localhost} and the client\_ip to
\texttt{127.0.0.1}, because the mail originates on the local server.  If
the new rule deliberately uses the architecture's overlapping rules feature
it will be obvious that the priority field needs to be set, on this rule
and possibly others; the priority field may be needed on unintentionally
overlapping rules too, but that is more difficult to determine.  Finally
the restriction\_name field should be set when the rule's action is
\action{DELIVERY\_REJECTED}; the name of the restriction should be clear
from the content of the log line.

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: programs generally create log lines by substituting variable words
into a fixed pattern, and analysis of the frequency with which each word
occurs can be used to determine whether the word is variable or part of the
fixed pattern.  This classification can be used to group similar log lines
and generate a regex to match each group of log lines.  The algorithm has
five steps:

\begin{description}

    \item [Pre-process the file.]  The new algorithm leverages the
        knowledge gained when writing rules and developing \parsername{};
        it performs many substitutions on the input log lines, replacing
        commonly occurring variable terms (e.g.\ email addresses,
        \acronym{IP} addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces many variable words with fixed
        words, improving the subsequent classification of words as fixed or
        variable.  The altered log lines are written to a temporary file,
        which subsequent stages will use instead of the original input
        file.

        In the original algorithm the purpose of the pre-processing stage
        was to reduce the memory consumption of the program.  In the first
        pass it generates a hash~\cite{hash-functions}\glsadd{hash} (from a
        small range of values) for each word of each log line, incrementing
        a counter for each hash.  The counters will be used in the next
        pass to filter out words: if the word's hash does not have a high
        frequency, the word itself cannot have a high frequency, so there
        is no need to maintain a counter for it, reducing the number of
        counters required and thus the program's memory consumption.

    \item [Calculate word frequencies.]  The position of words within a log
        line is important: a word occurring in two log lines does not
        indicate similarity unless it occupies the same position in both
        log lines.  If a variable term substituted into a log line contains
        spaces, it will appear to the algorithm as more than one word.
        This will alter the position of subsequent words, so a word
        occurring in different positions in two log lines \textit{may\/}
        indicate similarity, but the algorithm does not attempt to deal
        with this possibility.  The algorithm maintains a counter for each
        \texttt{(word, word's position within the log line)} tuple,
        incrementing it each time that word occurs in that position.

        The original algorithm hashes each word and checks that result's
        counter from the previous pass; if that counter has a high
        frequency a separate counter will be maintained for this
        \texttt{(word, word's position within the log line)} tuple.  This
        reduces the number of counters maintained by the algorithm in this
        step, reducing the memory requirements of the algorithm at the cost
        of increased CPU usage.

        As time goes by the amount of memory typically available to a
        program or algorithm increases, and the need to reduce memory
        requirements correspondingly decreases, so the modified algorithm
        omits the hashing step and maintains counters for all tokens.  Most
        of the infrequently occurring words will have been substituted with
        keywords during the first step, vastly reducing the number of
        tuples to maintain counters for; the original algorithm does not
        have the detailed knowledge leveraged by the new algorithm, because
        it is a generic tool.

    \item [Classify words based on their frequency.]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new regexes, and will be obvious
        when the new rules do not recognise some log lines they are
        expected to.  Variable terms are replaced by \texttt{.+} to match
        one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes are combined, and their counters added together.  If a
        regex's counter is lower than the threshold supplied by the user
        the regex is discarded; this second threshold is independent of the
        threshold used to differentiate between fixed and variable words,
        but once again 1\% of log lines is a good starting point.

    \item [Test the new regexes.]  The final step replaces keywords in the
        new regexes in the same way as \parsername{} does, and compiles
        each regex to check they are valid.  A match will be attempted
        between all of the new regexes and each of the unparsed log lines
        the regexes were built from; the user will be warned if a log line
        is not matched by any regex, or if a log line is matched by more
        than one regex.  The number of log lines each regex matches is
        counted, as is the number of log lines matched by all regexes,
        though a log line is counted once only, even if matched by more
        than one regex.

        This step is not performed by the original algorithm.

\end{description}

The new regexes are displayed for the user to add to the database, either
as new rules or merged into the regexes of existing rules; also displayed
are the number of log lines each regex was expected to match, and the
number it actually matched, to help the user notice problems in the new
regexes.  Discarding regexes will result in some of the log lines not being
matched; when the rules have been augmented with the new regexes, the
original log files should be parsed again, and any remaining unparsed log
lines used as input to this utility.

This utility is not intended to create perfect regexes, but it greatly
reduces the effort required to deal with unrecognised log lines.  The
regexes it generates will be self-contained: a parser that relies on using
cascaded parsing would require a modified algorithm, perhaps by replacing
the pre-processing stage with one that applies existing cascading rules to
each log line, and uses the resulting modified log lines for the remainder
of the algorithm.

\subsection{Rule Conditions}

\label{rule conditions in implementation}

Rule conditions as part of the architecture have already been documented in
\sectionref{rule conditions in architecture}, and they can be very complex
and difficult to evaluate.  In contrast, \parsername{} uses quite simple
rule conditions: each rule has a \texttt{program} attribute that specifies
the Postfix component whose log lines it recognises, and each log line
contains the name of the Postfix component that produced it; when trying to
recognise log lines, the framework will only use rules where the two are
equal.  This avoids needlessly trying rules that will not recognise the log
line, or worse, might recognise it unintentionally.  In addition the
framework supports generic rules, whose program attribute is ``*''; these
will be used if none of the program-specific rules recognise the log line.
If none of the rules are successful the framework will warn the user,
informing them that they need to augment their ruleset, and alerting them
that because \parsername{} failed to recognise some of their log lines, the
results stored in the database may be incomplete.

\subsection{Regex Components}

\label{regex components}

Each rule's regex will have keywords expanded when the rules are loaded,
for several reasons:

\begin{itemize}

    \item It simplifies both reading and writing of regexes and makes each
        regex largely self-documenting.  E.g.\ the meaning of
        \_\_CLIENT\_HOSTNAME\_\_ is immediately clear, whereas
        \verb!(?:unknown|(?:[-.\w]+))! needs to be deciphered each time it
        is encountered.

    \item It avoids needless repetition of complex regex components, and
        allows the components to be corrected or improved in one location.
        E.g.\ \_\_SENDER\_\_ is used in 68 rules; if a mistake is
        discovered it only needs to be corrected in one place.

    \item It enables automatic extraction and saving of captured data.  The
        regex snippets use Perl 5.10's named capture buffers~\cite{perlre}
        to capture data, so the mapping between captures and fields does
        not need to be explicitly specified by the rule.

\end{itemize}

For efficiency the keywords are expanded and every rule's regex is compiled
before attempting to parse the log file, otherwise every regex would be
recompiled each time it was used, resulting in a large, data dependent
slowdown, as described in \sectionref{Caching compiled regexes}.  Most of
the keywords are named after the fields in the connections or results
tables they populate: \_\_CLIENT\_HOSTNAME\_\_, \_\_CLIENT\_IP\_\_,
\_\_DELAY\_\_, \_\_DELAYS\_\_, \_\_ENHANCED\_STATUS\_CODE\_\_,
\_\_HELO\_\_, \_\_MESSAGE\_ID\_\_, \_\_QUEUEID\_\_, \_\_RECIPIENT\_\_,
\_\_SENDER\_\_, \_\_SERVER\_HOSTNAME\_\_, \_\_SERVER\_IP\_\_, \_\_SIZE\_\_,
and \_\_SMTP\_CODE\_\_.

Some of the keywords need more explanation:

\begin{eqlist}

    \squeezeitems{}

    \item [\_\_CHILD\_\_]  The queueid of a child mail; see
        \sectionref{Re-injected mails}.

    \item [\_\_COMMAND\_\_]  All \acronym{SMTP} commands.

    \item [\_\_CONN\_USE\_\_]  How many times the connection was reused;
        Postfix tries to reuse connections whenever possible to reduce the
        load on both the sending and receiving servers.

    \item [\_\_DATA\_\_]  This snippet is special: it matches a zero-length
        string, so it needs to be followed by a pattern that matches
        appropriately, but it captures whatever is matched so that it can
        be saved.  E.g.\ \verb!/(__DATA__connection (?:refused|reset!
        \newline{} \tab{}\verb!by peer))/! matches either ``connection
        refused'' or ``connection reset by peer'' and causes it to be saved
        to the data field of that log line's result.

    \item [\_\_PID\_\_]  The \acronym{pid} of a \daemon{smtpd} process that
        dies or is killed; see the \action{SMTPD\_DIED} action in
        \sectionref{actions in detail in implementation}.

    \item [\_\_RESTRICTION\_START\_\_]  Matches the standard information
        Postfix includes at the start of almost all log lines resulting
        from rejecting a delivery attempt.

    \item [\_\_SHORT\_CMD\_\_]  Postfix sometimes logs \acronym{SMTP}
        commands in a short, single word form; this snippet matches all of
        those, except \texttt{DATA}, which typically has a more specific
        rule.  Priorities could have been used instead of excluding
        \texttt{DATA}.

\end{eqlist}

Some similarity exists between regex components and cascaded parsing: each
regex component resembles a rule that recognises part of a log line and
consumes it, leaving the remainder to be recognised by other components or
cascaded rules.  The major difference between the two is that regex
components are explicitly used by the author of the ruleset, whereas
cascaded parsing would be dynamically applied by the framework.

\subsection{Overlapping Rules}

\label{overlapping rules in implementation}

The advantages and difficulties of overlapping rules have already been
addressed in \sectionref{overlapping rules in architecture} and will not be
repeated here.  \parsername{} does not try to detect overlapping rules;
that responsibility is left to the author of the rules.  A mechanism is
provided for ordering overlapping rules: the priority field in each rule;
negative priorities may be useful for catchall rules.

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

\begin{itemize}

    \item Sort rules by program and regex, then visually inspect the list,
        e.g.\ with \acronym{SQL} similar to:    \newline{}
        \verb!SELECT program, regex!            \newline{}
        \verb!    FROM rules!                   \newline{}
        \verb!    ORDER BY program, regex;!     \newline{}
        The rules are sorted first by program, then by regex, because rules
        with overlapping regexes will not overlap if their programs are
        different.  Note that this does not properly deal with generic
        rules whose program is \texttt{*}; those rules will be used on all
        log lines that have not been recognised by program-specific rules.

    \item Compare the results of parsing using different rule orderings, as
        described in \sectionref{rule ordering for efficiency}.  Parse
        several log files using optimal ordering, then dump a textual
        representation of the rules, connections, and results tables.
        Repeat with shuffled and reversed ordering, starting with a fresh
        database.  If the ruleset does not have overlapping rules the
        tables from each run will be identical; differences indicate
        overlapping rules.  The rules that overlap can be determined by
        examining the differences in the tables: each result contains a
        reference to the rule which created it, which will change if that
        rule overlaps with another.  Unfortunately this method cannot prove
        the absence of overlapping rules; it can detect overlapping rules,
        but only if the log files have log lines that are recognised by
        more than one rule.

\end{itemize}

\input{logparser-complications}

\section{Limitations And Possible Improvements}

\label{limitations and improvements in implementation}

Every piece of software suffers from some limitations, and almost always
has room for improvement.  Below are the limitations and possible
improvements that have been identified in \parsername{}.

\subsection{Limitations}

\label{logging helo}

\begin{enumerate}

    \item Each new Postfix release requires writing new rules or modifying
        existing rules to cope with the new or changed log lines.
        Similarly using a new \acronym{DNSBL}, a new policy server, or new
        administrator-defined rejection messages also requires new rules.

    \item The hostname used in the HELO command is not logged if the
        incoming delivery attempt is successful.  Configuring Postfix to do
        this is relatively simple; create a restriction that is guaranteed
        to warn for every accepted mail, as follows:

        \begin{enumerate}

            \item Create \texttt{/etc/postfix/log\_helo.pcre}
                containing:\newline{}
                \tab{}\texttt{/\^/~~~~WARN~Logging~HELO}

            \item Modify \texttt{smtpd\_data\_restrictions} in
                \texttt{/etc/postfix/main.cf} to contain:\newline{}
                \tab{}\texttt{check\_helo\_access~pcre:/etc/postfix/log\_helo.pcre}

        \end{enumerate}

        Although \texttt{smtpd\_helo\_restrictions} seems like the natural
        place to log the HELO hostname, when it is evaluated for the first
        recipient there will not yet be a queueid allocated for the
        delivery attempt, so the log line cannot be associated with the
        correct mail.  A queueid is guaranteed to have been allocated when
        the DATA command has been reached, and thus the queueid will be
        logged by any restrictions taking effect in
        \texttt{smtpd\_data\_restrictions}, and the log line can be
        associated with the correct mail.  Specifying a HELO-based
        restriction in \texttt{smtpd\_data\_restrictions} does not cause
        any problems; Postfix will perform the check correctly.

        Logging the HELO hostname in this fashion also partially prevents
        the complication described in \sectionref{Mail deleted before
        delivery is attempted} from occurring, but only when the mail has a
        single recipient; when a mail has a single recipient address it
        will be logged, but when a mail has multiple recipients no
        addresses are logged.

    \item \parsername{} does not create separate mails where one or more
        delivery attempts are rejected and subsequently a mail is accepted;
        it will appear in the database as one mail with lots of rejections
        followed by acceptance (this has already been mentioned in
        \sectionref{connection reuse}).  It does not appear to be possible
        to make this distinction given the data Postfix logs, though it
        might be possible to write a policy server to provide additional
        logging.

    \item \parsername{} will not detect that it is parsing the same log file
        twice, resulting in the database containing duplicate entries.

    \item \parsername{} does not distinguish between log files produced by
        different sources when parsing; all results will be saved to the
        same database.  This may be viewed as an advantage, because log
        files from different sources can be combined in the same database,
        or it may be viewed as a limitation because it is impossible to
        distinguish between log files from different sources in the same
        database.  If the results of parsing log files from different
        sources must remain separate, \parsername{} can easily be
        instructed to use a different database.

    \item The solution to complication \sectionref{Mails deleted during
        delivery} involves discarding data.

    \item Further complications may arise when parsing log files, and
        \parsername{} will need to be modified to deal with them.

    \item \parsername{} does not limit the size of the database, which will
        grow without bounds unless the user deletes connections and results
        from it.  This is both a limitation and a benefit: the benefit is
        that data will never be unexpectedly deleted, but the limitation is
        that the user must manage the size of the database.

\end{enumerate}

\subsection{Possible Improvements}

\begin{enumerate}

    \item Investigate and write the policy server referred to in limitation
        3 above.

    \item Improve the solution to complication \sectionref{Mails deleted
        during delivery} so that data is not discarded.

\end{enumerate}

\section{Summary}

This chapter has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the resemblance between a database schema and an
\acronym{API} is described, followed by a detailed description and a
diagram of the database schema.  The framework is documented next,
including the steps it takes during initialisation, the parsing process,
and the conveniences it offers to users.  The performance data collected by
the framework is described, as are the ways in which various optimisations
can be disabled to demonstrate their effect, and the debugging options the
framework provides.  The implementation of actions in \parsername{} is
documented, including how frequently each action is specified by rules, and
brief descriptions of why some actions are more popular than others.  All
the actions that are part of \parsername{} are described in detail,
followed by the process of adding a new action to the parser.  The final
component of the architecture, the rules, is also the most visible
component, and its implementation is examined in detail with: an example
rule and an explanation of every field; a description of how to add new
rules and determine the values that should be used for each field; an
explanation of the algorithm used by the utility that creates new regexes
based on unparsed log lines, and how it differs from the original algorithm
it is based on.  How \parsername{} uses rule conditions and overlapping
rules is discussed, accompanied by a description of the regex snippets the
framework provides to simplify the regexes used in rules.  The many
complications and difficulties encountered while writing \parsername{}, and
the solutions developed to overcome them, are documented in depth; also
documented are how some solutions interact, and which action or actions the
solution is implemented in.  This chapter concludes with a list of the
limitations identified in \parsername{}, followed by some possible
improvements suggested by the author.  The next chapter will evaluate this
implementation, examining both \parsernames{} efficiency and the coverage
it achieves when parsing log files.