### 2009-04-13 doc/logparser-conclusion.tex
Pulled the paper's conclusion into the thesis.

\chapter{Conclusion}

\label{conclusion}

XXX THIS NEEDS TO BE EXTENDED\@.

XXX THERE IS NO MENTION OF RESULTS/EFFICIENCY\@.

XXX PULL CONTENT FROM THE PAPER\@.

XXX INCLUDE A TIGHTLY FOCUSED FUTURE WORK SECTION\@.  PARSE LOG LINES FROM
OTHER PROGRAMS IF A WAY CAN BE FOUND TO ASSOCIATE THEM WITH EXISTING DATA
STRUCTURES\@.

Parsing Postfix log files appears at first sight to be an almost trivial
task, especially if one has previous experience in parsing log files, but
it turns out to be a much more taxing project than initially expected.  The
variety and breadth of log lines produced by Postfix is quite surprising,
because a quick survey of sample log files gives the impression that the
number of distinct log lines is quite small; this impression is because of
the uneven distribution exhibited by log lines produced in normal operation
(see \graphref{rule hits graph} for a vivid illustration of this).

Given the diverse nature of the log lines and the ease with which
administrators can cause new log lines to be logged (\sectionref{postfix
background}), enabling users to easily extend the parser to deal with new
log lines is a design imperative (\sectionref{parser design}).  Despite the
resulting initial increase in complexity the task is quite tractable,
though it does raise efficiency and optimisation questions (answered in
\sectionref{parser efficiency}).  Providing a tool to ease the generation
of regexes from unparsed log lines should greatly help users add rules
to parse previously unparsed log lines (\sectionref{creating new rules in
implementation}).

The architecture not only substantially eases the parsing of new log lines,
it makes adding new actions (\sectionref{adding new actions in
implementation}) a relatively simple task.  The simplicity of adding a new
action frees the implementer from worrying about how their action might
interfere with existing parsing, so they can focus on correctly
implementing their new action.

The division of the parser into rules, actions, and framework is unusual
because rules are separated so completely from the actions and framework;
although parsers are often divided (the combination of \texttt{lex} and
\texttt{yacc}~\cite{lex-and-yacc-book} being a common example), the parts
are usually quite internally interdependent, and will be combined into a
complete parser by the compilation process; in contrast \parsername{} keeps
the rules and actions separate until the parser runs.  The actions and
framework are not as completely separated, because the actions depend on
services provided by the framework; however the actions and framework are
not tightly integrated: it would be possible, with some work, to completely
separate the two.

The flow of control in this parser is quite different from that of most
other parsers.  In most traditional parsers the parser has a current state,
and each state has a fixed set of acceptable next states, with unacceptable
states causing parsing to fail --- e.g.\ when a \textbf{C} parser sees the
keyword \textbf{for} it expects to immediately see a left parenthesis, with
any other input causing parsing to fail (comments will already have been
removed by the preprocessor).  This parser is different: the rule which
recognises the next log line for a mail dictates the action that will be
invoked, which is equivalent to the next state in other parsers; thus the
next log line for a mail dictates the next state for that mail.

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications explained in
\sectionref{complications} only become apparent once the vast majority of
log lines have been parsed successfully.  Adding new rules to deal with
numerous, infrequently occurring log lines is a simple, albeit tiresome,
task, (though alleviated by the tool described in \sectionref{adding new
actions in implementation}), whereas dealing with mails where information
appears to be missing is much more grueling.  Trawling through the log
files, looking for something out of the ordinary, possibly hundreds or even
thousands of log lines away from the last mention of the queueid in
question, is extremely time consuming and error prone.  Sometimes the task
is not to spot the log line that is unusual, but to spot that a log line
that is normally present is missing, i.e.\ to realise that one log line
amongst thousands is absent.  In all cases the evidence must be used to
construct a hypothesis to explain the irregularities, and that hypothesis
must then be tested in the parser; if successful, the parser must be
modified to deal with the irregularities, without adversely affecting the
existing functionality.  The complications described in
\sectionref{complications} were solved in the order they are described in,
and that order closely resembles the frequency with which they occur; the
most frequently occurring complications dominate the warning messages
produced, and so naturally they are the first complications to be dealt
with.

A parser's ability to correctly parse its input is extremely important; the
parser's coverage of its test log files is discussed in \sectionref{parsing
coverage}.  Both its success at parsing individual log lines
(\sectionref{log-lines-covered}) and its correctness in reconstructing each
mail's journey through Postfix (\sectionref{mails-covered}) are described
in detail, including the results of manually verifying the correct parsing
of a subset of the test log files.

Experience implementing \parsername{} shows that full input coverage is
relatively easy to achieve with this architecture, and that with enough
time and effort a full understanding of the input is possible.  Postfix log
files would require substantial time and effort to correctly parse
regardless of the architecture used; this architecture enables an iterative
approach to be used (similar to Stepwise
Refinement~\cite{stepwise-refinement}), as is practiced in many other
software engineering disciplines.

This architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new rule with an appropriate regex and the task is complete.  Parsing a
new category of input is achieved by writing a new action and appropriate
rules; quite often the new action will not need to interact with existing
actions, but when interaction is required the framework provides the
necessary facilities.  The decoupling of rules from actions allows
different sets of rules to be used with the same actions, e.g.\ a parser
might have actions to process versions one and two of a file format; by
choosing the appropriate ruleset the parser will parse version one, or
version two, or both versions.  Decoupling also allows other approaches to
rule management, as discussed in \sectionref{rules in architecture}.  The
architecture makes it possible to apply commonly used programming
techniques (such as object orientation, inheritance, composition,
delegation, roles, modularisation, or closures) when designing and
implementing a parser, simplifying the process of working within a team or
when developing and testing additional functionality.  This architecture is
ideally suited to parsing inputs where the input is not fully understood or
does not follow a fixed grammar: the architecture warns about unparsed
inputs and other errors, but continues parsing as best it can, allowing the
developer of a new parser to decide which deficiencies are most important
and require attention first, rather than being forced to fix the first
error that arises.

The data gathered by the Postfix log parser provides the foundation for the
future of this project: applying machine-learning algorithms to the data to
analyse and optimise the set of anti-spam defences in use, followed by
identifying patterns in the data that could be used to write new filters to
recognise and reject spam rather than accepting it.  The parser provides
the data in a normalised form that is far easier to use as input to new or
existing algorithm implementations than trying to adapt each algorithm to
extract data directly from the log files.  The current focus is on
clustering and decision trees to optimise the order in which rules are
applied; future efforts will involve using data gathered by the parser to
train and test new filters.  This task is similar to analysing a black
box application based on its inputs and outputs, and this approach could be
applied to analyse the behaviour of any system given sufficient log
messages to parse.  An alternate approach to black box optimisation that
uses application profiling in conjunction with the application's error
messages to improve the error messages shown to users is described
in~\cite{black-box-error-reporting}; profiling data may be useful in
supplementing systems that fail to provide adequate log messages.

The Postfix log file parser based on this architecture provides a basis for
systems administrators to monitor the effectiveness of their anti-spam
measures and adapt their defences to combat new techniques used by those
sending spam.  This parser is a fully usable application, built to address
a genuine need, rather than a proof of concept whose sole purpose is to
illustrate a new idea; it deals with the oddities and difficulties that
occur in the real world, rather than a clean, idealised scenario developed
to showcase the best features of a new approach.