### 2009-04-15 doc/logparser-conclusion.tex
Corrections from proofreading.

\chapter{Conclusion}

\label{conclusion}

Parsing Postfix log files appears at first sight to be an almost trivial
task, especially if one has previous experience in parsing log files, but
it turns out to be a much more taxing project than initially expected.  The
variety and breadth of log lines produced by Postfix is quite surprising,
because a quick survey of sample log files gives the impression that the
number of distinct log line variants is quite small; this mistaken
impression is because of the uneven distribution exhibited by log lines
produced in normal operation, vividly illustrated by \graphref{rule hits
graph}.

Given the diverse nature of Postfix log lines, and the ease with which
administrators can cause new log lines to be logged (\sectionref{postfix
background}), enabling users to easily extend the parser to deal with new
log lines is a design imperative (\sectionref{parser design}).  Providing a
tool to ease the generation of regexes from unrecognised log lines should
greatly help users to extend their ruleset to recognise previously
unrecognised log lines (\sectionref{creating new rules in implementation}).
The architecture developed for this project not only makes it trivial to
recognise new inputs, adding new actions requires the minimum effort
possible, allowing the implementer to focus on correctly implementing their
new action.  The separation of the architecture into rules, actions, and
framework (\sectionref{parser design}) is unusual, partly because the three
are separated so completely.  Although parsers are often divided into
separate source code files (the combination of lex \&
yacc~\cite{lex-and-yacc-book} being a common example), the parts are
usually quite internally interdependent, and will be combined into a
complete parser by the compilation process; in contrast \parsername{} keeps
the rules and actions separate until the parser runs.

The flow of control in this architecture is quite different from most other
architectures.  Most traditional parsers have a current state: each state
has a fixed set of acceptable next states, and processing is determined by
the state transition that takes place; unacceptable state transitions cause
parsing to fail.  For example, when a \textbf{C} parser sees the keyword
\textbf{for} it expects to immediately see a left parenthesis, with any
other input causing parsing to fail (comments will already have been
removed by the preprocessor).  This architecture is different: the rule
that recognises the input dictates the action that will be invoked.  Rule
conditions (\sectionref{rule conditions in architecture}) allow a stateful
implementation of this architecture, where the list of rules used to
recognise an input is constrained by the parser's current state, but the
recognising rule still dictates the action that is invoked, and whether
directly or indirectly, the next state.

XXX SHOULD THIS NEXT PARAGRAPH BE IN THE PAST TENSE\@?

The real difficulties arose once the parser was successfully dealing with
90\% of the log lines, because most of the irregularities and complications
explained in \sectionref{complications} only became apparent when the vast
majority of log lines were being parsed successfully.  Adding new rules to
deal with the numerous infrequently occurring log line variants is a simple
if tiresome task, whereas dealing with mails that are missing information
or where Postfix's actions are not being correctly reconstructed is much
more grueling.  Trawling through log files, searching for something out of
the ordinary, possibly hundreds or even thousands of log lines away from
the last occurrence of the queueid for the mail in question, is extremely
time consuming and quite error prone.  Sometimes the task is not to spot
the unusual log line, but to spot that a log line that is normally present
is missing, i.e.\ to realise that one log line amongst thousands is absent.
In all cases the evidence must be used to construct a hypothesis to explain
the irregularities, and that hypothesis has to be tested in \parsername{};
if successful, the parser will be modified to deal with the irregularities,
without adversely affecting the existing parsing.  The complications
described in \sectionref{complications} are described in the order they
were solved in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they were the first
complications to be dealt with.

XXX WRITE ABOUT EFFICIENCY\@.

A parser's ability to correctly parse its input is extremely important; the
parser's coverage of \numberOFlogFILES{} log files is discussed in
\sectionref{parsing coverage}.  Both its success at recognising individual
log lines (\sectionref{log-lines-covered}) and its correctness in
reconstructing each mail's journey through Postfix
(\sectionref{mails-covered}) are described in detail, including the results
of manually verifying the correct parsing of a subset of a log file.
Experience implementing \parsername{} shows that full input coverage is
relatively easy to achieve with this architecture, and that with enough
time and effort a full understanding of the input is possible.  Postfix log
files would require substantial time and effort to correctly parse
regardless of the architecture used; this architecture enables an iterative
approach to be used (similar to Stepwise
Refinement~\cite{stepwise-refinement}), as is practiced in many other
software engineering disciplines.

This architecture's greatest strength is the ease with which
implementations of it can be adapted to deal with new requirements and
inputs.  Parsing a variation of an existing input is a trivial task: simply
modify an existing rule or add a new rule and the task is complete.
Parsing a new category of input is achieved by writing a new action and
appropriate rules; quite often the new action will not need to interact
with existing actions, but when interaction is required the framework
provides the necessary facilities.  The decoupling of rules from actions
allows different sets of rules to be used with one set of actions, e.g.\ a
parser might have actions to process versions one and two of a file format;
by choosing the appropriate ruleset the parser will parse version one, or
version two, or both versions.  Decoupling also allows other approaches to
rule management, as discussed in \sectionref{adding new rules in
architecture}.  The architecture makes it possible to apply commonly used
programming techniques (such as object orientation, inheritance,
composition, delegation, roles, modularisation, or closures) when designing
and implementing a parser, simplifying the process of working within a team
or when developing and testing additional functionality.  This architecture
is ideally suited to parsing inputs that are not fully understood or do not
follow a fixed grammar: the architecture warns about unparsed inputs and
other errors, but continues parsing as best it can, allowing the developer
of a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error that
arises.

The data gathered by \parsername{} provides the foundation for the future
of this project: applying machine-learning algorithms to the data to
analyse and optimise the set of anti-spam defences in use, followed by
identifying patterns in the data that could be used to write new anti-spam
techniques to recognise and reject spam rather than accepting it.  The
parser provides the data in a normalised form that is far easier to use as
input to new or existing implementations of machine-learning algorithms
than trying to adapt each algorithm to extract data directly from the log
files.  New policy servers (\sectionref{policy servers}), written to
implement new anti-spam measures, could be tested or trained by using the
data to simulate mail delivery attempts; this would allow simple, fast,
reproducible testing, without the risk of adversely affecting a production
mail server.  Development of \parsername{} is complete, i.e.\ it correctly
parses Postfix log files, and in future it will require only maintenance;
however, one avenue of future development under consideration is to extend
it to parse non-Postfix log lines, e.g.\ SpamAssassin or Amavisd-new log
lines.  \parsername{} can easily be extended to do this, but it lacks a
method of associating the non-Postfix log lines with the existing data
structures and state tables, so that all of the data for a mail delivery
attempt can be stored together.

The Postfix log file parser based on this architecture provides a basis for
systems administrators to monitor the effectiveness of their anti-spam
measures and adapt their defences to combat new techniques used by those
sending spam.  \parsername{} is a fully usable application, built to
address a genuine need, rather than a proof of concept whose sole purpose
is to illustrate a new idea; it deals with the oddities and difficulties
that occur in the real world, rather than a clean, idealised scenario
developed to showcase the best features of a new approach.