### 2009-04-15 doc/logparser.tex
Corrections from proofreading.

% $Id$
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Provides \FloatBarrier that stops floats going past it.
\usepackage{placeins}
% Relax the parameters Latex uses when placing floats.
\usepackage{flexiblefloats}
% Prevent floats from appearing above the line that includes them
\usepackage{flafter}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Better looking horizontal rules for tables.
\usepackage{booktabs}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

% Aligned list environments.
\usepackage{eqlist}
% This voodoo below gives us an eqlist environment with bold items.
\newcommand{\bolditem}[1]{%
    \bfseries#1%
}
\newenvironment{boldeqlist}
    {\begin{eqlist}[\renewcommand{\makelabel}{\bolditem}]}
    {\end{eqlist}}

% Enable customised lists.
\usepackage{enumitem}
% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

% Extra packages recommended by Springer.
%\usepackage{mathptmx}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{makeidx}
%\usepackage{multicol}

% Embed SVN Id
\usepackage{svn}

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix Log Files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

% Sort numbers where there are multiple citations.  Does not appear to have
% the expected effect (probably clashes with hyperref), though it does
% reduce the space between numbers.
\usepackage{cite}
% Check for unused references, and how the citation key (e.g.\ slct-paper)
% in the margin beside the reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used; neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the default spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[4][thbp]{%
    \begin{figure}[#1]
        \caption{#3}\label{#4}
        \includegraphics{#2}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[4][thbp]{%
    \begin{table}[#1]
        \caption{#3}\label{#4}
        \input{#2}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \toprule{}%
}

\newcommand{\tablebottomline}[0]{%
    \bottomrule{}%
}

\newcommand{\tablemiddleline}[0]{%
    \midrule{}%
}

% A command to format a Postfix daemon's name
\newcommand{\daemon}[1]{%
    \texttt{postfix/#1}%
}

\newcommand{\flowchart}[2]{%
    (action: \action{#1}\@; flowchart:~#2)%
}

\newcommand{\action}[1]{%
    \texttt{#1}%
}

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}

\newcommand{\tab}[0]{%
    \hspace*{1em}%
}

% This is silly, but it keeps chktex happy.
\newcommand{\singlequote}[0]{'}%

% Constant values.
\newcommand{\numberOFlogFILES}[0]{%
    93%
}

\newcommand{\numberOFlogFILESall}[0]{%
    774%
}

\newcommand{\numberOFrules}[0]{%
    184%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.500\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.000\%%
}

\newcommand{\numberOFlogLINES}[0]{%
    60,721,709%
}

\newcommand{\numberOFlogLINEShuman}[0]{%
    60.722 million%
}

\newcommand{\numberOFactions}[0]{%
    23%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    16,836%
}

\newcommand{\numberOFconnectionsINlogFILES}[0]{%
    13,850,793%
}

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\acronym{PLP}}
\newcommand{\parsernames}{\acronym['s]{PLP}}
\newcommand{\parsernamelong}{Postfix Log Parser}
\newcommand{\parsernameshort}{PLP}

\begin{document}

\title{Parsing Postfix Log Files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

\begin{abstract}

    \setcounter{page}{2}
    \phantomsection{}
    \addcontentsline{toc}{chapter}{Abstract}

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding Postfix log files, and subsequently in reconstructing the
    actions taken by Postfix when processing mail delivery attempts.  This
    thesis describes the creation of a parser for Postfix log files,
    documenting the architecture developed for this project and the parser
    that implements it, explaining the difficulties encountered and the
    solutions developed, with reference to an implementation that stores
    data gleaned from the log files in an SQL database.  The gathered data
    can be used to optimise current anti-spam measures, to produce
    statistics showing how effective those measures are, or to provide a
    baseline to test new anti-spam measures against.  The Postfix log file
    parser needs to be very precise and strict when parsing log files, yet
    must allow users to easily adapt or extend it to parse new log lines,
    without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  The newly developed
    architecture is designed to make the process of parsing new inputs as
    simple as possible, enabling users to trivially add new rules (to parse
    variants of known inputs) and relatively easily add new actions (to
    process a previously unknown category of inputs).  The architecture
    scales linearly or better as the number of rules and size of input
    increases, making it suitable for parsing large corpora or months of
    accumulated data.

    \SVN$Id$
    \begin{center}
        \SVNId{}
    \end{center}

    XXX MAIL MADS ABOUT EXAMINERS\@.

\end{abstract}

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\setcounter{page}{3}
\tableofcontents
\listoffigures
\listoftables

\newpage

% WSUIPA fonts.
\input{ipamacs}

\input{logparser-introduction}
% Reset the list of acronyms that have been used, so they will be expanded
% again the first time they are used.  The idea is to have them expanded
% once per chapter; I do not reset them before the appendices, but some may
% not have been used since they were reset, and thus will be expanded
% again.
\glsresetall{}
\input{logparser-background}
\glsresetall{}
\input{logparser-other-parsers-review}
\glsresetall{}
\input{logparser-architecture}
\glsresetall{}
\input{logparser-implementation}
\glsresetall{}
\input{logparser-results}
\glsresetall{}
\input{logparser-conclusion}

\appendix

\bibliographystyle{logparser-bibliography-style}
\bibliography{logparser-bibliography}

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

\end{document}