### 2009-04-17 doc/logparser.tex

% $Id$
\documentclass[a4paper,12pt,draft]{report}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Provides \FloatBarrier that stops floats going past it.
\usepackage{placeins}
% Relax the parameters Latex uses when placing floats.
\usepackage{flexiblefloats}
% Prevent floats from appearing above the line that includes them
\usepackage{flafter}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Better looking horizontal rules for tables.
\usepackage{booktabs}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

% Aligned list environments.
\usepackage{eqlist}
% This voodoo below gives us an eqlist environment with bold items.
\newcommand{\bolditem}[1]{%
    \bfseries#1%
}
\newenvironment{boldeqlist}
    {\begin{eqlist}[\renewcommand{\makelabel}{\bolditem}]}
    {\end{eqlist}}

% Enable customised lists.
\usepackage{enumitem}
% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

% Extra packages recommended by Springer.
%\usepackage{mathptmx}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{makeidx}
%\usepackage{multicol}

% Embed SVN Id
\usepackage{svn}

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix Log Files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

% Sort numbers where there are multiple citations.  Does not appear to have
% the expected effect (probably clashes with hyperref), though it does
% reduce the space between numbers.
\usepackage{cite}
% Check for unused references, and how the citation key (e.g.\ slct-paper)
% in the margin beside the reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used; neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the default spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[4][thbp]{%
    \begin{figure}[#1]
        \caption{#3}\label{#4}
        \includegraphics{#2}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[4][thbp]{%
    \begin{table}[#1]
        \caption{#3}\label{#4}
        \input{#2}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \toprule{}%
}

\newcommand{\tablebottomline}[0]{%
    \bottomrule{}%
}

\newcommand{\tablemiddleline}[0]{%
    \midrule{}%
}

% A command to format a Postfix daemon's name
\newcommand{\daemon}[1]{%
    \texttt{postfix/#1}%
}

\newcommand{\flowchart}[2]{%
    (action: \action{#1}\@; flowchart:~#2)%
}

\newcommand{\action}[1]{%
    \texttt{#1}%
}

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}

\newcommand{\tab}[0]{%
    \hspace*{1em}%
}

% This is silly, but it keeps chktex happy.
\newcommand{\singlequote}[0]{'}%

% Constant values.
\newcommand{\numberOFlogFILES}[0]{%
    93%
}

\newcommand{\numberOFlogFILESall}[0]{%
    774%
}

\newcommand{\numberOFrules}[0]{%
    184%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.500\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.000\%%
}

\newcommand{\numberOFlogLINES}[0]{%
    60,721,709%
}

\newcommand{\numberOFlogLINEShuman}[0]{%
    60.722 million%
}

\newcommand{\numberOFactions}[0]{%
    23%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    16,836%
}

\newcommand{\numberOFconnectionsINlogFILES}[0]{%
    13,850,793%
}

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\acronym{PLP}}
\newcommand{\parsernames}{\acronym['s]{PLP}}
\newcommand{\parsernamelong}{Postfix Log Parser}
\newcommand{\parsernameshort}{PLP}

\newcommand{\specialpage}[1]{%
    \phantomsection{}
    \addcontentsline{toc}{chapter}{#1}%
}
\newcommand{\specialpageheading}[1]{%
    \begin{center}
        \textbf{\Large #1}
    \end{center}
}
\newcommand{\specialpageandheading}[1]{%
    \specialpage{#1}
    \specialpageheading{#1}
}

\begin{document}

\pagestyle{empty}

\specialpage{Title}

\begin{center}

    {\Huge Parsing Postfix Log Files}

    \vfill{}

    {\LARGE John Tobin}

    \vfill{}

    A thesis submitted to the University of Dublin, in fulfilment of the
    requirements for the degree of Master of Science in Computer Science.

    \vfill{}

    April 2009

    \vfill{}

    \vfill{}

    \vfill{}

\end{center}

\newpage{}

% Declaration
\specialpageandheading{Declaration}

\vfill{}

\noindent{}I declare that this thesis, and the work described herein, is
entirely my own work, and has not been submitted as an exercise for a
degree at this or any other university.  This thesis may be borrowed or
copied upon request with the permission of the Librarian, Trinity College,
University of Dublin.

\vfill{}

\begin{flushright}
    \underline{\hspace*{10em}} \\~\\
    John Tobin \\
    \today{}
\end{flushright}

\vfill{}

\vfill{}

\vfill{}

\vfill{}

\newpage{}

% Acknowledgements
\specialpageandheading{Acknowledgements}

\noindent{}I would like to express my appreciation to my supervisor, Dr.\
Carl Vogel, for his advice, assistance, and guidance.

\bigskip{}

\noindent{}I would like to thank my wife, Ariane Tobin, for her patience,
good will, and encouragement.

\begin{abstract}

    \setcounter{page}{4}
    \phantomsection{}
    \addcontentsline{toc}{chapter}{Abstract}

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding Postfix log files, and subsequently in reconstructing the
    actions taken by Postfix when processing mail delivery attempts.  This
    thesis describes the creation of a parser for Postfix log files,
    documenting the architecture developed for this project and the parser
    that implements it, explaining the difficulties encountered and the
    solutions developed, and evaluating the parser on the criteria of
    efficiency and Postfix log file coverage.  The parser stores data
    gleaned from the log files in an SQL database; the gathered data can be
    used to optimise current anti-spam measures, to produce statistics
    showing how effective those measures are, or to provide a baseline to
    test new anti-spam measures against.  The Postfix log file parser needs
    to be very precise and strict when parsing, yet must allow users to
    easily adapt or extend it to parse new log lines, without requiring
    that the user have an in-depth knowledge and understanding of the
    parser's internal workings.  The newly developed architecture is
    designed to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of known inputs) and relatively easily add new actions (to process a
    previously unknown category of inputs).  The parser implemented for
    this project is evaluated on the criteria of efficiency and coverage of
    Postfix log files, demonstrating that the conflicting goals of
    efficiency and accuracy can be balanced, and one need not be sacrificed
    to achieve the other.

    \SVN$Id$

\end{abstract}

% Include the header across the top of each page, after the declaration.
\pagestyle{headings}

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\setcounter{page}{5}
\tableofcontents
\listoffigures
\listoftables

\newpage

% WSUIPA fonts.
\input{ipamacs}

\input{logparser-introduction}
% Reset the list of acronyms that have been used, so they will be expanded
% again the first time they are used.  The idea is to have them expanded
% once per chapter; I do not reset them before the appendices, but some may
% not have been used since they were reset, and thus will be expanded
% again.
\glsresetall{}
\input{logparser-background}
\glsresetall{}
\input{logparser-other-parsers-review}
\glsresetall{}
\input{logparser-architecture}
\glsresetall{}
\input{logparser-implementation}
\glsresetall{}
\input{logparser-results}
\glsresetall{}
\input{logparser-conclusion}

\appendix

\bibliographystyle{logparser-bibliography-style}
\bibliography{logparser-bibliography}

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

\end{document}