### 2009-04-22 doc/logparser-conclusion.tex

\chapter{Conclusion}

\label{conclusion}

Parsing Postfix log files appears at first sight to be an uncomplicated
task, especially if one has previous experience in parsing log files, but
it turns out to be a much more taxing project than initially expected.  The
variety and breadth of log lines produced by Postfix is quite surprising,
because a quick survey of sample log files gives the impression that the
number of distinct log line variants is quite small; this mistaken
impression comes from the uneven distribution exhibited by log lines
produced in normal operation, vividly illustrated in \graphref{rule hits
graph}.  Given the diverse nature of Postfix log lines, and the ease with
which administrators can cause new log lines to be logged
(\sectionref{postfix background}), enabling users to easily extend the
parser to deal with new log lines is a design imperative
(\sectionref{parser design}).  Providing a tool to ease the generation of
regexes from unrecognised log lines (\sectionref{creating new rules in
implementation}) should greatly help users who need to extend their ruleset
to recognise previously unrecognised log lines.

This architecture's greatest strength is the ease with which parsers based
on it can be adapted to deal with new requirements and inputs.  Parsing a
variation of an existing input is a trivial task: simply modify an existing
rule or add a new rule and the task is complete.  Parsing a new category of
input is achieved by writing a new action and a rule for each input
variant; quite often the new action will not need to interact with existing
actions, but when interaction is required the framework provides the
necessary facilities.  The architecture imposes very little red tape when
writing new actions, allowing the implementer to focus their time and
energy on correctly implementing their new action (\sectionref{actions in
architecture}).  The separation of the architecture into rules, actions,
and framework (\sectionref{parser design}) is unusual, partly because the
three are separated so completely.  Although parsers are often divided into
separate source code files (the combination of lex \&
yacc~\cite{lex-and-yacc-book} being a common example), the parts are
usually quite internally interdependent, and will be combined into a
complete parser by the compilation process; in contrast \parsername{} keeps
the rules and actions separate until the parser runs.  This separation
enables the optimisations discussed in \sectionref{parser efficiency}, and
it also allows different approaches to ruleset management, e.g.\ using
machine learning techniques to seamlessly create or alter rules to
recognise new inputs (\sectionref{adding new rules in architecture}).  The
decoupling of rules from actions allows different sets of rules to be used
with one set of actions, e.g.\ a parser might have actions to process
versions one and two of a file format; by choosing the appropriate ruleset
the parser will parse version one, or version two, or both versions.  A
general purpose framework can be written, so that writing a parser just
requires writing actions and rules.  The architecture makes it possible to
apply commonly used programming techniques (such as object orientation,
inheritance, composition, delegation, roles, modularisation, or closures)
when designing and implementing a parser, simplifying the process of
working within a team or developing and testing additional functionality.
This architecture is ideally suited to parsing inputs that are not fully
understood or do not follow a fixed grammar: the architecture warns about
unrecognised inputs and errors encountered by actions, but continues
parsing as best it can, allowing the developer of a new parser to decide
which deficiencies are most important and require attention first, rather
than being forced to fix the first error that arises.

The flow of control in this architecture is quite different from most other
architectures.  Most traditional parsers have a current state: each state
has a fixed set of acceptable next states, processing is determined by the
state transition that takes place, and unacceptable state transitions cause
parsing to fail.  This architecture is different: the rule that recognises
the input dictates the action that will be invoked.  Rule conditions
(\sectionref{rule conditions in architecture}) enable stateful parsing,
where the list of rules used to recognise an input is constrained by the
parser's current state, but the recognising rule still dictates the action
that is invoked and, whether directly or indirectly, the next state.

When writing \parsername{}, the real difficulties arose once the parser was
successfully recognising almost all of the log lines, because most of the
irregularities and complications explained in \sectionref{complications}
started to become apparent then.  Adding new rules to deal with numerous
infrequently occurring log line variants was a simple if tiresome task,
whereas dealing with mails that were missing information or where Postfix's
actions were not being correctly reconstructed was much more grueling.
Trawling through log files was extremely time consuming and quite error
prone, searching for something out of the ordinary that might help diagnose
the problem, and eventually finding it --- sometimes hundreds or even
thousands of log lines away from the last occurrence of the queueid for the
mail in question.  Sometimes the task was not to identify the unusual log
line, but to spot that a log line normally present was missing, i.e.\ to
realise that one log line amongst thousands was absent.  In all cases the
evidence was used to construct a hypothesis to explain the irregularities,
and that hypothesis was tested in \parsername{}; if successful, the parser
was modified to deal with the irregularities, without adversely affecting
existing parsing.  The complications documented in
\sectionref{complications} are presented in the order they were solved in,
and that order closely resembles the frequency in which they occur; the
most frequently occurring complications dominate the warning messages
produced, and so naturally they were the first complications to be dealt
with.

\parsername{} is not merely a proof of concept: it is intended to be used
for parsing real-world log files from production mail servers, and the
resulting data used to improve anti-spam defences.  This means that
efficiency is important: parsing must complete in an reasonable period of
time, so that the results can be used in a timely manner.  \parsernames{}
efficiency is evaluated in \sectionref{parser efficiency}, where
optimisations and the effect they have are explored.

A parser's ability to correctly parse its inputs is extremely important;
\parsernames{} coverage of \numberOFlogFILES{} log files is discussed in
\sectionref{parsing coverage}.  Both its success at recognising individual
log lines  and its correctness in reconstructing each mail's journey
through Postfix are described in detail, including the results of manually
verifying that a randomly selected portion of a log file was correctly
parsed.  Experience implementing \parsername{} shows that full input
coverage is relatively easy to achieve with this architecture, and that
with enough time and effort a full understanding of the input is possible.
Postfix log files would require substantial time and effort to correctly
parse regardless of the architecture used; this architecture enables an
iterative approach to be used~\cite{stepwise-refinement}, as is practiced
in many other software engineering disciplines.

The data gathered by \parsername{} provides the foundation for the future
of this project: using machine-learning algorithms to analyse the data and
optimise the set of anti-spam defences in use, followed by identifying
patterns in the data that could be used to write new anti-spam techniques
to recognise and reject spam rather than accepting it.  The database
provides the data in a normalised form that is far easier to use as input
to new or existing implementations of machine-learning algorithms than
trying to adapt each algorithm to extract data directly from log files.
New policy servers, written to implement new anti-spam measures, can be
tested or trained by using the collected data to simulate mail delivery
attempts; this would allow simple, fast, reproducible testing, without the
risk of adversely affecting a production mail server.  Development of
\parsername{} is finished, i.e.\ it correctly parses Postfix log files, and
in future it will only require maintenance; however, one avenue of future
development under consideration is to extend it to parse non-Postfix log
lines, e.g.\ SpamAssassin or Amavisd-new log lines.  \parsername{} can
easily be extended to do this, but it requires a method of associating the
non-Postfix log lines with the existing data structures and state tables,
so that all of the data for a mail delivery attempt can be stored together.

\parsername{} provides a basis for systems administrators to monitor the
effectiveness of their anti-spam measures and adapt their defences to
combat new techniques used by those sending spam.  \parsername{} is a fully
usable application, built to address a genuine need, rather than a proof of
concept whose sole purpose is to illustrate a new idea; it deals with the
oddities and difficulties that occur in the real world, rather than a
clean, idealised scenario developed to showcase the best features of a new
approach.