### 2009-04-24 doc/logparser-background.tex
Changes made after proofreading the preface and chapters 1-3.

\chapter{Background}

\label{background}

This chapter provides background information helpful in understanding the
remainder of this thesis.  It begins with a discussion of the motivation
underlying the project, followed by a brief introduction to \acronym{SMTP},
and finishes with a longer introduction to Postfix, concentrating on
anti-spam restrictions and policy servers.

\section{Motivation}

\label{motivation}

This work is part of a larger project to optimise a mail server's
Postfix-based anti-spam restrictions, generate statistics and graphs, and
provide a platform on which new restrictions can be developed and evaluated to determine
if they are beneficial in the fight against spam.  The program written for
this project, \parsername{}, parses Postfix log files and populates an
\acronym{SQL}
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

Determining the ten Postfix restrictions that reject the highest number of
delivery attempts is a short example of the analysis possible using data
from the database:

\begin{verbatim}
SELECT name, description, restriction_name, hits_total
    FROM rules
    WHERE action = 'DELIVERY_REJECTED'
    ORDER BY hits_total DESC
    LIMIT 10;
\end{verbatim}

If the database supports sub-selects (where the results of one query are
used as a parameter in another), percentages can be obtained for the top
ten restrictions using the query in \figureref{Sample SQL query}, producing
output similar to \tableref{Sample output from SQL query}.

\begin{figure}[thbp]
    \caption{Sample SQL query showing the percentage of delivery attempts
    rejected by the top ten restrictions}
    \empty{}\label{Sample SQL query}
    \input{build/include-sample-sql-query}
\end{figure}

\begin{table}[thbp]
    \caption{Sample output from an SQL query showing the percentage of
    delivery attempts rejected by the top ten restrictions}
    \empty{}\label{Sample output from SQL query}
    \begin{tabular}[]{lrr}
        \tabletopline{}%
        Restriction & Number of  & Percentage of    \\
                    & rejections & total rejections \\
        \tablemiddleline{}%
        \input{build/include-sample-sql-output}
        \tablebottomline{}%
    \end{tabular}

    All percentages in \tableref{Sample output from SQL query} are
    exclusive, i.e.\ the first restriction's number of rejections does not
    include the second restriction's number of rejections.

\end{table}

Another example is determining which restrictions are not effective: the
query in \figureref{Sample SQL query showing ineffective restrictions}
shows which restrictions rejected fewer than 100 delivery attempts in the
last log file parsed, and the percentage of total rejections in that log
file that each of those restrictions represents.

\begin{figure}[thbp]
    \caption{Sample SQL query showing ineffective restrictions}
    \empty{}\label{Sample SQL query showing ineffective restrictions}
    \begin{verbatim}
    SELECT name, description, restriction_name, hits,
            (hits * 100.0 /
                (SELECT SUM(hits)
                    FROM rules
                    WHERE action = 'DELIVERY_REJECTED'
                )
            ) || '%' AS percentage
        FROM rules
        WHERE action = 'DELIVERY_REJECTED'
            AND hits < 100
        ORDER BY hits ASC;
    \end{verbatim}
\end{figure}

The sample database queries yield summary statistics about the efficiency of
anti-spam techniques.  Analysis of this kind is much easier to perform when
the data source is a database that when trying to directly analyse log
files.

\section{Simple Mail Transfer Protocol (SMTP)}

\SMTPglossaryDescription{} Despite the simplicity of the protocol, many
viruses and spam sending programs fail to implement it properly, so
requiring strict adherence to the protocol specification is beneficial in
protecting against spam and viruses.\footnote{Originally all mail servers
adhered to the principle of \textit{Be liberal in what you accept, and
conservative in what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but
unfortunately that principle was written in a friendlier time.  Given the
deluge of spam that mail servers are subjected to daily, a more appropriate
maxim could be: \textit{Require strict adherence to relevant
\acronyms{RFC}; implement the strongest restrictions you can; relax the
restrictions and adherence only when legitimate mail is impeded.\/}  It is
neither as friendly nor as catchy, but it more accurately reflects the
current circumstances.} A typical \acronym{SMTP} conversation resembles the
following (the lines starting with a three digit number are sent by the
server, all other lines are sent by the client):

\begin{verbatim}
220 smtp.example.com ESMTP
HELO client.example.com
250 smtp.example.com
MAIL FROM: <alice@example.com>
250 2.1.0 Ok
RCPT TO: <bob@example.com>
250 2.1.5 Ok
DATA
354 End data with <CR><LF>.<CR><LF>
Message headers and body sent here.
.
250 2.0.0 Ok: queued as D7AFA38BA
QUIT
221 2.0.0 Bye
\end{verbatim}

An example deviation from the protocol:

\begin{verbatim}
220 smtp.example.com ESMTP
HELO client.example.com
250 smtp.example.com
MAIL FROM: Alice in Distribution alice@example.com
501 5.1.7 Bad sender address syntax
RCPT TO: Bob in Sales/Marketing bob@example.com
503 5.5.1 Error: need MAIL command
DATA
503 5.5.1 Error: need RCPT command
Message headers and body sent here.
502 5.5.2 Error: command not recognized
\end{verbatim}

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, newer programs
tend to be written by competent programmers or send mail using well written
programs (e.g.\ Postfix or Sendmail on Unix hosts, Microsoft Outlook on
Windows hosts).  Traditionally a mail server would have done its best to
deal with deficient clients, with the intention of accepting as much mail
sent to its users as possible, e.g.\ by accepting sender or recipient
addresses that were not enclosed in \texttt{<>}, or by ignoring the absence
of a HELO command.\footnote{The HELO command is the first command sent by
the client in the \acronym{SMTP} conversation; it is required by the
protocol, but its original purpose was to prevent mail loops by detecting a
server trying to deliver mail to itself, so some implementations omitted
it.}  Given the volume of spam sent daily, this liberal approach is no
longer viable.

\section{Postfix}

\label{postfix background}

Postfix is a \acronym{MTA} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to employ those restrictions which they judge
suitable for their server's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
different levels of stricture and permissiveness.  Administrators can
supply their own rejection messages to make it clear to senders why their
mail was rejected.  Policy servers (soon to be described in
\sectionref{policy servers}) provide a simple way to write new restrictions
without having to modify Postfix's source code.  Unfortunately this
flexibility has a cost: the complexity of Postfix's log files.  Although it
is usually relatively simple to use standard Unix text processing utilities
to determine the fate of an individual mail, sometimes it can be quite
difficult.  For most mails the journey through Postfix is simple and brief,
but the remainder can be quite complex (see \sectionref{complications} for
details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple components that each perform one of the tasks
required of an \acronym{MTA}, e.g.\ receive mail, send mail, deliver mail
to a user's mailbox;\ full details can be found in~\cite{postfix-overview}.
Postfix's design is strongly influenced by security concerns: those
components that interact with other hosts are not
privileged,\footnote{Privilege means the power to perform actions that are
limited to the administrator, and are not available to ordinary users.} so
bugs in those components will not give an attacker extra privileges; those
components that are privileged do not interact with other hosts, making it
much more difficult for an attacker to exploit any bugs that may exist in
those components.

\subsection{Mixing And Matching Postfix Restrictions}

\label{Mixing and matching Postfix restrictions}

Postfix restrictions are documented fully in~\cite{smtpd_per_user_control,
policy-servers, smtpd_access_readme}; the following is only a brief
introduction.

Postfix uses one restriction list (containing zero or more restrictions)
for each stage of the \acronym{SMTP} conversation: client connection, HELO
command, MAIL FROM command, RCPT TO commands, DATA command, and end of
data.  The appropriate restriction list is evaluated for each stage, though
by default the restriction lists for client connection, HELO, and MAIL FROM
commands will not be evaluated until the first RCPT TO command is received,
because some clients do not deal properly with rejections before this
stage; a benefit of this delay is that Postfix has more information
available when logging rejections.

Postfix uses simple lookup tables to make decisions when evaluating some
restrictions, e.g.\newline{}
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}

\begin{boldeqlist}

    \item [check\_client\_access] The name of the restriction to evaluate.

    \item [cidr] The type of the lookup table.

    \item [/etc/postfix/client\_access] The file containing the lookup
        table.

\end{boldeqlist}

The restriction \texttt{check\_client\_access} checks if the IP address of
the connected client is found in the specified table and returns the
associated result if found; the method of searching the file is dependant
on the type of the lookup table~\cite{postfix-lookup-tables}.  Other
restrictions determine their result by consulting external sources,
e.g.\newline{} \tab{}\texttt{reject\_rbl\_client
dnsbl.example.com}\newline{} checks the \acronym{DNSBL}
\texttt{dnsbl.example.com} and rejects the command if the client's IP
address is listed.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{access-man-page}.  The meaning of \textit{permit\/} and
\textit{reject\/} is obvious; \textit{dunno\/} means to stop evaluating the
current restriction and continue processing the remainder of the
restriction list, allowing exceptions to more general rules.  The
configuration example below shows how to require that all machines on the
local network except for the web server authenticate before sending mail;
the web server is exempt because the legacy applications running on it lack
authentication support.  The restriction list is evaluated from top to
bottom: \texttt{permit\_sasl\_authenticated} permits authenticated clients;
an exception is made for the web server in the
\texttt{check\_client\_access} check, and then other local machines are
rejected, because to continue past \texttt{permit\_sasl\_authenticated} the
client must not have authenticated.

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    <other restrictions>
    permit_sasl_authenticated,
    check_client_access /etc/postfix/allow_webserver.cidr,
    <other restrictions>

  /etc/postfix/allow_webserver.cidr:
    192.0.2.80/32   dunno
    192.0.2.0/24    reject "Please authenticate to send mail"
\end{verbatim}

That example also shows how to supply a custom rejection message.  When the
result of a lookup is the name of another restriction Postfix will evaluate
that restriction; this allows restrictions to be chosen based on the client
IP address, client hostname, HELO hostname, sender address, or recipient
address.  For example, the administrator may require that all clients on
the local network have valid DNS entries, to prevent people sending mail
from unknown machines; one example of how to achieve this is:

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    <other restrictions>
    check_client_access /etc/postfix/require_dns_entries.cidr,
    <other restrictions>

  /etc/postfix/require_dns_entries.cidr:
    192.0.2.0/24    reject_unknown_client_hostname
\end{verbatim}

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex user-defined sequences
of lookups, restrictions, and exceptions.  Postfix tries to protect the
administrator from misconfiguration in as far as it reasonably can, e.g.\
the restriction \texttt{check\_helo\_mx\_access} cannot cause a mail to be
accepted, because the parameter it checks (the hostname given in the HELO
command) is under the control of the remote client.  Despite this, it is
possible for the administrator to make catastrophic mistakes, e.g.\
rejecting all mail --- the administrator must be cognisant of the effects
their configuration changes will have.  This is similar to one of UNIX's
design philosophies: \textit{``UNIX was not designed to stop its users from
doing stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

\subsection{Policy Servers}

\label{policy servers}

A policy server~\cite{policy-servers} is an external program consulted by
Postfix to determine the fate of an \acronym{SMTP} command.  The policy
server is given state information by Postfix (sample state information is
shown in \tableref{Example attributes sent to policy servers}) and returns
a result (reject, permit, dunno, or a rejection name) as described in
\sectionref{Mixing and matching Postfix restrictions}.  A policy server can
perform more complex checks than those provided by Postfix, such as
allowing addresses associated with the payroll system to send mail on the
third Tuesday after pay day only, to help prevent problems from phishing
mails using faked sender addresses.  For example, a phishing mail might
pretend that the payroll system had a disastrous disk failure, and until
the server is replaced all salary payments will have to be processed
manually, so please reply to this mail with your name, address, and bank
account details; the criminal can then use any details sent to him to help
with identity theft.

\begin{table}[thbp]
    \caption{Example attributes sent to policy servers, taken
    from~\cite{policy-servers}}
    \empty{}\label{Example attributes sent to policy servers}
    \centering{}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Attribute name          & Attribute value           \\
        \tablemiddleline{}%
        request                 & smtpd\_access\_policy     \\
        protocol\_state         & RCPT                      \\
        protocol\_name          & SMTP                      \\
        helo\_name              & some.domain.tld           \\
        queue\_id               & 8045F2AB23                \\
        sender                  & foo@bar.tld               \\
        recipient               & bar@foo.tld               \\
        recipient\_count        & 0                         \\
        client\_address         & 1.2.3.4                   \\
        client\_name            & another.domain.tld        \\
        reverse\_client\_name   & another.domain.tld        \\
        instance                & 123.456.7                 \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

Some widely deployed policy servers:

\begin{itemize}

    \item Checking if the client satisfies a domain's \acronym{SPF}
        records, \urlLastChecked{http://www.openspf.org/}{2009/04/21}.
        \acronym{SPF} records specify which mail servers are allowed to
        send mail using sender addresses from a particular domain.  The
        intention is to reduce spam from faked sender addresses,
        backscatter, and joe~jobs~\glsadd{Joe-job}.  There has been
        considerable resistance to \acronym{SPF} because it breaks or
        vastly complicates some commonly-used features of \acronym{SMTP},
        e.g.\ forwarding mail from one domain to another when a user moves.

    \item Greylisting,
        \urlLastChecked{http://www.greylisting.org/}{2009/04/21}, is a
        technique that temporarily rejects a delivery attempt when the
        tuple of \newline{} \tab{}\texttt{(sender address, recipient
        address, remote IP address)}\newline{} has not been seen before; on
        second and subsequent delivery attempts from that tuple the mail
        will be accepted.  This blocks spam from some senders because
        maintaining a list of failed addresses and retrying after a
        temporary failure is uneconomical for a spam sender, but a
        legitimate mail server must retry deliveries that fail temporarily.
        Sadly spam senders are using increasingly complex and well written
        programs to distribute spam, frequently using an ISP provided
        \acronym{SMTP} server from a compromised machine on the ISP's
        network.  Greylisting will slowly become less effective as spam
        senders adapt, but it does block a large percentage of spam mail at
        the moment; the most effective restrictions from the
        \numberOFlogFILES{} log files used when generating the results in
        \sectionref{Evaluation} are shown in \tableref{Sample output from
        SQL query}.  That table shows that greylisting is worth using at
        the moment, particularly when you take into account its position as
        the final restriction that a mail must overcome in the
        configuration used on the mail server that generated the log files:
        on that server greylisting only takes effect for mails that have
        passed all other restrictions.  Some problems may be encountered
        when using greylisting: some servers fail to retry after a
        temporary failure, or legitimate mail may be delayed, particularly
        when coming from a pool of servers.

    \item Scoring systems such as postfwd,
        \urlLastChecked{http://postfwd.org/}{2009/04/21}, perform tests on
        features of the delivery attempt (e.g.\ IP address, sender
        address), incrementing or decrementing a score based on the
        results; if the final score is higher than a threshold the delivery
        attempt is rejected.  The administrator must manually whitelist
        clients if they are to bypass a Postfix restriction, whereas using
        a threshold that requires a delivery attempt to hit several scored
        restrictions will allow delivery attempt that would be rejected by
        a boolean restriction.

\end{itemize}

\section{Summary}

This chapter has provided background information useful in understanding
this thesis, starting with the motivation behind the project, continuing
with an introduction to \acronym{SMTP}, and finishing with Postfix, its
anti-spam restrictions, and its support for policy servers.