### 2009-04-24 doc/logparser-introduction.tex

\chapter{Introduction}

\label{introduction}

The architecture and implementation described in this thesis were developed
as the foundation of a larger project to improve anti-spam defences, by
analysing the performance of the set of anti-spam techniques currently in
use, optimising the order and membership of the set based on that analysis,
and developing supplemental anti-spam techniques where deficiencies are
identified.  Most anti-spam techniques are content-based
(e.g.~\cite{a-plan-for-spam, relaxed-online-svms, word-stemming}) and
require a mail to be accepted before determining if it is spam, but
rejecting mail during the delivery attempt is preferable: senders of
non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; resource usage is reduced on the accepting mail
server, allowing more intensive content-based techniques to be used on the
remaining mail that is accepted; users have less spam mail to wade through.
Improving the performance of anti-spam techniques that are applied when
mail is being transferred via \acronym{SMTP} is the goal of this project,
by providing a platform for reasoning about the performance of anti-spam
techniques.

The approach chosen to measure performance is to analyse the log files
produced by the \acronym{MTA} in use, Postfix, rather than modifying its
source code to generate statistics: this approach improves the chances of
other Postfix users testing, using, and benefiting from the software
developed for this project.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by Postfix: the
parser must be designed and implemented so that adding support for parsing
new inputs is a simple task, because the log lines to be parsed will change
over time.  This variety in log lines occurs for several reasons:

\begin{itemize}

    \squeezeitems{}

    \item Log lines differ amongst versions of Postfix.

    \item The mail server administrator can define custom rejection
        messages.

    \item Policy servers (see \sectionref{policy servers}) may log
        different messages depending on the characteristics of the
        connection.

    \item Every \acronym{DNSBL}\footnote{This thesis is supplied with a
        glossary (\textsection\ref{Glossary}) and a list of acronyms
        (\textsection\ref{Acronyms}).} returns a different explanatory
        message.

\end{itemize}

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam techniques are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the first real foray the author, a Systems
Administrator for a network of over 2000 computers and over 1800 users,
took into processing Postfix log files.}  Very few will have performed
in-depth parsing and analysis of their log files, where the parsing must
correlate each connection or mail's log lines rather than
processing each log line in isolation.  One of the barriers to this kind of
processing is the unstructured nature of Postfix log files, where each log
line was added on an ad hoc basis when a requirement was discovered or new
functionality was added.\footnote{A history of all changes made to Postfix
is distributed with the source code, available from
\urlLastChecked{http://www.postfix.org/}{2009/02/23}.}  Further
complication arises because the set of log lines is not fixed, and log
lines can differ in many ways between servers, even within the same
organisation, where servers may be configured differently or running
different versions of Postfix.

The only prior published work on the subject of parsing Postfix log files
that the author is aware of is \textit{Log Mail Analyser: Architecture and
Practical Utilizations\/}~\cite{log-mail-analyser}, which aims to extract
data from Postfix log files, correlate it, and present it in a form
suitable for a systems administrator to search using the myriad of standard
Unix text processing utilities already available; it is reviewed alongside
the other parsers in the State of the Art Review in chapter~\ref{state of
the art review}.  It was hoped to reuse an existing parser rather than
writing one from scratch, but the effort required to adapt and improve an
existing parser was judged to be greater than the effort to write a new
one, as described in the State of the Art Review.

Once it was decided that a new parser would be written, an architecture was
required to base the implementation on.  Existing architectures are
tailored towards parsing inputs with a fixed grammar or a tightly
constrained format, whereas Postfix log files lack any form of constraint,
as outlined earlier.  A new architecture was designed and developed for
this parser, with the hope that it will be useful to others.  The resulting
architecture is conceptually simple: provide a few generic functions
(\textit{actions\/}), each capable of dealing with an entire category of
inputs (e.g.\ rejecting a mail delivery attempt), accompanied by a
multitude of precise patterns (\textit{rules\/}), each of which recognises
one input variant and only that variant (e.g.\ rejection by a specific
\acronym{DNSBL}), and specifies which action will process the inputs it
recognises.  This architecture is ideally suited to parsing inputs that are
not fully understood or do not conform to a fixed grammar: the architecture
warns about unrecognised inputs and other errors, but continues parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require immediate attention, rather
than being forced to fix the first error that arises.
The architecture is designed to enable the users of a parser to
easily extend it to parse their particular inputs, without requiring much
work or a high level of understanding of the parsing process and the
parser's internal workings.

This architecture is the basis of \parsername{}, a program that parses
Postfix log files and places the resulting data into a \acronym{SQL}
database for later analysis.  The gathered data can be used to optimise
current anti-spam defences, to produce statistics showing how effective
each technique in use is, or to provide a baseline to test new anti-spam
measures against.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times or
providing a dedicated service for them; detecting virus outbreaks that
propagate via mail; billing customers on a shared server.  Preserving the
raw data enables a multitude of uses far beyond those conceived of by the
author.

\section{Thesis Layout}

Chapter~\ref{background} provides background information useful in
understanding \acronym{SMTP}, Postfix, and the motivation behind the
project.

Chapter~\ref{state of the art review} reviews the previously published
research in this area and other Postfix log file parsers, discussing why
they were deemed unsuitable for the task, including why they could not be
improved or expanded upon.

Chapter~\ref{parser architecture} describes the parser architecture
developed for this project, beginning with an overview, then describing
each of the components of the architecture in detail.  This chapter
concentrates on the abstract, theoretical, implementation-independent
aspects of the architecture; discussion of the practical aspects is
deferred until chapter~\ref{Postfix Parser Implementation}.

Chapter~\ref{Postfix Parser Implementation} documents \parsername{}, the
parser based on the architecture described in chapter~\ref{parser
architecture}.  The practical difficulties of implementing each of the
components of the architecture are described, accompanied by the many
complications encountered when parsing Postfix log files, and other details
of the implementation.

Chapter~\ref{Evaluation} evaluates \parsernames{} efficiency, exploring
the various optimisations implemented in the parser and the effect they
have.  It also discusses the coverage achieved by \parsername{} over
\numberOFlogFILES{} log files, with separate sections for the number of log
lines correctly recognised, and the number of connections and mails whose
journey through Postfix was correctly reconstructed.

Chapter~\ref{conclusion} contains the conclusion of the thesis.

The bibliography (appendix~\ref{bibliography}) contains references to the
resources used in designing the architecture and writing \parsername{}.

Appendix~\ref{Glossary} provides a glossary of terms used in the thesis.

Appendix~\ref{Acronyms} contains a list of acronyms used in the thesis;
uncommon acronyms will have an entry in the glossary too.

Appendix~\ref{Postfix Daemons} provides a brief description of each Postfix
daemon.

\section{Previously Published Work}

Portions of chapters~\ref{introduction} and~\ref{parser
architecture}--\ref{Evaluation} have previously been published at an
international conference~\cite{sgai-2008}, and later reprinted in a
journal~\cite{elsevier-2009}.  Publication of the conference paper was
supported by Science Foundation Ireland RFP~05/RF/CMS002.