### 2009-04-26 doc/logparser-architecture.tex
Changes based on proofreading chapters 4 and 5.

\chapter{Parser Architecture}

\label{parser architecture}

To avoid cluttering the description of the parser architecture with the
details of implementing a parser for Postfix log files, each topic has been
given its own chapter.  This chapter presents the architecture developed
for this project, beginning with the overall architecture and design,
followed by detailed documentation of the three components of the
architecture: Framework, Actions, and Rules.  This chapter is focused on
the theoretical, implementation-independent aspects of the architecture;
the practical difficulties of writing a parser for Postfix log files are
covered in detail in \sectionref{Postfix Parser Implementation}.

\section{Architecture Overview}

\label{parser design}

It should be clear from the earlier Postfix background (\sectionref{postfix
background}) that Postfix log files may vary widely from host to host.
With this in mind, one of the architecture's design aims was to make
parsing new inputs as effortless as possible, to enable administrators to
properly parse their own log files.  The solution developed is to divide
the architecture into three parts: Framework, Actions, and Rules.  Each
will be documented separately, but first an overview:

\begin{boldeqlist}

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It manages the parsing process, providing shared
        data storage, loading and validation of rules, storage of results,
        and other support functions.

    \item [Actions] Each action performs the processing required for a
        single \textit{category\/} of inputs, e.g.\ rejection of a delivery
        attempt.  Actions are invoked to process an input once it has been
        recognised by a rule.

    \item [Rules]  Rules are responsible for classifying inputs: each rule
        recognises one input \textit{variant\/} --- a single input category
        may have many input variants.  Each rule also specifies the action
        to be invoked when an input has been recognised; rules thus provide
        an extensible method of associating inputs with actions.

\end{boldeqlist}

For each input, the framework tries each rule in turn until it finds a rule
that recognises the input, then invokes the action specified by that rule.
If the input is not recognised by any of the rules, the framework issues a
warning; the framework will usually continue parsing after this, although
some parsers might prefer to stop immediately.

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework manages the parsing process and
provides services to the actions; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately and correctly processing inputs and the information provided by
rules; rules are responsible for recognising inputs, and extracting data
from those inputs for processing by actions.

Separating the rules from the actions and framework makes it possible to
parse new inputs without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to recognise inputs is
trivial in comparison to understanding an entire parser, identifying the
correct location to change, and making the appropriate changes.  Changes to
a parser must be made without adversely affecting existing parsing,
including any edge cases that are not immediately obvious; an edge case
that occurs only four times in \numberOFlogFILES{} log files is described
in Yet More Aborted Delivery Attempts (\sectionref{yet more aborted
delivery attempts}).  The more intrusive the changes are, the more likely
they are to introduce a bug, so reducing the extent of the changes is
important.  Requiring changes to a parser's source code also complicates
upgrades of the parser, because the changes must be preserved during the
upgrade, and they may clash with changes made by the developer.  This
architecture allows the user to add new rules to a parser without having to
edit it, unless the new inputs cannot be processed by the existing actions.
If the new inputs do require new functionality, new actions can be added to
the parser without having to modify existing actions; only when the new
actions need to cooperate with existing actions will more extensive changes
be required.

Some similarity exists between this architecture and William A.\ Wood's
\acronym{ATN}~\cite{nlpip,atns}, used in Computational Linguistics to
create grammars that parse or generate sentences.  The resemblance between
the two (shown in \tableref{Similarities between ATN and this
architecture}) is accidental, but clearly the two different approaches
share a similar division of responsibilities, despite having different
semantics.

% Do Not Reformat!

\begin{table}[thbp]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \centering{}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Framework     & Determines the sequence
                                          of transitions              \\
                        &               & or actions that
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assemble data and
                                          impose conditions the       \\
                        &               & input must meet to be
                                          accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for
                                          recognising inputs.         \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

The architecture can be thought of as implementing transduction: it takes
data in one form and transforms it to another form; \parsername{}
transforms log files to a \acronym{SQL} database.

Unlike traditional parsers such as those used when compiling a programming
language, this architecture does not require a fixed grammar specification
that inputs must adhere to.  The architecture is capable of dealing with
interleaved inputs, out of order inputs, and ambiguous inputs where
heuristics must be applied --- all have arisen and been successfully
accommodated in \parsername{}.  This architecture is ideally suited to
parsing inputs where the input is not fully understood or does not conform
to a fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues parsing as best it can, allowing the developer of a
new parser to decide which deficiencies are most important and the order to
address them in, rather than being forced to fix the first error that
arises.

\section{Framework}

\label{framework in architecture}

The framework manages the parsing process and provides support functions
for the actions, freeing the programmers writing actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other, and allows new rules to be written
without needing changes to the source code of a parser.  The framework is
the core of the architecture and is deliberately quite simple: the rules
deal with the variation in inputs, and the actions deal with the
intricacies and complications encountered during parsing.  Finding the rule
that recognises the input is a very simple process, as shown by the
pseudo-code in \figureref{Pseudo-code describing how inputs are
recognised}.  The framework tries each rule until it finds one that
recognises the input, then it invokes the action specified by the rule.
The framework issues a warning if the input is not recognised by any of the
rules.

\begin{figure}[thbp]
    \caption{Pseudo-code showing how the framework finds the rule that
    recognises the input}
    \empty{}\label{Pseudo-code describing how inputs are recognised}
\begin{verbatim}
INPUT:
for each input {
    for each rule defined by the user {
        if this rule recognises the input {
            invoke the action specified by the rule
            next INPUT
        }
    }
    warn the user that the input was not recognised
}
\end{verbatim}
\end{figure}

Most parsers will require the same basic functionality from the framework;
it is responsible for managing the parsing process from start to finish,
which will generally involve the following:

\begin{description}

    \item [Register actions]  Each action needs to be registered with the
        framework so that the framework knows about it: the list of
        registered actions will be used when validating rules.

    \item [Load and validate rules]  The framework loads the rules from
        wherever they are stored: a simple file, a database, or possibly
        even a web server or other network service --- though that would
        have serious security implications.  It validates each rule to
        catch problems as early in the parsing process as possible; the
        checks will be implementation-specific to some extent, but will
        generally include the following:

        \begin{itemize}

            \squeezeitems{}

            \item Ensuring the action specified by the rule has been
                registered with the framework.

            \item Checking for conflicts in the data to be extracted, e.g.\
                setting the same variable twice.

            \item Checking that the regex in the rule is valid.

        \end{itemize}

        Some optimisation steps may also be performed during loading of
        rules, as described in \sectionref{parser efficiency}.

    \item [Convert physical inputs to logical inputs] Each rule recognises
        a single input at a time: there is no facility for rules to consume
        more input or push unused input back onto the input stream,
        although actions may use cascaded parsing (explained in
        \sectionref{actions in architecture}) to push input back onto the
        input stream.  A physical input (e.g.\ a single line read from the
        input stream) may contain multiple or partial logical inputs, and
        the framework must pre-process these to provide a logical input for
        the rules to recognise.  This simplifies the rules and actions
        considerably, at the cost of added complexity in the framework;
        during the design phase it was decided that it was easier to deal
        with the problem of parsing multiple or partial inputs once, rather
        than dealing with it in every rule and action.  This is trivial for
        Postfix log files because they have a one-to-one mapping between
        physical and logical inputs; mapping between physical and logical
        inputs may be more difficult for other types of input.  Some input
        types may require pre-processing equivalent to parsing the physical
        inputs; in such cases the framework should take the approach
        adopted by many other parsers: combine the physical inputs into one
        complete input, use the rules to recognise the start of the input,
        discard the recognised portion if successful, and repeat until the
        input has been exhausted.

    \item [Classify the input]  The pseudo-code in \figureref{Pseudo-code
        describing how inputs are recognised} shows how rules are
        successively tried until one is found that recognises the input.
        That pseudo-code is very simple: there may be efficiency concerns
        (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or rule priorities
        (\sectionref{rules in architecture}) that complicate the process.

    \item [Invoke actions]  Once a rule has been found that recognises the
        input, the specified action will be invoked.  The framework
        marshals the data extracted by the rule, invokes the action, and
        pushes the modified input onto the input stream if the action uses
        cascaded parsing (see \sectionref{actions in architecture}).

    \item [Shared storage]  Parsers commonly need to save some state
        information about the input being parsed, e.g.\ a compiler tracking
        which variables are in lexical scope as it moves from one lexical
        block to another.  The framework provides shared storage to deal
        with this and any other storage needs the actions may have.
        Actions may need to exchange data to correctly parse the input,
        e.g.\ setting or clearing flags, maintaining a list of previously
        used identifiers, or ensuring at a higher level that the input
        being parsed meets certain requirements.

    \item [Save and load state]  The architecture can save the contents of
        the shared storage it provides for actions, and reload it later so
        that information is not lost between parsing runs.  \parsername{}
        does this because mails may take some time to deliver
        and thus have their log lines split between log files; a compiler
        might store data structures it builds as it parses different files.

    \item [Specialised support functions]  Actions may need support or
        utility functions; the framework may be a good location for support
        functions, but if another way exists to make those functions
        available to all actions it may be preferable to use that way
        instead, maintaining a clear separation of concerns.

\end{description}

\section{Actions}

\label{actions in architecture}

Each action is a separate procedure written to process a particular
category of input, e.g.\ rejection of a delivery attempt.  One input
category may have many input variants; in general each action will handle
one input category, with each rule recognising one input variant.  It is
anticipated that parsers based on this architecture will have many actions,
and each action may be invoked by many rules, with the aim of having simple
rules and tightly focused actions.  An action may need to process different
input variants in slightly different ways, but large variation in the
processing performed by an action indicates the need for a new action and a
new category of input; if an action becomes overly complicated it starts to
turn into a monolithic parser, with too much logic contained in a single
procedure.

The ability to easily add special purpose actions to deal with difficulties
and new requirements that are discovered during parser development is one
of the strengths of this architecture.  When a new requirement arises an
independent action can be written to satisfy it, instead of having to
modify a single monolithic function that processes every input, with all
the attendant risks of adversely affecting the existing parsing.  Sometimes
the new action will require the cooperation of other actions, e.g.\ to set
or check a flag, so actions are not always self-contained, but there will
still be a far lower degree of coupling and interdependency than in a
monolithic parser.

During development of \parsername{} it became apparent that in addition to
the obvious variety in log lines there were many complications to overcome.
Some were the result of deficiencies in Postfix's logging, and some of
those deficiencies were rectified by later versions of Postfix, e.g.\
identifying bounce notifications (\sectionref{identifying bounce
notifications}); others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in \parsername{}: adding new actions was enough to overcome
several of the complications; others required modifications to a single
existing action to work around a difficulty; the remainder were resolved by
adapting existing actions to cooperate and exchange extra data (via the
framework), changing their behaviour as appropriate based on that extra
data.  Every architecture should aim to make the easy things easy and the
hard things possible; the successful implementation of \parsername{}
demonstrates that this architecture achieves that aim.

Actions may modify the input they process and return it to the framework,
where it will be parsed as if read from the input stream, allowing for a
simplified version of cascaded parsing~\cite{cascaded-parsing}.  This
powerful facility allows several rules and actions to parse a single input,
potentially simplifying both rules and actions.  A simple example is to
have one rule and action removing comments from inputs, so that other rules
and actions do not have to handle comments at all; obviously if comment
characters can be escaped or embedded in quoted strings the implementation
must be careful not to remove those.  For some inputs this kind of
pre-processing can greatly simplify parsing, echoing the simplification
provided by the framework presenting rules and actions with logical inputs
rather than physical inputs.  A more complex example of cascaded parsing is
evaluating simple arithmetic expressions, where sub-expressions enclosed in
parentheses must be evaluated first; cascaded parsing can be used to parse
and evaluate the sub-expressions, substituting the result into the original
expression for subsequent re-evaluation.  Actions do not need to be
specially registered with the framework or be declared in a particular way
to use cascaded parsing: actions that do not use cascaded parsing will
return nothing, those that do will simply return a string to be parsed.

This section is quite brief, because the architecture deliberately imposes
as few restrictions, conditions, and conventions as possible on actions, to
allow maximum flexibility for parsers based on this architecture.

\section{Rules}

\label{rules in architecture}

Rules are responsible for recognising inputs: each rule should recognise
one and only one input variant; an input category with multiple input
variants should have multiple rules, one for each variant.  Rules will
typically use a regex when recognising inputs, but other approaches may
prove useful for some applications, e.g.\ comparing fixed strings to the
input, or checking the length of the input; for the remainder of this
thesis it will be assumed that a regex is used.  Each rule must specify the
regex to recognise inputs and the action to invoke when recognition is
successful, but implementations are free to add any other attributes they
require; \sectionref{rules table} describes the attributes used in
\parsername{}, and some generally useful attributes will be discussed later
in this section.

Using the rules is simple: the first rule to recognise the input determines
the action that will be invoked; there is no backtracking to try alternate
rules, and no attempt is made to pick a \textit{best\/} rule.
\sectionref{attaching conditions to rules} contains an example which
requires that the rules are used in a specific order to correctly parse the
input, so a mechanism is needed to allow the author of the rules to specify
that ordering.  Each rule can have a priority attribute: when recognising
inputs, the framework should try the rules in the order specified by their
priority attributes, giving the ruleset author fine-grained control over
the order that rules are used in.  The priority attribute may be
implemented as a number, or as a range of values, e.g.\ low, medium, and
high, or in a different fashion entirely if it suits the implementation.
Rule ordering for efficiency is a separate topic that is covered in
\sectionref{rule ordering for efficiency}; overlapping rules are discussed
in \sectionref{overlapping rules in architecture}.

In \acronym{CFG} terms the rules could be described as:

\tab{}$\text{\textless{}input\textgreater{}}~\mapsto{}~\text{rule-1}~|~\text{rule-2}~|~\text{rule-3}~|~\dots~|~\text{rule-n}$

This is not entirely correct because the rules are not truly context free:
rule conditions (described in \sectionref{attaching conditions to rules})
restrict which rules will be used to recognise each input, imposing a
context of sorts.

\subsection{Adding New Rules}

\label{adding new rules in architecture}

The framework issues a warning for each unparsed input, so it is clearly
evident when the ruleset needs to be augmented.  Parsing new inputs is
achieved in one of three ways:

\begin{enumerate}

    \item Modify an existing rule's regex, because the new input is part of
        an existing variant.

    \item Write a new rule that pairs an existing action with a new regex,
        adding a new variant to an existing category.

    \item Create a new category of inputs, write a new action to process
        inputs from the new category, and write a new rule pairing the new
        action with a new regex.

\end{enumerate}

Decoupling the rules from the actions and framework enables other rule
management approaches to be used, e.g.\ instead of manually editing
existing rules or adding new rules, machine learning techniques could be
used to automate the process.  If this approach was taken, the choice of
machine learning technique would be constrained by the size of typical data
sets (see \sectionref{parser efficiency}).  Techniques requiring the full
data set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser could also dynamically create new rules in response to certain
inputs, e.g.\ parsing a subroutine declaration could cause a rule to be
created that parses calls to that subroutine, checking that the arguments
used agree with the subroutine's signature.  These avenues of research and
development have not been pursued by the author, but the architecture
allows them to easily be undertaken independently.

\subsection{Attaching Conditions To Rules}

\label{attaching conditions to rules}

\label{rule conditions in architecture}

Rules can have conditions attached that will be evaluated by the framework
before attempting to use a rule to recognise an input: if the condition is
true the rule will be used, if not the rule will be skipped.  Conditions
can be as simple or complex as the parser requires, though naturally as the
complexity rises so too does the difficulty in understanding how different
rules and actions interact.  The framework has to evaluate each condition,
so as the complexity of conditions increases so will the complexity of the
code required to evaluate them.  Beyond a certain level of complexity,
conditions should probably be written in a proper programming language,
e.g.\ taking advantage of dynamic languages' support for evaluating code at
run-time, or embedding a language like Lua into the parser.  If an
implementation is going to use conditions so complex that they will require
a Turing-complete programming language, the design may need to be
revisited, including the decision to use this architecture --- there may be
other architectures more suitable.

Conditions that only examine the input will be the easiest conditions to
understand, because they can be understood in isolation; they do not depend
on variables set by actions or other rules.  Conditions that examine the
input can be complex if required, but simple conditions can be quite useful
too, e.g.\ every Postfix log line contains the name of the Postfix
component that produced it, so every rule used in \parsername{} has a
condition specifying the component whose log lines it recognises, reducing
the number of rules that will be used when recognising a log line (see
\sectionref{rule conditions in implementation} for details) and increasing
the chance that the log line will be correctly recognised.

Conditions can also check the value of variables that have been set by
either actions or rules; it is easier to understand how a variable's value
will be used and changed if it is set by rules only, rather than by
actions, because the chain of checking and setting variables can be
followed from rule to rule.  The downside to actions setting variables that
are used in rule conditions is action at a distance: understanding when a
rule's condition will be true or false requires understanding not just
every other rule but also every action.  The framework probably does not
need to support a high level of complexity and flexibility when rules are
setting variables; however, if the framework supports complex conditions,
that code can probably be easily extended to support complex variable
assignments too.  The level of complexity the framework supports when
evaluating conditions and setting variables has two costs that must be
taken into account when designing a parser: the difficulty of
implementation, and the difficulty of understanding or writing correct
rules; the flexibility provided by complex conditions may be outweighed by
the difficulty in understanding the interactions between rules that use
them.

An example of how rule conditions can be used is parsing C-style comments,
which start with \texttt{/*} and end with \texttt{*/}; the start and end
tokens can be on one line, or may have many lines between them.
\Tableref{Rules to parse C-style comments} shows the regexes, conditions,
and state changes of the four rules required to parse C-style comments.
These are simplified rules, e.g.\ rules one and two will incorrectly
recognise the comment start token if it is within a quoted string.  Rules
one and two will be used when the parser is parsing code, not comments:
rule one recognises a comment that is contained within one line and leaves
the parser's state unchanged; rule two recognises the start of a comment
and changes the parser's state to parsing comments instead of parsing code.
Rules three and four will be used when the parser is parsing a comment:
rule three recognises the end of a comment and switches the parser's state
back to parsing code; rule four recognises a comment line without an end
token and keeps the parser's state unchanged.  It is important that the
rules are applied in the order listed in \tableref{Rules to parse C-style
comments} because rule two overlaps with rule one, and rule four overlaps
with rule three; \sectionref{rules in architecture} has explained how this
is achieved.  \sectionref{overlapping rules in architecture} will discuss
the benefits and difficulties of using overlapping rules.

\begin{table}[thbp]
    \caption{Rules to parse C-style comments}
    \empty{}\label{Rules to parse C-style comments}
    \begin{tabular}{llll}
        \tabletopline{}%
        No.   & Regex             & Condition                 & Variable Changes          \\
        \tablemiddleline{}%
        1     & \verb!/\*.*?\*/!  & state == parsing code     &                           \\
        2     & \verb!/\*.*!      & state == parsing code     & state = parsing comment   \\
        3     & \verb!.*\*/!      & state == parsing comment  & state = parsing code      \\
        4     & \verb!.*!         & state == parsing comment  &                           \\
        \tablebottomline{}%
    \end{tabular}

    Note that ``state'' is merely a descriptive name for the variable; the
    variable can be called anything at all.

\end{table}

\subsection{Overlapping Rules}

\label{overlapping rules in architecture}

When adding new rules, the rule author must be aware that the new rule may
overlap with one or more existing rules, i.e.\ some inputs could be parsed
by more than one rule.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first rule to
recognise the input wins, and the order in which rules are used might
change between parser invocations.  Overlapping rules are frequently a
requirement, allowing a more specific rule to recognise some inputs and a
more general rule to recognise the remainder, e.g.\ separating
\acronym{SMTP} delivery to specific sites from \acronym{SMTP} delivery to
the rest of the world.  Using overlapping rules simplifies both the general
rule and the more specific rule.  Overlapping rules should have a priority
attribute to specify their relative ordering; negative priorities may be
useful for catchall rules.  The architecture does not try to detect
overlapping rules: that responsibility is left to the author of the rules.

Overlapping rules can be detected by visual inspection, or a program could
be written to analyse the regexes in a ruleset.  Traditional regexes are
equivalent in computational power to \acronym{FA} and can be converted to
\acronym{FA}, so regex overlap can be detected by finding a non-empty
intersection of two \acronym{FA}\@.  Perl 5.10 regexes~\cite{perlre} are
more powerful than traditional regexes: they can match correctly balanced
brackets nested to an arbitrary depth, e.g.\
\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!  matches
\verb!z<123<pq<>rs>j<r>ml>s!.  Matching balanced brackets requires the
regex engine to maintain state on a stack, so Perl 5.10 regexes are
equivalent in computational power to \acronym{PDA}; detecting overlap may
require calculating the intersection of two \acronym{PDA} instead of two
\acronym{FA}.  \acronym{PDA} intersection is not closed, i.e.\ the result
cannot always be implemented using a \acronym{PDA}, so intersection may be
intractable sometimes, e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.
Detecting overlap amongst $n$ regexes requires calculating
$\frac{n\left(n-1\right)}{2}$ intersections, resulting in
$O\left(n^{2}x\right)$ complexity, where $x$ is the cost of calculating
\acronym{FA} or \acronym{PDA} intersection.  This is certainly not a task
to be performed every time a parser runs: naive detection of overlap
amongst \parsernames{} \numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

When detecting overlap, any conditions attached to rules must be taken into
account, because two rules whose regexes overlap may have conditions
attached that prevent the rules overlapping.  A less naive approach to
overlap detection would first check for overlapping conditions amongst
rules, and then check for overlap between the regexes of each pair of rules
with overlapping conditions.  Rule overlap is not transitive, e.g.\ given
these three conditions:

\begin{enumerate}

    % Reduce inter-item spacing because each item is a single line.
    \squeezeitems{}

    \item total $<$ 10

    \item total $>$ 20

    \item total $<$ 30

\end{enumerate}

\noindent{}The first and second conditions do not overlap, but the third
condition overlaps with both the first and second conditions.  When rules
are paired based on how their conditions overlap, the complexity of
detecting overlap amongst $n$ rules is $O\left(n^{2}y+|o|x\right)$, where:

\begin{tabular}[]{rcl}

    $y$ & = & cost of checking for overlap between two conditions \\
    $o$ & = & set of pairs of rules with overlapping conditions   \\
    $x$ & = & cost of checking for overlap between two regexes    \\

\end{tabular}

In the worst case, $|o|$ above will be equal to $n^{2}$.  For this approach
to be more efficient than the naive approach, $y$ must be significantly
lower than $x$.  If $y$ is higher than $x$ then the checks for overlap
should be performed in the opposite order: check for regex overlap first,
then check for condition overlap only between pairs of rules with
overlapping regexes.

Once conditions pass a certain level of complexity, determining if two
conditions overlap becomes intractable, because it requires so much
knowledge of other rules and possibly even actions.  For example, given two
rules with conditions \verb!verbose == true! and \verb!silent == true!,
logically these rules should not overlap, yet there is nothing to stop both
variables being set to true by one or more rules or actions.  If variables
used in conditions can be set by actions, determining if two conditions
overlap is impossible: the Halting Problem shows that it is impossible for
one program to answer questions about another program's behaviour if the
inquiring program is implemented on a computational machine whose power is
equal or less than the power of the computational machine the other program
is implemented upon.

\subsection{Pathological Rules}

It is possible to write pathological regexes, which fall into two main
categories: regexes that match inputs they should not, and regexes that
consume excessive amounts of CPU time during matching.  Defining a regex
that matches inputs it should not is trivial: \verb!/^/! matches the start
of every input.  This regex would be found by a tool that detects
overlapping rules, and would easily be noticed by visual inspection, but
more complex regexes would be harder to find.  Regexes that match inputs
more than they should are a problem not because of excessive resource
usage, but because they may prevent the correct rule from recognising the
input.  If an adaptive ordering system is used to prioritise rules that
frequently recognise inputs (see \sectionref{rule ordering for
efficiency}), then a rule with a regex that matches inputs it should not
may be promoted up through the list, displacing an increasing number of
correct rules as it rises.

Excessive CPU time is usually consumed when a regex fails to match an
input, and the regex engine backtracks many times because of alteration or
nested quantifiers; successful matching is generally quite fast with such
regexes, so problematic regexes may go unnoticed for a long time.  For
example, with most regex engines matching double quoted strings with
\verb!\"([^"\\]+|\\.)*"\!  is very fast when there is a match, but when the
match fails its computational complexity for a string of length $n$ is
$O(2^{n})$; see~\cite{mastering-regular-expressions} for in-depth
discussion of nested quantifiers, backtracking, alteration, and capturing
groups.  Pathological regexes that consume excessive CPU time can be
difficult to detect, whether by visual inspection or by machine inspection,
but if a regex is converted to a \acronym{FA} or the internal
representation used by the regex engine, it may be possible to determine if
nested quantifiers or other troublesome constructs are present.  Modern
regex engines have addressed many of these problems, e.g.\ the regex to
match double quoted strings given above fails immediately with Perl 5.10,
regardless of the input length, because the regex engine looks for both of
the required double quotes first.  Similarly, Perl 5.10's regex engine
optimises alterations starting with literal text into a trie, which has
matching time proportional to the length of the alternatives, rather than
the number of alternatives.  Perl regexes can use \verb!(?>pattern)!, which
matches \verb!pattern! the first time the regex engine passes over it, and
does not change what it originally matched if the regex engine backtracks
over it, alleviating problems caused by excessive backtracking; Prolog
users will notice a similarity to the \verb'!' (cut) operator.  A
presentation showing some of Perl 5.10's new regex features is available at
\urlLastChecked{http://www.regex-engineer.org/slides/perl510_regex.html}{2009/03/03}.

Conditions can vary in complexity from simple equality through to a
Turing-complete language, so enumerating pathological conditions is
difficult if not pointless.  Conditions that check variables or the input
in uncomplicated ways may exhibit unexpected or incorrect behaviour, but
are unlikely to exhibit pathological behaviour.  Deciding if a more complex
condition's behaviour is pathological or simply a bug is difficult and to
some extent is a matter of opinion.  When this architecture has received
more widespread usage, consensus should be reached on the topic of
pathological conditions.

\section{Summary}

This chapter has presented the parser architecture developed for this
project.  It started with a high level view of the architecture, describing
how it achieves its design aim of being easily extensible for users, and
the advantages that being easily extensible brings to parser authors.  The
three main components of the architecture were documented in detail,
explaining each component's responsibilities and the functionality it
provides, plus any difficulties associated with the components.  The
framework provides several support functions and manages the parsing
process, enabling simple rules and actions to be written.  The actions are
simple to understand, because the architecture does not impose any
structure or requirements upon them: parser authors are free to do anything
they want within an action.  The architecture's support for cascaded
parsing was described in the actions section, with an example to illustrate
how it can be useful for general parsing.  The rules section was the
longest section in this chapter, because although the rules appear to be
quite simple --- recognise an input and specify the action to invoke ---
they have subtle behaviour that needs to be clearly explained.  When
extending a ruleset, a decision needs to be taken about whether the input
should be recognised by extending an existing rule, by adding a new rule to
an existing input category, or by adding a new input category, action, and
rule.  Rules can have conditions attached to them, restricting the set of
rules used to recognise an input; the complexity of the conditions used
greatly influences the difficulty of writing a correct ruleset or
understanding and extending an existing ruleset.  Overlapping rules are
frequently a requirement in a parser, and their use can greatly simplify
some rules, but they can be a source of bugs because they can recognise
inputs unexpectedly.  The framework does not try to detect overlapping
rules, because overlap amongst rules may be valid and is quite often
intentional; that responsibility falls to the author of the ruleset.  The
difficulty involved in detecting overlap is proportional to the complexity
of a ruleset's regexes and conditions, and may be possible for a human yet
intractable or impossible for a program.  The rules section concludes with
a discussion of pathological rules, concentrating on pathological regexes.