### 2009-04-28 doc/logparser-glossary.tex

% vim: set filetype=tex :
% The contents of the glossary.
\newglossaryentry{SQLite}{name={SQLite},description={
    \textit{SQLite is a small C library that implements a self-contained,
    embeddable, zero-configuration SQL database engine.\/} SQLite is an
    \acronym{SQL} implementation focusing on correctness, simplicity, and
    speed.  Unlike other \acronym{SQL} implementations it does not require
    a separate server process, greatly simplifying deployment of an
    application utilising it.
}}

\newglossaryentry{Phishing}{name={Phishing},description={
    Phishing~\cite{Wikipedia-phishing} is an attempt to acquire information
    by masquerading as an entity trusted by the user, e.g.\ a bank.
}}

\newglossaryentry{Backscatter}{name={Backscatter},description={
    When a spam sender or virus sends mail with forged sender addresses,
    innocent mail servers are flooded with undeliverable mail notifications
    from badly configured mail servers that either do not verify recipient
    addresses before accepting deliver attempts, or automatically reply to
    mail identified as spam; this is called backscatter.
}}

\newglossaryentry{Joe-job}{name={Joe~Job},description={
    A joe~job describes a large amount of spam mail sent using a faked
    sender address with the intention of sullying the good name of the user
    of that address.  Joe~jobs are one cause of
    backscatter\glsadd{Backscatter}.
}}

\newglossaryentry{Epoch}{name={Epoch},description={
    Most operating systems store the current time and timestamps of files
    as seconds elapsed since the epoch, the beginning of time as far as the
    operating system is concerned.  On Unix and Unix-derived systems the
    epoch is 1970/01/01 00:00:00; on other operating systems it may be
    different.
}}

\newglossaryentry{<>}{name={$<>$},sort={<>},description={
    $<>$ is the sender address used for sending bounce notifications.  In
    the \acronym{SMTP} conversation, all addresses are enclosed in $<>$, so
    \texttt{username@domain} is sent as \texttt{$<$username@domain$>$};
    thus $<>$ is actually an empty address, but it is always written as
    $<>$ for clarity.  Mail servers must not unconditionally reject mail
    sent from $<>$, or they are in violation of
    \acronym{RFC}~2821~\cite{RFC2821}.
}}

\newglossaryentry{queueid}{name={queueid},description={
    Each mail in Postfix's queue is assigned a queueid to uniquely identify
    it.  Queueids are assigned from a limited pool, so although they are
    guaranteed to be unique for the lifetime of the mail they are assigned
    to, they may be reused later.
}}

\newglossaryentry{hash}{name={hash},description={
    A hashing function transforms a string of characters to a number.  A
    common usage is to maintain a data structure indexed by strings in an
    efficient manner.  A full description is beyond the scope of this
    thesis, further information can be found in~\cite{hash-functions}.
}}

\newglossaryentry{awk}{name={AWK},description={
    AWK is a general purpose programming language designed for processing
    text that is available as a standard utility on all Unix systems.
}}

\newglossaryentry{syslog}{name={Syslog},description={
    Syslog is the standard logging mechanism used on Unix systems: a
    program sends log messages to syslog, then syslog filters and stores
    the messages as configured by the administrator.
}}

\newglossaryentry{mail bomb}{name={Mail Bomb},description={
    A mail bomb occurs when an attacker sends a huge volume of mail to a
    victim.  At best a mail bomb is irritating to the victim; at worst the
    deluge of mail can be severe enough to interrupt service for the victim
    or other users, by causing mail to be delayed or rejected because the
    victim or the server has reached a limit, e.g.\ too many mails in their
    mailbox, or too much disk space consumed by the mail queue.
}}

\newglossaryentry{mail loop}{name={Mail Loop},description={
    Sometimes mail sent to one address must be forwarded to a different
    address instead, e.g.\ because a person has changed jobs.  A mail loop
    occurs when the chain of recipient addresses constitute a cyclic
    directed graph; e.g.\ mail for \texttt{alice@example.net} is delivered
    to \texttt{bob@example.com}, and mail for \texttt{bob@example.com} is
    delivered to \texttt{alice@example.net}.  Postfix adds a
    \texttt{Delivered-To:} header to all mails it delivers locally to
    detect mail loops, but in some cases this header is removed by the
    program that forwards the mail.
}}