### 2009-08-24 doc/logparser.tex
Provide a brief introduction to Postfix in the abstract and introduction, as instructed by Howard Hamilton.

% $Id$
\documentclass[a4paper,12pt,draft]{report}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Provides \FloatBarrier that stops floats going past it.
\usepackage{placeins}
% Relax the parameters Latex uses when placing floats.
\usepackage{flexiblefloats}
% Prevent floats from appearing above the line that includes them
\usepackage{flafter}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% College expect one and a half or double spacing for the initial
% submission.
\usepackage{setspace}
\onehalfspacing{}

% Better looking horizontal rules for tables.
\usepackage{booktabs}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

% Aligned list environments.
\usepackage{eqlist}
% This voodoo below gives us an eqlist environment with bold items.
\newcommand{\bolditem}[1]{%
    \bfseries#1%
}
\newenvironment{boldeqlist}
    {\begin{eqlist}[\renewcommand{\makelabel}{\bolditem}]}
    {\end{eqlist}}

% Enable customised lists.
\usepackage{enumitem}
% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

% Extra packages recommended by Springer.
%\usepackage{mathptmx}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{makeidx}
%\usepackage{multicol}

% Embed SVN Id
\usepackage{svn}

\newcommand{\parserblurb}[3]{%
    \begin{quotation}
        \noindent{}\textit{#1\/}

        \noindent{}\url{#2}\newline
        Last checked #3.
    \end{quotation}
}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\newcommand{\mytitle}[0]{%
    Parsing Log Files Produced by the Postfix Mail Transfer Agent%
}
\hypersetup{
    pdftitle    = {\mytitle},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

% Sort numbers where there are multiple citations.  Does not appear to have
% the expected effect (probably clashes with hyperref), though it does
% reduce the space between numbers.
\usepackage{cite}
% Check for unused references, and how the citation key (e.g.\ slct-paper)
% in the margin beside the reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used; neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the default spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[4][thbp]{%
    \begin{figure}[#1]
        \caption{#3}\label{#4}
        \includegraphics{#2}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[4][thbp]{%
    \begin{table}[#1]
        \caption{#3}\label{#4}
        \input{#2}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \toprule{}%
}

\newcommand{\tablebottomline}[0]{%
    \bottomrule{}%
}

\newcommand{\tablemiddleline}[0]{%
    \midrule{}%
}

% A command to format a Postfix daemon's name
\newcommand{\daemon}[1]{%
    \texttt{postfix/#1}%
}

\newcommand{\flowchart}[2]{%
    (action: \action{#1}\@; flowchart:~#2)%
}

\newcommand{\action}[1]{%
    \texttt{#1}%
}

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} \newline{} #1 Last checked #3%
}

\newcommand{\tab}[0]{%
    \hspace*{1em}%
}

% This is silly, but it keeps chktex happy.
\newcommand{\singlequote}[0]{'}%

% Constant values.
\newcommand{\numberOFlogFILES}[0]{%
    93%
}

\newcommand{\numberOFlogFILESall}[0]{%
    774%
}

\newcommand{\numberOFlogFILESallYEARS}[0]{%
    2 years, 1\textonehalf{} months%
}

\newcommand{\numberOFrules}[0]{%
    184%
}

\newcommand{\numberOFrulesMINUSten}[0]{%
    174%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.500\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    60.000\%%
}

\newcommand{\numberOFlogLINES}[0]{%
    60,721,709%
}

\newcommand{\numberOFlogLINEShuman}[0]{%
    60.722 million%
}

\newcommand{\numberOFactions}[0]{%
    23%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    16,836%
}

\newcommand{\numberOFconnectionsINlogFILES}[0]{%
    13,850,793%
}

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}[1][]{\acronym[#1]{PLP}}
\newcommand{\parsernames}[0]{\acronym['s]{PLP}}
\newcommand{\parsernamelong}[0]{Postfix Log Parser}
\newcommand{\parsernameshort}[1][]{PLP#1}

\newcommand{\specialpage}[1]{%
    \phantomsection{}
    \addcontentsline{toc}{chapter}{#1}%
}
\newcommand{\specialpageheading}[1]{%
    \begin{center}
        \textbf{\Large #1}
    \end{center}
}
\newcommand{\specialpageandheading}[1]{%
    \specialpage{#1}
    \specialpageheading{#1}
}

\newcommand{\titleandauthor}[0]{%
\begin{center}

    {\Huge \mytitle{}}

    \vfill{}

    {\LARGE John Tobin}

    \vfill{}

\end{center}
}

\begin{document}

\pagestyle{empty}

\specialpage{Title}

\titleandauthor{}

\begin{center}

    A thesis submitted to the University of Dublin, in fulfilment of the
    requirements for the degree of Master of Science in Computer Science.

    \vfill{}

    April 2009

    \vfill{}

    \vfill{}

    \vfill{}

\end{center}

\newpage{}

% Declaration
\specialpageandheading{Declaration}

\vfill{}

\noindent{}I declare that this thesis, and the work described herein, is
entirely my own work, and has not been submitted as an exercise for a
degree at this or any other university.  This thesis may be borrowed or
copied upon request with the permission of the Librarian, Trinity College,
University of Dublin.

\vfill{}

\begin{flushright}
    \underline{\hspace*{15em}} \\~\\
    John Tobin \\
    \today{}
\end{flushright}

\vfill{}

\vfill{}

\vfill{}

\vfill{}

\newpage{}

% Acknowledgements
\specialpageandheading{Acknowledgements}

\bigskip{}

\noindent{}I am indebted to my supervisor, Dr.\ Carl Vogel, for his advice,
assistance, and guidance.

\bigskip{}

\noindent{}I am grateful to my wife, Ariane Tobin, for her patience, good
will, and encouragement; without her, I would not have accomplished this.

\newpage{}

% Abstract
\specialpage{Abstract}

~\empty{}

\vfill{}
\vfill{}
\vfill{}

%\titleandauthor{}

\begin{center}
    \textbf{Abstract}
\end{center}

Postfix is a widely used Mail Transfer Agent, transferring hundreds of
millions of mails between senders and recipients every day.  Parsing log
files produced by Postfix is much more difficult than it first appears, but
it \textit{is\/} possible to achieve a high degree of accuracy in
understanding Postfix log files, and subsequently in reconstructing the
actions taken by Postfix when processing mail delivery attempts.  This
thesis describes the creation of a parser for Postfix log files,
documenting the architecture developed for this project and the parser that
implements it, the difficulties encountered and the solutions developed.
The parser stores data gleaned from the log files in an SQL database;
future projects or programs could use the gathered data to optimise current
anti-spam measures, to produce statistics showing how effective those
measures are, or to provide a baseline to test new anti-spam measures
against.  The Postfix log file parser needs to be very precise and strict
when parsing, yet must allow users to easily adapt or extend it to parse
new log lines, without requiring that the user have an in-depth knowledge
and understanding of the parser's internal workings.  The newly developed
architecture is designed to make the process of parsing new inputs as
simple as possible, enabling users to trivially add new rules (to parse
variants of known inputs) and relatively easily add new actions (to process
a previously unknown category of inputs).  The parser implemented for this
project is evaluated on the criteria of efficiency and coverage of Postfix
log files, demonstrating that the conflicting goals of efficiency and
accuracy can be balanced, and that one need not be sacrificed to achieve
the other.

\SVN$Id$

\vfill{}
\vfill{}
\vfill{}

~\empty{}

\newpage{}

% Include the header across the top of each page, after the declaration.
\pagestyle{headings}

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\setcounter{page}{5}
\tableofcontents
\listoffigures
\listoftables

\newpage{}

% WSUIPA fonts.
\input{ipamacs}

\input{logparser-introduction}
% Reset the list of acronyms that have been used, so they will be expanded
% again the first time they are used.  The idea is to have them expanded
% once per chapter; I do not reset them before the appendices, but some may
% not have been used since they were reset, and thus will be expanded
% again.
\glsresetall{}
\input{logparser-background}
\glsresetall{}
\input{logparser-other-parsers-review}
\glsresetall{}
\input{logparser-architecture}
\glsresetall{}
\input{logparser-implementation}
\glsresetall{}
\input{logparser-results}
\glsresetall{}
\input{logparser-conclusion}

\appendix

\bibliographystyle{logparser-bibliography-style}
\bibliography{logparser-bibliography}

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\glsresetall{}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\glsresetall{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

\end{document}